---
title: "R Notebook"
output: html_notebook
---

## Load libraries
```{r}
library(tidyverse)
# library(ggtree)
# library(velocyto.R)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
source("../utils/utilFxns.R")
source("../utils/plottingFxns.R")
```

# Periphery
## Load data
Set data directory and load files. 
NOTE: velocyto is not running properly, so use cellranger for now
```{r}
data_dir <- "/Users/joshhahn/Google Drive File Stream/My Drive/shekharlab_data/projects/Macaque/"
M1PerCD90S1 <- Read10X(paste0(data_dir,"MacaqueCD90S1/outs/filtered_gene_bc_matrices/MacFas_5"))
M1PerCD90S2 <- Read10X(paste0(data_dir,"MacaqueCD90S2/outs/filtered_gene_bc_matrices/MacFas_5"))
M1PerCD90S3 <- Read10X(paste0(data_dir,"MacaqueCD90S3/outs/filtered_gene_bc_matrices/MacFas_5"))
M1PerCD90S4 <- Read10X(paste0(data_dir,"MacaqueCD90S4/outs/filtered_gene_bc_matrices/MacFas_5"))
M2PerCD90S1 <- Read10X(paste0(data_dir,"MacaqueCD90S5/outs/filtered_gene_bc_matrices/MacFas_5"))
M2PerCD90S2 <- Read10X(paste0(data_dir,"MacaqueCD90S6/outs/filtered_gene_bc_matrices/MacFas_5"))
M3PerCD90S1 <- Read10X(paste0(data_dir,"MacaqueCD90S7/outs/filtered_gene_bc_matrices/MacFas_5"))
M3PerCD90S2 <- Read10X(paste0(data_dir,"MacaqueCD90S8/outs/filtered_gene_bc_matrices/MacFas_5"))
M3PerCD90S3 <- Read10X(paste0(data_dir,"MacaqueCD90S9/outs/filtered_gene_bc_matrices/MacFas_5"))

M1PerCD73S1 <- Read10X(paste0(data_dir,"MacaqueCD73dpS1/outs/filtered_gene_bc_matrices/MacFas_5"))
M1PerCD73S2 <- Read10X(paste0(data_dir,"MacaqueCD73dpS2/outs/filtered_gene_bc_matrices/MacFas_5"))
M2PerCD73S1 <- Read10X(paste0(data_dir,"MacaqueCD73dpS3/outs/filtered_gene_bc_matrices/MacFas_5"))
M2PerCD73S2 <- Read10X(paste0(data_dir,"MacaqueCD73dpS4/outs/filtered_gene_bc_matrices/MacFas_5"))
M2PerMixedS1 <- Read10X(paste0(data_dir,"MacaquePeripheryMixedS1/outs/filtered_gene_bc_matrices/MacFas_5"))

M4PerCD73S1 <- Read10X(paste0(data_dir,"MacaqueCD73DP2S1/outs/filtered_gene_bc_matrices/MacFas_5"))
M4PerCD73S2 <- Read10X(paste0(data_dir,"MacaqueCD73DP2S2/outs/filtered_gene_bc_matrices/MacFas_5"))

Macaque_mat <- cbind(M1PerCD90S1, M1PerCD90S2, M1PerCD90S3, M1PerCD90S4, M2PerCD90S1, M2PerCD90S2, M3PerCD90S1, M3PerCD90S2, M3PerCD90S3, M1PerCD73S1, M1PerCD73S2, M2PerCD73S1, M2PerCD73S2, M2PerMixedS1, M4PerCD73S1, M4PerCD73S2)
```


Create Seurat object and set file information
```{r}
Macaque <- CreateSeuratObject(Macaque_mat, names.delim = ":")

Macaque@meta.data[colnames(M1PerCD90S1), 'orig.file'] = "MacaqueCD90S1"
Macaque@meta.data[colnames(M1PerCD90S2), 'orig.file'] = "MacaqueCD90S2"
Macaque@meta.data[colnames(M1PerCD90S3), 'orig.file'] = "MacaqueCD90S3"
Macaque@meta.data[colnames(M1PerCD90S4), 'orig.file'] = "MacaqueCD90S4"
Macaque@meta.data[colnames(M2PerCD90S1), 'orig.file'] = "MacaqueCD90S5"
Macaque@meta.data[colnames(M2PerCD90S2), 'orig.file'] = "MacaqueCD90S6"
Macaque@meta.data[colnames(M3PerCD90S1), 'orig.file'] = "MacaqueCD90S7"
Macaque@meta.data[colnames(M3PerCD90S2), 'orig.file'] = "MacaqueCD90S8"
Macaque@meta.data[colnames(M3PerCD90S3), 'orig.file'] = "MacaqueCD90S9"
Macaque@meta.data[colnames(M1PerCD73S1), 'orig.file'] = "MacaqueCD73dpS1"
Macaque@meta.data[colnames(M1PerCD73S2), 'orig.file'] = "MacaqueCD73dpS2"
Macaque@meta.data[colnames(M2PerCD73S1), 'orig.file'] = "MacaqueCD73dpS3"
Macaque@meta.data[colnames(M2PerCD73S2), 'orig.file'] = "MacaqueCD73dpS4"
Macaque@meta.data[colnames(M2PerMixedS1), 'orig.file'] = "MacaquePeripheryMixedS1"
Macaque@meta.data[colnames(M4PerCD73S1), 'orig.file'] = "MacaqueCD73DP2S1"
Macaque@meta.data[colnames(M4PerCD73S2), 'orig.file'] = "MacaqueCD73DP2S2"

Macaque@meta.data[colnames(M1PerCD90S1), 'animal'] = 1
Macaque@meta.data[colnames(M1PerCD90S2), 'animal'] = 1
Macaque@meta.data[colnames(M1PerCD90S3), 'animal'] = 1
Macaque@meta.data[colnames(M1PerCD90S4), 'animal'] = 1
Macaque@meta.data[colnames(M2PerCD90S1), 'animal'] = 2
Macaque@meta.data[colnames(M2PerCD90S2), 'animal'] = 2
Macaque@meta.data[colnames(M3PerCD90S1), 'animal'] = 3
Macaque@meta.data[colnames(M3PerCD90S2), 'animal'] = 3
Macaque@meta.data[colnames(M3PerCD90S3), 'animal'] = 3
Macaque@meta.data[colnames(M1PerCD73S1), 'animal'] = 4
Macaque@meta.data[colnames(M1PerCD73S2), 'animal'] = 4
Macaque@meta.data[colnames(M2PerCD73S1), 'animal'] = 5
Macaque@meta.data[colnames(M2PerCD73S2), 'animal'] = 5
Macaque@meta.data[colnames(M2PerMixedS1), 'animal'] = 5
Macaque@meta.data[colnames(M4PerCD73S1), 'animal'] = 6
Macaque@meta.data[colnames(M4PerCD73S2), 'animal'] = 6

Macaque@meta.data[, 'tag'] = "none"
Macaque@meta.data[colnames(M1PerCD90S1), 'tag'] = "M1PerCD90S1"
Macaque@meta.data[colnames(M1PerCD90S2), 'tag'] = "M1PerCD90S2"
Macaque@meta.data[colnames(M1PerCD90S3), 'tag'] = "M1PerCD90S3"
Macaque@meta.data[colnames(M1PerCD90S4), 'tag'] = "M1PerCD90S4"
Macaque@meta.data[colnames(M2PerCD90S1), 'tag'] = "M2PerCD90S1"
Macaque@meta.data[colnames(M2PerCD90S2), 'tag'] = "M2PerCD90S2"
Macaque@meta.data[colnames(M3PerCD90S1), 'tag'] = "M3PerCD90S1"
Macaque@meta.data[colnames(M3PerCD90S2), 'tag'] = "M3PerCD90S2"
Macaque@meta.data[colnames(M3PerCD90S3), 'tag'] = "M3PerCD90S3"
Macaque@meta.data[colnames(M1PerCD73S1), 'tag'] = "M1PerCD73S1"
Macaque@meta.data[colnames(M1PerCD73S2), 'tag'] = "M1PerCD73S2"
Macaque@meta.data[colnames(M2PerCD73S1), 'tag'] = "M2PerCD73S1"
Macaque@meta.data[colnames(M2PerCD73S2), 'tag'] = "M2PerCD73S2"
Macaque@meta.data[colnames(M2PerMixedS1), 'tag'] = "M2PerMixedS1"
Macaque@meta.data[colnames(M4PerCD73S1), 'tag'] = "M4PerCD73S1"
Macaque@meta.data[colnames(M4PerCD73S2), 'tag'] = "M4PerCD73S2"

saveRDS(Macaque, "../Species_Objects/Macaque_initial.rds")
```

```{r}
VlnPlot(Macaque, features = "nCount_RNA", pt.size = 0, group.by = "tag")
VlnPlot(Macaque, features = "nFeature_RNA", pt.size = 0, group.by = "tag")
```


## Cell class annotation

```{r}
Macaque <- ClusterSeurat(Macaque)
saveRDS(Macaque, file = "../Species_Objects/Macaque_initial.rds")
```

### Visualize initial clusters
```{r}
# Macaque <- readRDS("../Species_Objects/Macaque_initial.rds")
DimPlot(Macaque, label = TRUE)
DimPlot(Macaque, group.by = "tag", cells = sample(colnames(Macaque)))
DimPlot(Macaque, group.by = "animal", cells = sample(colnames(Macaque)))
VlnPlot(Macaque, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Macaque, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

### Look for major retinal cell classes
```{r}
# Macaque <- DendroOrder(Macaque)
Macaque <- UpperCase_genes(Macaque)

RGC_markers= c("RBPMS", "SLC17A6", "POU6F2", "SNCG", "SYT2")
BC_markers=c("VSX2", "CABP5", "OTX2", "PRKCA", "TMEM215")
AC_markers=c("TFAP2A","TFAP2B", "GAD1", "C1QL1", "C1QL2", "SLC6A9")
HC_markers=c("RET", "ONECUT1","ONECUT2", "CALB1", "TPM3", "NRSN1")
PR_markers=c("PDE6H", "CRX", "SAG", "PDC", "RHO", "ARR3")
MG_markers=c("SLC1A3","RLBP1", "GLUL", "APOE")
Other_markers = c("FN1", "GSN","CLDN5","RGS5")

DotPlot(Macaque, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, PR_markers, MG_markers, Other_markers), group.by = "dendro_order") + RotatedAxis()
```

### Annotate clusters using DE genes and Dot plot of marker genes
```{r}
Idents(Macaque) <- "seurat_clusters"
Macaque_genes <- FindAllMarkers(Macaque)
saveRDS(Macaque_genes, file = "../Marker_Genes/Macaque_genes_initial.rds")
```


```{r}
Macaque@meta.data$cell_class = "N/A"
Macaque@meta.data[WhichCells(Macaque, idents =c(9,25,27,21,2,43, 26, 11)),]$cell_class = "RGC"
Macaque@meta.data[WhichCells(Macaque, idents = c(36,12,16,39, 31, 23, 10, 31, 23)),]$cell_class = "BC"
Macaque@meta.data[WhichCells(Macaque, idents = c(13,15,33,38, 7, 29, 34, 24, 32, 35,19,30,40, 1, 20, 0, 5)),]$cell_class = "AC"
Macaque@meta.data[WhichCells(Macaque, idents = c(15, 33)),]$cell_class = "HC"
Macaque@meta.data[WhichCells(Macaque, idents = c(42,28,14,37, 39)),]$cell_class = "PR"
Macaque@meta.data[WhichCells(Macaque, idents = c(4,22,8,45,6,17,18,44)),]$cell_class = "MG"
Macaque@meta.data[WhichCells(Macaque, idents = c(41,3)),]$cell_class = "Other"


DimPlot(Macaque, group.by = "cell_class")

saveRDS(Macaque, "../Species_Objects/Macaque_initial.rds")

```

### Investigate unmarked clusters to see if they have any unique markers
```{r}
# 10 - BP, expresses VSX1
Maca_mark10 <- FindMarkers(Macaque, ident.1 = "10")
head(Maca_mark10, 20)
# 26 - annotate as RGC, likely low quality
Maca_mark26 <- FindMarkers(Macaque, ident.1 = "26")
head(Maca_mark26, 20)
# 39 - PR, top gene is RHO
Maca_mark39 <- FindMarkers(Macaque, ident.1 = "39")
head(Maca_mark39, 20)
# 11 - inconclusive markers, but annotate as RGC based on proximity in dendrogram and UMAP to other RGC clusters
Maca_mark11 <- FindMarkers(Macaque, ident.1 = "11")
head(Maca_mark11, 20)
# 1 -  AC - SLC6A9
Maca_mark1 <- FindMarkers(Macaque, ident.1 = "1")
head(Maca_mark1, 20)
# 20 - AC - SLC6A9
Maca_mark20 <- FindMarkers(Macaque, ident.1 = "20")
# 31 - BC- OTX2
Maca_mark31 <- FindMarkers(Macaque, ident.1 = "31")
head(Maca_mark31, 20)
# 23 - BC - CABP5
Maca_mark23 <- FindMarkers(Macaque, ident.1 = "23")
head(Maca_mark23, 20)
# 0 - AC - related to cluster 5
Maca_mark0 <- FindMarkers(Macaque, ident.1 = "0")
head(Maca_mark0, 20)
# 5 - AC - SLC6A9
Maca_mark5 <- FindMarkers(Macaque, ident.1 = "5")
head(Maca_mark5, 20)
# 3 - other
Maca_mark3 <- FindMarkers(Macaque, ident.1 = "3")
# 41 - other
Maca_mark41 <- FindMarkers(Macaque, ident.1 = "41")
```


### Plot cell type by file
```{r}
Macaque <- readRDS("../Species_Objects/Macaque_initial.rds")
counts <- table(Macaque@meta.data$cell_class, Macaque@meta.data$orig.file)
counts
counts <- t(t(counts) / colSums(counts))
barplot(counts, legend = rownames(counts), col= c("red","orange","yellow","green","blue", "black", "white"))
```

### Collapse positive and negative gene names - unsuccessful
```{r}
# V1
Macaque <- readRDS("../Species_Objects/Macaque_initial.rds")
N_genes_index <- grep("-N$", rownames(Macaque))
N_genes_names <- grep("-N$", rownames(Macaque), value = TRUE)
for(i in 1:length(N_genes_names)){
  N_gene <- N_genes_names[i]
  N_index <- N_genes_index[i]
  P_gene <- gsub("-N$", "-P", N_gene)
  # Add -N and -P counts together in the -N row
  Macaque@assays$RNA@counts[N_gene,] <- Macaque@assays$RNA@counts[N_gene,] + Macaque@assays$RNA@counts[P_gene,] 
  # Rename the -N row
  rownames(Macaque@assays$RNA@counts)[N_index] <- gsub("-N$", "", N_gene)
}

# Remove all -P rows
Macaque@assays$RNA@counts <- Macaque@assays$RNA@counts[-c(grep("-P$", rownames(Macaque))),]

# V2
groupings <- rep(0, length(rownames(Macaque)))
index = 1
new_rownames <- c()

for(i in 1:length(rownames(Macaque)) ){
  # If -N or -P has been detected previously, skip to the next loop call
  if(groupings[i] > 0){
    next()
  }

  gene <- rownames(Macaque)[i]
  groupings[i] <- index
  
  if(substring(gene, first = nchar(gene)-1) == "-N"){
    # Find matching gene, and set it's grouping variable
    groupings[grep(gsub("-N$", "-P", gene), rownames(Macaque))] <- index
    # Update new rownames
    new_rownames <- c(new_rownames, gsub("-N$", "", gene))
  }
  else if(substring(gene, first = nchar(gene)-1) == "-P"){
    # Find matching gene, and set it's grouping variable
    groupings[grep(gsub("-P$", "-N", gene), rownames(Macaque))] <- index
    # Update new rownames
    new_rownames <- c(new_rownames, gsub("-P$", "", gene))
  }
  else{
    # Update new rownames
    new_rownames <- c(new_rownames, gene)
  }
  index <- index + 1
}
    

# V3
Macaque <- readRDS("../Species_Objects/Macaque_initial.rds")
new_mat <- Macaque@assays$RNA@counts[1:2,]
use <- rep(TRUE, length(rownames(Macaque)))
for(i in 1:length(N_genes_names)){
  if(use[i]){
    gene <- rownames(Macaque)[i]
    if(substring(gene, first = nchar(gene)-1) == "-N"){
    # Find matching gene, and skip it later
    use[grep(gsub("-N$", "-P", gene), rownames(Macaque))] <- FALSE
    # Update new rownames
    new_rownames <- c(new_rownames, gsub("-N$", "", gene))
  }
  else if(substring(gene, first = nchar(gene)-1) == "-P"){
    # Find matching gene, and set it's grouping variable
    groupings[grep(gsub("-P$", "-N", gene), rownames(Macaque))] <- index
    # Update new rownames
    new_rownames <- c(new_rownames, gsub("-P$", "", gene))
  }
  else{
    # Update new rownames
    new_rownames <- c(new_rownames, gene)
  }
  N_index <- N_genes_index[i]
  P_gene <- gsub("-N$", "-P", N_gene)
  # Add -N and -P counts together in the -N row
  Macaque@assays$RNA@counts[N_gene,] <- Macaque@assays$RNA@counts[N_gene,] + Macaque@assays$RNA@counts[P_gene,] 
  # Rename the -N row
  rownames(Macaque@assays$RNA@counts)[N_index] <- gsub("-N$", "", N_gene)
}

# Remove all -P rows
Macaque@assays$RNA@counts <- Macaque@assays$RNA@counts[-c(grep("-P$", rownames(Macaque))),]
```

### Collapse -N -P genes
```{r}
# V4
Macaque <- readRDS("../Species_Objects/Macaque_initial.rds")
saveRDS(Macaque, "../Species_Objects/Macaque_initial_uncollapsed.rds")
count_mat <- Macaque@assays$RNA@counts

# Extract rows that do not end in -N or -P
N_genes_index <- grep("-N$", rownames(count_mat))
P_genes_index <- grep("-P$", rownames(count_mat))
good_mat <- count_mat[-c(N_genes_index, P_genes_index),]
bad_mat <- count_mat[c(N_genes_index, P_genes_index),]

# Factor gene names, use to collapse bad_mat
groupings <- as.factor(substr(rownames(bad_mat),1,nchar(rownames(bad_mat))-2))
collapsed_mat <- rowsum(as.data.frame(bad_mat), group = groupings)

# Bind good_mat and collapsed_mat together to create new count matrix
new_mat <- rbind(good_mat, as(as.matrix(collapsed_mat), "sparseMatrix"))

# Generate new object
Macaque_v2 <- CreateSeuratObject(new_mat)
Macaque_v2 <- ClusterSeurat(Macaque_v2)

# Import metadata from old object
Macaque_v2@meta.data <- Macaque@meta.data
saveRDS(Macaque_v2, "../Species_Objects/Macaque_initial.rds")
```



## Import cluster labels
```{r}
meta <- read.csv("../Species_Reference/Macaque_NN_RGC_AC_BC_HC_PR_metadata_3.txt")
meta <- meta[2:dim(meta)[1],]
rownames(meta) <- meta$NAME
meta$NAME <- NULL

# Format velocyto labels to match appropriately
barcode <- function(string){
  tag <- Macaque@meta.data[string, "tag"]
  name <- paste0(tag, "_", string)
  return(name)
}
new_names <- unlist(lapply(colnames(Macaque), barcode))
Macaque@meta.data$barcode <- new_names
Macaque@meta.data$annotated <- meta[Macaque@meta.data$barcode, "Cluster"]

# Remove cells not present in the atlas
Macaque@meta.data$keep <- !is.na(Macaque@meta.data$annotated)
Idents(Macaque) <- "keep"
Macaque <- subset(Macaque, idents = TRUE)
Macaque@meta.data$keep <- NULL
Macaque@meta.data$cell_class <- NULL

saveRDS(Macaque, "../Species_Objects/Macaque_initial_annotated.rds")
```

Separate into RGC objects and BC objects
```{r}
Macaque <- readRDS("../Species_Objects/Macaque_initial_annotated.rds")

Idents(Macaque) <- "annotated"
Macaque_RGC <- subset(Macaque, idents = c(paste0("pRGC", 5:18), "PG_ON", "PG_OFF", "MG_ON", "MG_OFF"))

saveRDS(Macaque_RGC, "../Species_Objects/MacaqueRGC_ann.rds")

Macaque_BC <- subset(Macaque, idents = c("FMB","DB1", "DB2","DB3a","DB3b", "IMB","DB4","BB/GB*","DB6","DB5*", "RB"))

saveRDS(Macaque_BC, "../Species_Objects/MacaqueBC_ann.rds")
```

## Integrate RGC and BC objects

Batch correct Macaque RGCs and run clustering pipeline

```{r}
Macaque_RGC <- readRDS("../Species_Objects/MacaqueRGC_ann.rds")

obj.list <- SplitObject(Macaque_RGC, split.by = "animal")
for (i in 1:length(obj.list)) {
      obj.list[[i]] <- NormalizeData(obj.list[[i]], verbose = FALSE)
      obj.list[[i]] <- FindVariableFeatures(obj.list[[i]], selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}
obj.anchors <- FindIntegrationAnchors(object.list = obj.list)
Macaque_RGC <- IntegrateData(anchorset = obj.anchors)
DefaultAssay(Macaque_RGC) <- "integrated"

Macaque_RGC <- ScaleData(Macaque_RGC, verbose = FALSE)
Macaque_RGC <- RunPCA(Macaque_RGC, npcs = 30, verbose = FALSE)
Macaque_RGC <- RunUMAP(Macaque_RGC, reduction = "pca", dims = 1:30)

DimPlot(Macaque_RGC, group.by = "annotated", label = TRUE)

saveRDS(Macaque_RGC, "~/Species_Objects/MacaqueRGC_int_ann_v1.rds")
```

```{r}
Macaque_BC <- readRDS("../Species_Objects/MacaqueBC_ann.rds")

obj.list <- SplitObject(Macaque_BC, split.by = "animal")
for (i in 1:length(obj.list)) {
      obj.list[[i]] <- NormalizeData(obj.list[[i]], verbose = FALSE)
      obj.list[[i]] <- FindVariableFeatures(obj.list[[i]], selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}
obj.anchors <- FindIntegrationAnchors(object.list = obj.list)
Macaque_BC <- IntegrateData(anchorset = obj.anchors)
DefaultAssay(Macaque_BC) <- "integrated"

Macaque_BC <- ScaleData(Macaque_BC, verbose = FALSE)
Macaque_BC <- RunPCA(Macaque_BC, npcs = 30, verbose = FALSE)
Macaque_BC <- RunUMAP(Macaque_BC, reduction = "pca", dims = 1:30)

DimPlot(Macaque_BC, group.by = "annotated", label = TRUE)

saveRDS(Macaque_BC, "~/Species_Objects/MacaqueBC_int_ann_v1.rds")
```

Convert UMAP and diagonal gene plots to files
```{r}
Macaque_RGC <- readRDS("../Species_Objects/MacaqueRGC_int_ann_v1.rds")
Macaque_BC <- readRDS("../Species_Objects/MacaqueBC_int_ann_v1.rds")

Idents(Macaque_RGC) <- "annotated"
Idents(Macaque_BC) <- "annotated"

pdf("../Figures/Species_UMAPs/MacaqueRGC.pdf", w=4, h=4, useDingbats = FALSE)
DimPlot(Macaque_RGC, label = TRUE) + NoLegend()
dev.off()

pdf("../Figures/Species_UMAPs/MacaqueBC.pdf", w=4, h=4, useDingbats = FALSE)
DimPlot(Macaque_BC, label = TRUE) + NoLegend()
dev.off()
```

# Fovea
### Load data
Set data directory and load files. 
NOTE: velocyto is not running properly, so use cellranger for now
```{r}
data_dir <- "/Users/joshhahn/Google Drive File Stream/My Drive/shekharlab_data/projects/Macaque/"
M1FoveaS1 <- Read10X(paste0(data_dir,"M1Fovea1/MacFas_5"))
colnames(M1FoveaS1) <- paste0("M1FoveaS1:", colnames(M1FoveaS1))
M1FoveaS2 <- Read10X(paste0(data_dir,"M1Fovea2/MacFas_5"))
colnames(M1FoveaS2) <- paste0("M1FoveaS2:", colnames(M1FoveaS2))
M1FoveaS3 <- Read10X(paste0(data_dir,"M1Fovea3/MacFas_5"))
colnames(M1FoveaS3) <- paste0("M1FoveaS3:", colnames(M1FoveaS3))
M1FoveaS4 <- Read10X(paste0(data_dir,"M1Fovea4/MacFas_5"))
colnames(M1FoveaS4) <- paste0("M1FoveaS4:", colnames(M1FoveaS4))
M1FoveaS5 <- Read10X(paste0(data_dir,"M1Fovea5/MacFas_5"))
colnames(M1FoveaS5) <- paste0("M1FoveaS5:", colnames(M1FoveaS5))
M1FoveaS6 <- Read10X(paste0(data_dir,"M1Fovea6/MacFas_5"))
colnames(M1FoveaS6) <- paste0("M1FoveaS6:", colnames(M1FoveaS6))
M1FoveaS7 <- Read10X(paste0(data_dir,"M1Fovea7/MacFas_5"))
colnames(M1FoveaS7) <- paste0("M1FoveaS7:", colnames(M1FoveaS7))
M1FoveaS8 <- Read10X(paste0(data_dir,"M1Fovea8/MacFas_5"))
colnames(M1FoveaS8) <- paste0("M1FoveaS8:", colnames(M1FoveaS8))


M2FoveaS1 <- Read10X(paste0(data_dir,"M2Fovea1/MacFas_5"))
colnames(M2FoveaS1) <- paste0("M2FoveaS1:", colnames(M2FoveaS1))
M2FoveaS2 <- Read10X(paste0(data_dir,"M2Fovea2/MacFas_5"))
colnames(M2FoveaS2) <- paste0("M2FoveaS2:", colnames(M2FoveaS2))
M2FoveaS3 <- Read10X(paste0(data_dir,"M2Fovea3/MacFas_5"))
colnames(M2FoveaS3) <- paste0("M2FoveaS3:", colnames(M2FoveaS3))
M2FoveaS4 <- Read10X(paste0(data_dir,"M2Fovea4/MacFas_5"))
colnames(M2FoveaS4) <- paste0("M2FoveaS4:", colnames(M2FoveaS4))
M2FoveaS5 <- Read10X(paste0(data_dir,"M2Fovea5/MacFas_5"))
colnames(M2FoveaS5) <- paste0("M2FoveaS5:", colnames(M2FoveaS5))
M2FoveaS6 <- Read10X(paste0(data_dir,"M2Fovea6/MacFas_5"))
colnames(M2FoveaS6) <- paste0("M2FoveaS6:", colnames(M2FoveaS6))
M2FoveaS7 <- Read10X(paste0(data_dir,"M2Fovea7/MacFas_5"))
colnames(M2FoveaS7) <- paste0("M2FoveaS7:", colnames(M2FoveaS7))
M2FoveaS8 <- Read10X(paste0(data_dir,"M2Fovea8/MacFas_5"))
colnames(M2FoveaS8) <- paste0("M2FoveaS8:", colnames(M2FoveaS8))

M3FoveaS1 <- Read10X(paste0(data_dir,"M3FoveaS1/MacFas_5"))
colnames(M3FoveaS1) <- paste0("M3FoveaS1:", colnames(M3FoveaS1))
M3FoveaS2 <- Read10X(paste0(data_dir,"M3FoveaS2/MacFas_5"))
colnames(M3FoveaS2) <- paste0("M3FoveaS2:", colnames(M3FoveaS2))
M3FoveaS3 <- Read10X(paste0(data_dir,"M3FoveaS3/MacFas_5"))
colnames(M3FoveaS3) <- paste0("M3FoveaS3:", colnames(M3FoveaS3))

M4FoveaS1 <- Read10X(paste0(data_dir,"M4FoveaS1/MacFas_5"))
colnames(M4FoveaS1) <- paste0("M4FoveaS1:", colnames(M4FoveaS1))
M4FoveaS2 <- Read10X(paste0(data_dir,"M4FoveaS2/MacFas_5"))
colnames(M4FoveaS2) <- paste0("M4FoveaS2:", colnames(M4FoveaS2))
M4FoveaS3 <- Read10X(paste0(data_dir,"M4FoveaS3/MacFas_5"))
colnames(M4FoveaS3) <- paste0("M4FoveaS3:", colnames(M4FoveaS3))


Macaque_mat <- cbind(M1FoveaS1,M1FoveaS2,M1FoveaS3,M1FoveaS4,M1FoveaS5,M1FoveaS6,M1FoveaS7,M1FoveaS8,M2FoveaS1, M2FoveaS2,M2FoveaS3,M2FoveaS4,M2FoveaS5,M2FoveaS6,M2FoveaS7,M2FoveaS8,M3FoveaS1,M3FoveaS2,M3FoveaS3,M4FoveaS1,M4FoveaS2,M4FoveaS3)
```

### Collapse -N -P genes
```{r}
# Extract rows that do not end in -N or -P
N_genes_index <- grep("_n$", rownames(Macaque_mat))
P_genes_index <- grep("_p$", rownames(Macaque_mat))
good_mat <- Macaque_mat[-c(N_genes_index, P_genes_index),]
bad_mat <- Macaque_mat[c(N_genes_index, P_genes_index),]

# Factor gene names, use to collapse bad_mat
groupings <- as.factor(substr(rownames(bad_mat),1,nchar(rownames(bad_mat))-2))
collapsed_mat <- rowsum(as.data.frame(bad_mat), group = groupings)

# Bind good_mat and collapsed_mat together to create new count matrix
Macaque_mat <- rbind(good_mat, as(as.matrix(collapsed_mat), "sparseMatrix"))
```


### Create object
Create Seurat object and set file information
```{r}
Macaque <- CreateSeuratObject(Macaque_mat, names.delim = ":")

Macaque@meta.data[colnames(M1FoveaS1), 'orig.file'] = "M1FoveaS1"
Macaque@meta.data[colnames(M1FoveaS2), 'orig.file'] = "M1FoveaS2"
Macaque@meta.data[colnames(M1FoveaS3), 'orig.file'] = "M1FoveaS3"
Macaque@meta.data[colnames(M1FoveaS4), 'orig.file'] = "M1FoveaS4"
Macaque@meta.data[colnames(M1FoveaS5), 'orig.file'] = "M1FoveaS5"
Macaque@meta.data[colnames(M1FoveaS6), 'orig.file'] = "M1FoveaS6"
Macaque@meta.data[colnames(M1FoveaS7), 'orig.file'] = "M1FoveaS7"
Macaque@meta.data[colnames(M1FoveaS8), 'orig.file'] = "M1FoveaS8"

Macaque@meta.data[colnames(M2FoveaS1), 'orig.file'] = "M2FoveaS1"
Macaque@meta.data[colnames(M2FoveaS2), 'orig.file'] = "M2FoveaS2"
Macaque@meta.data[colnames(M2FoveaS3), 'orig.file'] = "M2FoveaS3"
Macaque@meta.data[colnames(M2FoveaS4), 'orig.file'] = "M2FoveaS4"
Macaque@meta.data[colnames(M2FoveaS5), 'orig.file'] = "M2FoveaS5"
Macaque@meta.data[colnames(M2FoveaS6), 'orig.file'] = "M2FoveaS6"
Macaque@meta.data[colnames(M2FoveaS7), 'orig.file'] = "M2FoveaS7"
Macaque@meta.data[colnames(M2FoveaS8), 'orig.file'] = "M2FoveaS8"

Macaque@meta.data[colnames(M3FoveaS1), 'orig.file'] = "M3FoveaS1"
Macaque@meta.data[colnames(M3FoveaS2), 'orig.file'] = "M3FoveaS2"
Macaque@meta.data[colnames(M3FoveaS3), 'orig.file'] = "M3FoveaS3"

Macaque@meta.data[colnames(M4FoveaS1), 'orig.file'] = "M4FoveaS1"
Macaque@meta.data[colnames(M4FoveaS2), 'orig.file'] = "M4FoveaS2"
Macaque@meta.data[colnames(M4FoveaS3), 'orig.file'] = "M4FoveaS3"

saveRDS(Macaque, "../Species_Objects/MacaqueFovea_initial.rds")
```

### Import cluster labels
```{r}
meta <- read.csv("../Species_Reference/Macaque_NN_RGC_AC_BC_HC_PR_metadata_3.txt")
meta <- meta[2:dim(meta)[1],]
rownames(meta) <- meta$NAME
meta$NAME <- NULL

Macaque@meta.data$barcode <- str_replace_all(str_remove_all(colnames(Macaque), "S"), ":", "_")
Macaque@meta.data$annotated <- meta[Macaque@meta.data$barcode, "Cluster"]

# Remove cells not present in the atlas
Macaque@meta.data$keep <- !is.na(Macaque@meta.data$annotated)
Idents(Macaque) <- "keep"
Macaque <- subset(Macaque, idents = TRUE)
Macaque@meta.data$keep <- NULL
Macaque@meta.data$cell_class <- NULL

saveRDS(Macaque, "../Species_Objects/MacaqueFovea_initial_annotated.rds")
```

### Separate into RGC objects and BC objects
```{r}
Macaque <- readRDS("../Species_Objects/MacaqueFovea_initial_annotated.rds")

Idents(Macaque) <- "annotated"
Macaque_RGC <- subset(Macaque, idents = c(paste0("fRGC", 5:16), "PG_ON", "PG_OFF", "MG_ON", "MG_OFF"))

saveRDS(Macaque_RGC, "../Species_Objects/MacaqueFovea_RGC_ann.rds")

Macaque_BC <- subset(Macaque, idents = c("FMB","DB1","OFFx", "DB2","DB3a","DB3b", "IMB","DB4","BB/GB*","DB6","DB5*", "RB"))

saveRDS(Macaque_BC, "../Species_Objects/MacaqueFovea_BC_ann.rds")
```



## Rods
```{r}
Macaque <- readRDS("../Species_Objects/Macaque_initial_annotated.rds")

Idents(Macaque) <- "annotated"
Macaque_Rods <- subset(Macaque, idents = c("Rods"))
Macaque_Rods <- ClusterSeurat(Macaque_Rods)
saveRDS(Macaque_Rods, file = "../Species_Objects/MacaqueRods_v1.rds")
```



# Velocyto Data

## Fovea
### Load velocyto data
```{r}
# Load test data
data_dir <- "/Users/joshhahn/Google Drive File Stream/My Drive/shekharlab_data/projects/Macaque_v2/"

Macaque_mat <- NULL

for(i in c(1:2, 4:8)){
  file_dir <- paste0(data_dir, "M1Fovea", i, "/velo_outs/")
  #Convert(paste0(file_dir, "full.h5ad"), dest = "h5seurat")
  file_obj <- LoadH5Seurat(paste0(file_dir, "full.h5seurat"))
  count_mat <- file_obj@assays$matrix@counts
  colnames(count_mat) <- paste0("M1Fovea", i, substr(colnames(count_mat), 27, 44))
  if(is.null(Macaque_mat)){
    Macaque_mat <- count_mat
  }
  else{
    Macaque_mat <- cbind(Macaque_mat, count_mat)
  }
}

for(i in 1:8){
  file_dir <- paste0(data_dir, "M2Fovea", i, "/velo_outs/")
  #Convert(paste0(file_dir, "full.h5ad"), dest = "h5seurat")
  file_obj <- LoadH5Seurat(paste0(file_dir, "full.h5seurat"))
  count_mat <- file_obj@assays$matrix@counts
  colnames(count_mat) <- paste0("M2Fovea", i, substr(colnames(count_mat), 27, 44))
  Macaque_mat <- cbind(Macaque_mat, count_mat)
}

for(i in 1:3){
  file_dir <- paste0(data_dir, "M3Fovea", i, "/velo_outs/")
  #Convert(paste0(file_dir, "full.h5ad"), dest = "h5seurat")
  file_obj <- LoadH5Seurat(paste0(file_dir, "full.h5seurat"))
  count_mat <- file_obj@assays$matrix@counts
  colnames(count_mat) <- paste0("M3Fovea", i, substr(colnames(count_mat), 27, 44))
  Macaque_mat <- cbind(Macaque_mat, count_mat)
}

for(i in 1:3){
  file_dir <- paste0(data_dir, "M4Fovea", i, "/velo_outs/")
  #Convert(paste0(file_dir, "full.h5ad"), dest = "h5seurat")
  file_obj <- LoadH5Seurat(paste0(file_dir, "full.h5seurat"))
  count_mat <- file_obj@assays$matrix@counts
  colnames(count_mat) <- paste0("M4Fovea", i, substr(colnames(count_mat), 27, 44))
  Macaque_mat <- cbind(Macaque_mat, count_mat)
}

Macaque_Fovea <- CreateSeuratObject(Macaque_mat, names.delim = ":")
saveRDS(Macaque_Fovea, "../Species_Objects/MacaqueFovea_Velocyto_initial.rds")
```



### Create Object
```{r}

Macaque_Fovea <- readRDS("../Species_Objects/MacaqueFovea_Velocyto_initial.rds")

meta <- read.csv("../Species_Reference/Macaque_NN_RGC_AC_BC_HC_PR_metadata_3.txt")
meta <- meta[2:dim(meta)[1],]
rownames(meta) <- meta$NAME
meta$NAME <- NULL

Macaque_Fovea@meta.data$barcode <- str_replace_all(str_replace_all(colnames(Macaque_Fovea), ":", "_"), "x", "-1")
Macaque_Fovea@meta.data$annotated <- meta[Macaque_Fovea@meta.data$barcode, "Cluster"]

# Remove cells not present in the atlas
Macaque_Fovea@meta.data$keep <- !is.na(Macaque_Fovea@meta.data$annotated)
Idents(Macaque_Fovea) <- "keep"
Macaque_Fovea <- subset(Macaque_Fovea, idents = TRUE)
Macaque_Fovea@meta.data$keep <- NULL
Macaque_Fovea@meta.data$cell_class <- NULL

saveRDS(Macaque_Fovea, "../Species_Objects/MacaqueFovea_Velo_int_ann.rds")
Macaque_Fovea <- readRDS("../Species_Objects/MacaqueFovea_Velo_int_ann.rds")
```

### Format Cell class
```{r}
Macaque_Fovea <- readRDS("../Species_Objects/MacaqueFovea_Velo_int_ann.rds")
Idents(Macaque_Fovea) <- "annotated"

Macaque_Fovea@meta.data$cell_class <- "Other"


Macaque_Fovea@meta.data[WhichCells(Macaque_Fovea, idents = c(paste0("fRGC", 5:16), "PG_ON", "PG_OFF", "MG_ON", "MG_OFF")), "cell_class"] <- "RGC"
Macaque_Fovea@meta.data[WhichCells(Macaque_Fovea, idents = c("FMB","DB1","OFFx", "DB2","DB3a","DB3b", "IMB","DB4","BB/GB*","DB6","DB5*", "RB")), "cell_class"] <- "BP"

Macaque_Fovea@meta.data[WhichCells(Macaque_Fovea, idents = paste0("fGa", 1:19)), "cell_class"] <- "GabaAC"
Macaque_Fovea@meta.data[WhichCells(Macaque_Fovea, idents = paste0("fGl", 1:8)), "cell_class"] <- "GlyAC"
Macaque_Fovea@meta.data[WhichCells(Macaque_Fovea, idents = c("H1", "H2")), "cell_class"] <- "HC"
Macaque_Fovea@meta.data[WhichCells(Macaque_Fovea, idents = "MG"), "cell_class"] <- "MG"
Macaque_Fovea@meta.data[WhichCells(Macaque_Fovea, idents = "Mic"), "cell_class"] <- "MicroG"
Macaque_Fovea@meta.data[WhichCells(Macaque_Fovea, idents = c("M/L Cones", "S Cones")), "cell_class"] <- "Cone"
Macaque_Fovea@meta.data[WhichCells(Macaque_Fovea, idents = "Rods"), "cell_class"] <- "Rod"

Macaque_Fovea@meta.data[WhichCells(Macaque_Fovea, idents = c("Endo", "Pericytes")), "cell_class"] <- "Other"

saveRDS(Macaque_Fovea, "../Species_Initial/MacaqueFovea_initial.rds")
saveRDS(Macaque_Fovea@meta.data, "../Metadata/MacaqueFovea_metadata.rds")
```

### Separate into RGC / BC
Batch correction was not performed, as downstream tasks use just the raw RGC counts
```{r}
Idents(Macaque_Fovea) <- "cell_class"

M_Fov_RGC <- subset(Macaque_Fovea, idents = "RGC")
M_Fov_RGC <- ClusterSeurat(M_Fov_RGC)
Idents(M_Fov_RGC) <- "annotated"
saveRDS(M_Fov_RGC, "../Species_Objects/MacaqueFovea_RGC_velo_ann_v1.rds")
  
M_Fov_BC <- subset(Macaque_Fovea, idents = "BP")
M_Fov_BC <- ClusterSeurat(M_Fov_BC)
saveRDS(M_Fov_BC, "../Species_Objects/MacaqueFovea_BC_velo_ann_v1.rds")



```




## Periphery
### Load velocyto data
```{r}
# Load test data
data_dir <- "/Users/joshhahn/Google Drive/My Drive/shekharlab_data/projects/Macaque_v2/"

Macaque_mat <- NULL

for(i in c(1:9)){
  file_dir <- paste0(data_dir, "MacaqueCD90S", i, "/velo_outs/")
  #Convert(paste0(file_dir, "full.h5ad"), dest = "h5seurat")
  file_obj <- LoadH5Seurat(paste0(file_dir, "full.h5seurat"))
  count_mat <- file_obj@assays$matrix@counts
  colnames(count_mat) <- paste0("MacaqueCD90S", i, substr(colnames(count_mat), 27, 44))
  if(is.null(Macaque_mat)){
    Macaque_mat <- count_mat
  }
  else{
    Macaque_mat <- cbind(Macaque_mat, count_mat)
  }
}

for(i in 1:3){
  file_dir <- paste0(data_dir, "MacaqueCD73S", i, "/velo_outs/")
  #Convert(paste0(file_dir, "full.h5ad"), dest = "h5seurat")
  file_obj <- LoadH5Seurat(paste0(file_dir, "full.h5seurat"))
  count_mat <- file_obj@assays$matrix@counts
  colnames(count_mat) <- paste0("MacaqueCD73S", i, substr(colnames(count_mat), 27, 44))
  Macaque_mat <- cbind(Macaque_mat, count_mat)
}

for(i in 1:2){
  file_dir <- paste0(data_dir, "MacaqueCD73DP2S", i, "/velo_outs/")
  #Convert(paste0(file_dir, "full.h5ad"), dest = "h5seurat")
  file_obj <- LoadH5Seurat(paste0(file_dir, "full.h5seurat"))
  count_mat <- file_obj@assays$matrix@counts
  colnames(count_mat) <- paste0("MacaqueCD73DP2S", i, substr(colnames(count_mat), 27, 44))
  Macaque_mat <- cbind(Macaque_mat, count_mat)
}


file_dir <- paste0(data_dir, "MacaquePeripheryMixedS1", "/velo_outs/")
#Convert(paste0(file_dir, "full.h5ad"), dest = "h5seurat")
file_obj <- LoadH5Seurat(paste0(file_dir, "full.h5seurat"))
count_mat <- file_obj@assays$matrix@counts
colnames(count_mat) <- paste0("MacaquePeripheryMixedS1", substr(colnames(count_mat), 27, 44))
Macaque_mat <- cbind(Macaque_mat, count_mat)
```

### Create Object
```{r}
Macaque_Peri <- CreateSeuratObject(Macaque_mat, names.delim = ":")
saveRDS(Macaque_Peri, "../Species_Objects/MacaquePeriphery_Velocyto_initial.rds")

Macaque_Peri <- readRDS("../Species_Objects/MacaquePeriphery_Velocyto_initial.rds")

Idents(Macaque_Peri) <- "orig.ident"
Macaque_Peri@meta.data$tag <- NA
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD90S1"), "tag"] <- "M1PerCD90S1"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD90S2"), "tag"] <- "M1PerCD90S2"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD90S3"), "tag"] <- "M1PerCD90S3"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD90S4"), "tag"] <- "M1PerCD90S4"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD90S5"), "tag"] <- "M2PerCD90S1"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD90S6"), "tag"] <- "M2PerCD90S2"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD90S7"), "tag"] <- "M3PerCD90S1"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD90S8"), "tag"] <- "M3PerCD90S2"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD90S9"), "tag"] <- "M3PerCD90S3"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD73S1"), "tag"] <- "M1PerCD73S1"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD73S2"), "tag"] <- "M1PerCD73S2"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD73S3"), "tag"] <- "M2PerCD73S1"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD73DP2S1"), "tag"] <- "M4PerCD73S1"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaqueCD73DP2S2"), "tag"] <- "M4PerCD73S2"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MacaquePeripheryMixedS1"), "tag"] <- "M2PerMixedS1"

Macaque_Peri@meta.data$code <- substr(colnames(Macaque_Peri), nchar(colnames(Macaque_Peri))-16, nchar(colnames(Macaque_Peri))-1)
Macaque_Peri@meta.data$barcode <- paste0(Macaque_Peri@meta.data$tag, "_", Macaque_Peri@meta.data$code, "-1")


meta <- read.csv("../Species_Reference/Macaque_NN_RGC_AC_BC_HC_PR_metadata_3.txt")
meta <- meta[2:dim(meta)[1],]
rownames(meta) <- meta$NAME
meta$NAME <- NULL

Macaque_Peri@meta.data$annotated <- meta[Macaque_Peri@meta.data$barcode, "Cluster"]

# Remove cells not present in the atlas
Macaque_Peri@meta.data$keep <- !is.na(Macaque_Peri@meta.data$annotated)
Idents(Macaque_Peri) <- "keep"
Macaque_Peri <- subset(Macaque_Peri, idents = TRUE)
Macaque_Peri@meta.data$keep <- NULL
Macaque_Peri@meta.data$cell_class <- NULL

saveRDS(Macaque_Peri, "../Species_Objects/MacaquePeriphery_Velo_int_ann.rds")
```


### Format Cell class
```{r}
Macaque_Peri <- readRDS("../Species_Objects/MacaquePeriphery_Velo_int_ann.rds")
Idents(Macaque_Peri) <- "annotated"

Macaque_Peri@meta.data$cell_class <- "Other"


Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = c(paste0("pRGC", 5:18), "PG_ON", "PG_OFF", "MG_ON", "MG_OFF")), "cell_class"] <- "RGC"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = c("FMB","DB1", "DB2","DB3a","DB3b", "IMB","DB4","BB/GB*","DB6","DB5*", "RB")), "cell_class"] <- "BP"

Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = paste0("pGa", 1:26)), "cell_class"] <- "GabaAC"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = paste0("pGl", 1:8)), "cell_class"] <- "GlyAC"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = c("H1", "H2")), "cell_class"] <- "HC"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "MG"), "cell_class"] <- "MG"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "Mic"), "cell_class"] <- "MicroG"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = c("M/L Cones")), "cell_class"] <- "Cone"
Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = "Rods"), "cell_class"] <- "Rod"

Macaque_Peri@meta.data[WhichCells(Macaque_Peri, idents = c("Endo", "Pericytes")), "cell_class"] <- "Other"

saveRDS(Macaque_Peri, "../Species_Initial/MacaquePeriphery_initial.rds")
saveRDS(Macaque_Peri@meta.data, "../Metadata/MacaquePeriphery_metadata.rds")
```

### Separate into RGC / BC
Batch correction was not performed, as downstream tasks use just the raw RGC counts
```{r}
Idents(Macaque_Peri) <- "cell_class"

M_Per_RGC <- subset(Macaque_Peri, idents = "RGC")
M_Per_RGC <- ClusterSeurat(M_Per_RGC)
M_Per_RGC@meta.data$type <- M_Per_RGC@meta.data$annotated
Idents(M_Per_RGC) <- "type"
M_Per_RGC@meta.data$type <- factor(M_Per_RGC@meta.data$type, unique(M_Per_RGC@meta.data$type))

saveRDS(M_Per_RGC, "../Species_Objects/MacaquePeri_RGC_velo_ann_v2.rds")
saveRDS(M_Per_RGC@meta.data, "../Metadata/RGC/MacaquePeri_RGC_velo_ann_v2_metadata.rds")
  
M_Per_BC <- subset(Macaque_Peri, idents = "BP")
M_Per_BC <- ClusterSeurat(M_Per_BC)
Idents(M_Per_BC) <- "annotated"
saveRDS(M_Per_BC, "../Species_Objects/MacaquePeri_BC_velo_ann_v1.rds")
```

# Combine fovea and periphery into full Macaque object
```{r}
Fovea <- readRDS("../Species_Initial/MacaqueFovea_initial.rds")
Periphery <- readRDS("../Species_Initial/MacaquePeriphery_initial.rds")
Full <- merge(Fovea, Periphery)

saveRDS(Full, "../Species_Initial/Macaque_initial.rds")
saveRDS(Full@meta.data, "../Metadata/Macaque_metadata.rds")
```



```{r}
object <- readRDS("../Species_Objects/MacaqueFovea_initial_annotated.rds")

numPCs = 30
object <- NormalizeData(object)
object <- FindVariableFeatures(object)
object <- ScaleData(object)

object <- RunPCA(object)
object <- RunUMAP(object, dims = 1:numPCs)

saveRDS(object, "../Species_Objects/MacaqueFovea_initial_annotated.rds")
```

```{r}
object <- readRDS("../Species_Objects/MacaqueFovea_initial_annotated.rds")

numPCs = 30
object <- NormalizeData(object)
object <- FindVariableFeatures(object)
object <- ScaleData(object)

object <- RunPCA(object)
object <- FindNeighbors(object, dims = 1:numPCs)
object <- RunTSNE(object, dims = 1:numPCs)
object <- RunUMAP(object, dims = 1:numPCs)

saveRDS(object, "../Species_Objects/MacaqueFovea_initial_annotated.rds")
```


```{r}
object <- readRDS("../Species_Objects/MacaqueFovea_Velo_int_ann.rds")

numPCs = 30
object <- NormalizeData(object)
object <- FindVariableFeatures(object)
object <- ScaleData(object)

object <- RunPCA(object)
object <- RunUMAP(object, dims = 1:numPCs)

saveRDS(object, "../Species_Objects/MacaqueFovea_initial_annotated.rds")

```

```{r}
Macaque <- readRDS("../Species_Objects//Macaque_initial_annotated.rds")
Mac_Fov <- readRDS("../Species_Initial/Ma")
```


