---
title: "R Notebook"
output: html_notebook
---

# Preprocessing 

Load libraries
```{r}
library(reshape2)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
source("../utils/utilFxns.R")
source("../utils/plottingFxns.R")
```

Set data directory and load loom files
```{r}
data_dir <- "/Users/joshhahn/Google Drive/My Drive/shekharlab_data/projects/Opossum/"

#Convert(paste0(data_dir, "Oppasum1_neun_neg_1/velo_outs/full.h5ad"), dest = "h5seurat")
O1_NN1 <- LoadH5Seurat(paste0(data_dir, "Oppasum1_neun_neg_1/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "Oppasum1_neun_neg_2/velo_outs/full.h5ad"), dest = "h5seurat")
O1_NN2 <- LoadH5Seurat(paste0(data_dir, "Oppasum1_neun_neg_2/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "Oppasum1_neun_pos/velo_outs/full.h5ad"), dest = "h5seurat")
O1_Npos <- LoadH5Seurat(paste0(data_dir, "Oppasum1_neun_pos/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "Oppasum2_neun_neg_1/velo_outs/full.h5ad"), dest = "h5seurat")
O2_NN1 <- LoadH5Seurat(paste0(data_dir, "Oppasum2_neun_neg_1/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "Oppasum2_neun_neg_2/velo_outs/full.h5ad"), dest = "h5seurat")
O2_NN2 <- LoadH5Seurat(paste0(data_dir, "Oppasum2_neun_neg_2/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "Oppasum2_neun_pos/velo_outs/full.h5ad"), dest = "h5seurat")
O2_Npos <- LoadH5Seurat(paste0(data_dir, "Oppasum2_neun_pos/velo_outs/full.h5seurat"))




#Convert(paste0(data_dir, "op_neunneg_B4/velo_outs/full.h5ad"), dest = "h5seurat")
nn_b4 <- LoadH5Seurat(paste0(data_dir, "op_neunneg_B4/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "op_neunpos_B2/velo_outs/full.h5ad"), dest = "h5seurat")
np_b2 <- LoadH5Seurat(paste0(data_dir, "op_neunpos_B2/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "op_neunpos_B5/velo_outs/full.h5ad"), dest = "h5seurat")
np_b5 <- LoadH5Seurat(paste0(data_dir, "op_neunpos_B5/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "OpNeuNs1/velo_outs/full.h5ad"), dest = "h5seurat")
OpN1 <- LoadH5Seurat(paste0(data_dir, "OpNeuNs1/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "OpNeuNs2/velo_outs/full.h5ad"), dest = "h5seurat")
OpN2 <- LoadH5Seurat(paste0(data_dir, "OpNeuNs2/velo_outs/full.h5seurat"))



```

Combine matrices and create new Seurat object. 
```{r}
Opossum_mat <- cbind(O1_NN1@assays$matrix@counts, O1_NN2@assays$matrix@counts, O1_Npos@assays$matrix@counts, O2_NN1@assays$matrix@counts, O2_NN2@assays$matrix@counts, O2_Npos@assays$matrix@counts, nn_b4@assays$matrix@counts, np_b2@assays$matrix@counts, np_b5@assays$matrix@counts, OpN1@assays$matrix@counts, OpN2@assays$matrix@counts)

Opossum <- CreateSeuratObject(Opossum_mat, names.delim = ":")

Opossum@meta.data[colnames(O1_NN1), 'orig.file'] = "Oppasum1_neun_neg_1"
Opossum@meta.data[colnames(O1_NN2), 'orig.file'] = "Oppasum1_neun_neg_2"
Opossum@meta.data[colnames(O1_Npos), 'orig.file'] = "Oppasum1_neun_pos"
Opossum@meta.data[colnames(O2_NN1), 'orig.file'] = "Oppasum2_neun_neg_1"
Opossum@meta.data[colnames(O2_NN2), 'orig.file'] = "Oppasum2_neun_neg_2"
Opossum@meta.data[colnames(O2_Npos), 'orig.file'] = "Oppasum2_neun_pos"
Opossum@meta.data[colnames(np_b2), 'orig.file'] = "op_neunpos_B2"
Opossum@meta.data[colnames(np_b5), 'orig.file'] = "op_neunpos_B5"
Opossum@meta.data[colnames(nn_b4), 'orig.file'] = "op_neunneg_B4"
Opossum@meta.data[colnames(OpN1), 'orig.file'] = "OpNeuNs1"
Opossum@meta.data[colnames(OpN2), 'orig.file'] = "OpNeuNs2"



```


```{r}
VlnPlot(Opossum, features = "nCount_RNA", pt.size = 0, group.by = "orig.file")
VlnPlot(Opossum, features = "nFeature_RNA", pt.size = 0, group.by = "orig.file")
```

```{r}
saveRDS(Opossum, file = "../Species_Objects/Opossum_initial.rds")
```

# Cell class annotation

```{r}
#Opossum <- readRDS("../Species_Objects/Opossum_initial.rds")
Opossum <- ClusterSeurat(Opossum, numPCs = 20, cluster_resolution = 0.3)
saveRDS(Opossum, file = "../Species_Objects/Opossum_initial.rds")
```


Visualize initial clusters
```{r}
DimPlot(Opossum, label = TRUE)
DimPlot(Opossum, group.by = "orig.file")
VlnPlot(Opossum, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Opossum, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```


Look for major retinal cell classes
```{r}
Opossum <- DendroOrder(Opossum)

RGC_markers= c("ENSMODG00000019006", "SLC17A6", "POU6F2", "THY1", "NEFL", "NEFM", "SNCG")

# 6374 is Vsx2, 20977 is Grik1, 
BC_markers= c("VSX2", "OTX2", "TRPM1", "ENSMODG00000006374", "CABP5", "ENSMODG00000020977", "ENSMODG00000004147", "GRM6")

# Tfap2a is 10299, Tfap2b is 18831, Tfap2x is 16445, GAD1 is 8529, GAD2 is 8626
AC_markers= c("ENSMODG00000010299" ,"ENSMODG00000018831", "ENSMODG00000016445", "ENSMODG00000008529","ENSMODG00000008626", "SLC6A9", "C1QL1", "C1QL2")

# Onecut1 - 3 is 1801, 20448, 4149
HC_markers= c("ENSMODG00000001801", "ENSMODG00000020448","ENSMODG00000004149", "LHX1", "CALB1", "TPM3")
Cone_markers= c("PDE6H", "CRX", "ARR3")
Rod_markers= c("SAG", "PDC", "ENSMODG00000006488")
MG_markers= c("SLC1A3","RLBP1", "APOE")
Other_markers= c("S100B", "GFAP", "C1QA", "C1QB", "CLDN5", "IGFBP7")

DotPlot(Opossum, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), group.by = "dendro_order") + RotatedAxis()
```

```{r}
plot(Opossum@tools$BuildClusterTree)
```

Annotate clusters using DE genes and Dot plot of marker genes
```{r}
old_Opossum <- readRDS("../Species_Initial/Opossum_initial.rds")
Opossum@meta.data$prev_class = "N/A"
Opossum@meta.data[rownames(old_Opossum@meta.data), "prev_class"] <- old_Opossum@meta.data$cell_class
table(Opossum@meta.data$dendro_order, Opossum@meta.data$prev_class)
```

Annotate clusters using DE genes and Dot plot of marker genes

```{r}
Opossum@meta.data$cell_class = "Nope"
Opossum@meta.data[WhichCells(Opossum, idents =c(42,1,9,15,2,27)),]$cell_class = "RGC"
Opossum@meta.data[WhichCells(Opossum, idents = c(6,38,23,24,25,28)),]$cell_class = "BP"
Opossum@meta.data[WhichCells(Opossum, idents = c(32)),]$cell_class = "HC"
Opossum@meta.data[WhichCells(Opossum, idents = c(41,10)),]$cell_class = "MG"
Opossum@meta.data[WhichCells(Opossum, idents = c(5,36,11,8,20)),]$cell_class = "GlyAC"
Opossum@meta.data[WhichCells(Opossum, idents = c(12,22,33,4,37,31,19,3,7,40,39,26,16,18,34,13,14)),]$cell_class = "GabaAC"
Opossum@meta.data[WhichCells(Opossum, idents = c(0,17)),]$cell_class = "Rod"
Opossum@meta.data[WhichCells(Opossum, idents = c(21)),]$cell_class = "Cone" 
Opossum@meta.data[WhichCells(Opossum, idents = c(29,30,35)),]$cell_class = "Other"


```


Investigate unmarked clusters to see if they have any unique markers
```{r}
Mouse <- readRDS("../Species_Initial/Mouse_initial.rds")

ortho3[grep("Tfap", ortho3$Gene.name), c("Opossum.gene.name", "Opossum.gene.stable.ID")]


# 
ident = c(26,31)
mark <- FindMarkers(Opossum, ident.1 = ident, max.cells.per.ident = 1000)
ortho3[ortho3$Opossum.gene.stable.ID %in% head(rownames(mark),10), "Gene.name"]

```

## Update gene names
```{r}
ortho <- read.csv("../Orthology/mart_Opossum.txt", sep = "\t")
ortho <- ortho[, c("Opossum.gene.stable.ID", "Opossum.gene.name")]
ortho <- ortho[!duplicated(ortho), ]
ortho <- ortho[ortho$Opossum.gene.name != "", ]
ortho <- ortho[!duplicated(ortho$Opossum.gene.name),]
rownames(ortho) <- ortho$Opossum.gene.stable.ID
```



```{r}
replace_index <- rownames(Opossum) %in% rownames(ortho)
replace_id <- rownames(Opossum)[replace_index]
rownames(Opossum@assays$RNA@counts)[replace_index] <- ortho[replace_id, "Opossum.gene.name"]
rownames(Opossum@assays$RNA@data)[replace_index] <- ortho[replace_id, "Opossum.gene.name"]

```


```{r}
DimPlot(Opossum, group.by = "cell_class")

counts <- table(Opossum@meta.data$cell_class, Opossum@meta.data$orig.file)
counts
saveRDS(Opossum, "../Species_Initial/Opossum_initial.rds")
saveRDS(Opossum@meta.data, "../Metadata/Opossum_metadata.rds")

```




# RGC clustering analysis
Create separate RGC object
```{r}
Opossum <- readRDS("../Species_Objects/Opossum_initial.rds")

# Only keep RGCs
Idents(Opossum) <- "cell_class"
Opossum_RGC <- subset(Opossum, cells = WhichCells(Opossum, idents ="RGC"))

# Remove cells with low and high counts or features
VlnPlot(Opossum_RGC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Opossum_RGC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

```


```{r}
Opossum_RGC <- subset(Opossum_RGC, orig.file %in% c("Oppasum1_neun_pos","Oppasum2_neun_pos", "op_neunpos_B2", "op_neunpos_B5", "OpNeuNs1", "OpNeuNs2"))

Opossum_RGC <- subset(Opossum_RGC, subset = nCount_RNA < 40000)

VlnPlot(Opossum_RGC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Opossum_RGC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)


saveRDS(Opossum_RGC, file = "../Species_Objects/OpossumRGC_v1.rds")
```

### Run clustering pipeline

```{r}
Opossum_RGC <- ClusterSeurat(Opossum_RGC, integrate.by = "orig.file", numPCs = 20)
saveRDS(Opossum_RGC, file = "../Species_Objects/OpossumRGC_int_v1.rds")
```

### Visualize initial clusters
```{r}
# Opossum_RGC <- readRDS("~/Species_Objects/OpossumRGC_integrated_v1.rds")
DimPlot(Opossum_RGC, label = TRUE)
DimPlot(Opossum_RGC, group.by = "orig.file", cells = sample(colnames(Opossum_RGC)))
VlnPlot(Opossum_RGC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Opossum_RGC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

Examine markers for suspicious clusters. There don't appear to be contaminant clusters.
```{r}
Opossum_RGC <- DendroOrder(Opossum_RGC)

RGC_markers= c("RBPMS2","RBPMS", "SLC17A6", "THY1", "NEFL", "NEFM", "SNCG")
BC_markers=c("VSX2", "OTX2", "GRM6", "TRPM1", "VSX1", "CABP5", "GRIK1", "PRKCA")
AC_markers=c("POU6F2", "TFAP2A", "TFAP2B", "TFAP2C", "GAD1", "GAD2", "SLC6A9",  "C1QL1", "C1QL2")
HC_markers=c("ONECUT1", "ONECUT2", "LHX1", "CALB1", "TPM3")
Cone_markers=c("PDE6H", "CRX", "ARR3")
Rod_markers=c("SAG", "PDC", "RHO")
MG_markers=c("SLC1A3","RLBP1", "APOE")
Other_markers=c("S100B", "GFAP", "C1QA", "C1QB", "CLDN5", "IGFBP7")


DotPlot(Opossum_RGC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), assay = "RNA", group.by = "dendro_order") + RotatedAxis()
```

```{r}
# Remove cluster 0,1 29, 32: AC doublet
Opossum_RGC <- DropClusters(Opossum_RGC, idents = c(0,1,39,13,31), refactor = FALSE)

# Clusters 16,19, 5,9,27, 10, 17 all express ONECUT, HC doublet?
Opossum_RGC <- DropClusters(Opossum_RGC, idents = c(12,5,35,37,16,8))
```


### Refine Clusters
Increasing cluster resolution doesn't seem to do much. Stick with 0.8
```{r}
DefaultAssay(Opossum_RGC) <- "integrated"
Opossum_RGC <- FindNeighbors(Opossum_RGC, dims = 1:20)
Opossum_RGC <- RunUMAP(Opossum_RGC, dims = 1:20)

Opossum_RGC <- FindClusters(Opossum_RGC, resolution = 1.4)

DimPlot(Opossum_RGC, label = TRUE, group.by = "integrated_snn_res.0.5")
DimPlot(Opossum_RGC, label = TRUE, group.by = "integrated_snn_res.0.8")
DimPlot(Opossum_RGC, label = TRUE, group.by = "integrated_snn_res.1.1")
DimPlot(Opossum_RGC, label = TRUE, group.by = "integrated_snn_res.1.4")



```


Use resolution of 1.1
```{r}
DefaultAssay(Opossum_RGC) <- "RNA"
Idents(Opossum_RGC) <- "integrated_snn_res.1.1"
Opossum_RGC@meta.data$seurat_clusters <- Idents(Opossum_RGC)
Opossum_RGC <- DendroOrder(Opossum_RGC)
```

Merge any clusters / drop clusters
```{r}
DotPlot(Opossum_RGC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), assay = "RNA", group.by = "dendro_order") + RotatedAxis()
```

```{r}
old_Opossum <- readRDS("../Species_Objects/OpossumRGC_int_v3.rds")
common <- colnames(old_Opossum)[colnames(old_Opossum) %in% colnames(Opossum_RGC)]
MakePrettyConfusionMatrix(table(old_Opossum@meta.data[common, "type"], Opossum_RGC@meta.data[common, "type"]), xlab.use = "New Types", ylab.use = "Old Types")
```



```{r}
# Drop 34, 41 AC
Opossum_RGC <- DropClusters(Opossum_RGC, idents = c(34,31), refactor = FALSE)
# Merge 1 and 20: difference in expression levels
Opossum_RGC <- MergeClusters(Opossum_RGC, idents = c(8,24), refactor = FALSE)
Opossum_RGC <- MergeClusters(Opossum_RGC, idents = c(1, 25), refactor = TRUE)



ident.1 = 1
ident.2 = 25
mark_dif <- FindMarkers(Opossum_RGC, ident.1 = ident.1, ident.2 = ident.2)
mark_dif <- mark_dif[order(mark_dif$avg_log2FC),]

DotPlot(Opossum_RGC, idents = c(ident.1, ident.2), features = c(head(rownames(mark_dif), 10), tail(rownames(mark_dif),10))) + RotatedAxis()

mark.1 <- FindMarkers(Opossum_RGC, ident.1 = ident.1, max.cells.per.ident = 1000)
mark.2 <- FindMarkers(Opossum_RGC, ident.1 = ident.2, max.cells.per.ident = 1000)
DotPlot(Opossum_RGC, features = unique(c(head(rownames(mark.1)), head(rownames(mark.2))))) + RotatedAxis()
```

### Summary Plots
```{r}
Opossum_RGC@meta.data$type <- Opossum_RGC@meta.data$seurat_clusters
DimPlot(Opossum_RGC, label = TRUE, group.by = "type")
saveRDS(Opossum_RGC, "../Species_Objects/OpossumRGC_int_v4.rds")
saveRDS(Opossum_RGC, "../Metadata/RGC/OpossumRGC_int_v4_metadata.rds")
```

```{r}
Opossum_markers <- FindAllMarkers(Opossum_RGC, only.pos = TRUE, max.cells.per.ident = 1000)

Opossum_RGC <- DendroOrder(Opossum_RGC)
Idents(Opossum_RGC) <- "dendro_order"

PlotUniqueMarkers(Opossum_RGC, Opossum_markers)

DotPlot(Opossum_RGC, features = TopMarkers(Opossum_markers, num_markers = 2), group.by = "dendro_order") + RotatedAxis()

```

```{r}
DefaultAssay(Opossum_RGC) <- "RNA"
Idents(Opossum_RGC) <- "type"
saveRDS(Opossum_RGC, "../Species_Objects/OpossumRGC_int_v3.rds")
saveRDS(Opossum_RGC@meta.data, "../Metadata/RGC/OpossumRGC_int_v3.rds")

```


# BC clustering analysis
```{r}
Opossum <- readRDS("../Species_Objects/Opossum_initial.rds")

# Only keep BCs
Idents(Opossum) <- "cell_class"
Opossum_BC <- subset(Opossum, cells = WhichCells(Opossum, idents ="BP"))

# Remove cells with low and high counts or features
VlnPlot(Opossum_BC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Opossum_BC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)
```

```{r}
# Remove cells from NEUNpos samples
Opossum_BC <- subset(Opossum_BC, subset = (orig.file != "Oppasum1_neun_pos") & (orig.file != "Oppasum2_neun_pos") & (orig.file != "op_neunpos_B2") & (orig.file != "op_neunpos_B5"))
Opossum_BC <- subset(Opossum_BC, subset = nCount_RNA < 20000)

saveRDS(Opossum_BC, file = "../Species_Objects/OpossumBC_v1.rds")
```

### Run clustering pipeline

```{r}
# Opossum_BC <- readRDS(file = "../Species_Objects/OpossumBC_v1.rds")
Opossum_BC <- ClusterSeurat(Opossum_BC, integrate.by = "orig.file", numPCs = 20)
saveRDS(Opossum_BC, "../Species_Objects/OpossumBC_int_v1.rds")
```

### Visualize initial clusters
```{r}
DimPlot(Opossum_BC, label = TRUE)
DimPlot(Opossum_BC, group.by = "orig.file", cells = sample(colnames(Opossum_BC)))
VlnPlot(Opossum_BC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Opossum_BC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```


Examine markers for suspicious clusters. 
```{r}
Opossum_BC <- DendroOrder(Opossum_BC)
DotPlot(Opossum_BC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), assay = "RNA", group.by = "dendro_order") + RotatedAxis()
```

```{r}
# 18 expreses AC markers
# 23,20,25,14, 24, 19, 5 express Rho
Opossum_BC <- DropClusters(Opossum_BC, idents = c(18,23,20,25,14, 24, 19, 5))
```

### Refine Clusters

Increase cluster resolution. Stick with 0.8
```{r}
DefaultAssay(Opossum_BC) <- "integrated"

numPCs <- 20
Opossum_BC <- FindNeighbors(Opossum_BC, dims = 1:numPCs)
Opossum_BC <- RunUMAP(Opossum_BC, dims = 1:numPCs)

Opossum_BC <- FindClusters(Opossum_BC, resolution = 1.4)

DimPlot(Opossum_BC, label = TRUE, group.by = "integrated_snn_res.0.5")
DimPlot(Opossum_BC, label = TRUE, group.by = "integrated_snn_res.0.8")
DimPlot(Opossum_BC, label = TRUE, group.by = "integrated_snn_res.1.1")
DimPlot(Opossum_BC, label = TRUE, group.by = "integrated_snn_res.1.4")

```

Merge Clusters
Stick with 0.8, results are similar
```{r}
Idents(Opossum_BC) <- "integrated_snn_res.0.8"
DefaultAssay(Opossum_BC) <- "RNA"
Opossum_BC@meta.data$seurat_clusters <- Idents(Opossum_BC)
Opossum_BC <- MergeClusters(Opossum_BC, idents = c(0,1), refactor = TRUE)
Opossum_BC@meta.data$type <- Opossum_BC@meta.data$seurat_clusters
```

### Summary Plots
```{r}
# 6374 is Vsx2, 20977 is Grik1, 


Opossum_BC <- DendroOrder(Opossum_BC)

# 42052 is Cabp5
pan_BC <- c("VSX2", "OTX2", "CABP5")
# 4147 is Prkca, 12947 is Gramd1b, 
Rod <- c("PRKCA", "GRAMD1B")
# 20977 is Grik1
Off <- c("GRIK1")
# 19439 is Isl1
On <- c("GRM6", "ISL1")

DotPlot(Opossum_BC, features = c(pan_BC, Rod, Off, On), group.by = "dendro_order", assay = "RNA") + RotatedAxis()


Opossum_BC@meta.data$on_off <- factor(Opossum_BC@meta.data$type, levels = rev(c(1,2,4,9,10,11,12,15,16,17,3,5,6,7,8,13,14)))

DotPlot(Opossum_BC, features = c(pan_BC, Rod, Off, On), group.by = "on_off", assay = "RNA") + RotatedAxis()
```


```{r}
Idents(Opossum_BC) <- "type"
DefaultAssay(Opossum_BC) <- "RNA"
DimPlot(Opossum_BC, label = TRUE)
```

```{r}
saveRDS(Opossum_BC, "../Species_Objects/OpossumBC_int_v3.rds")
saveRDS(Opossum_BC@meta.data, "../Metadata/BC/OpossumBC_int_v3_metadata.rds")

```

#PR
```{r}
Idents(Opossum) <- "cell_class"
Opossum_PR <- subset(Opossum, cells = WhichCells(Opossum, idents = c("Rod", "Cone")))

# Remove cells with low and high counts or features
VlnPlot(Opossum_PR, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Opossum_PR, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

Opossum_PR <- ClusterSeurat(Opossum_PR,  numPCs = 30, integrate.by = "orig.file")
saveRDS(Opossum_PR, file = "../Species_Objects/Opossum_PR_int_v1.rds")

DimPlot(Opossum_PR, label = TRUE)
DimPlot(Opossum_PR, group.by = "orig.file", cells = sample(colnames(Opossum_PR)))
VlnPlot(Opossum_PR, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Opossum_PR, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Opossum_PR, features = HC_markers, assay = "RNA") + RotatedAxis()
DotPlot(Opossum_PR, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "OPN1SW", "OPN1LW", "OPN", "OPN5"), assay = "RNA") + RotatedAxis()

#27 Cone SW
#keep 3 and 10 seperate, Cone LW
#rest all rod
Opossum_PR <- MergeClusters(Opossum_PR, idents = c(0,1,2,4,5,6,7,8,9,11,12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,28,29,30,31), refactor = FALSE)
DimPlot(Opossum_PR, label = TRUE)
Opossum_PR@meta.data$type = "ML_cone"
Idents(Opossum_PR) <- "seurat_clusters"
Opossum_PR@meta.data[WhichCells(Opossum_PR, idents = 27), "type"] <- "S_cone"
Opossum_PR@meta.data[WhichCells(Opossum_PR, idents = 0), "type"] <- "Rod"
DimPlot(Opossum_PR, group.by = "type", label = TRUE)
DotPlot(Opossum_PR, group.by = "type", features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "OPN1SW", "OPN1LW", "OPN"), assay = "RNA") + RotatedAxis()
saveRDS(Opossum_PR, file = "../Species_Objects/Opossum_PR_v2.rds")
```

#HC
```{r}
# Only keep HCs
Idents(Opossum) <- "cell_class"
Opossum_HC <- subset(Opossum, cells = WhichCells(Opossum, idents ="HC"))

#Remove cells with low and high counts or features
VlnPlot(Pig_HC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Pig_HC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)


Opossum_HC <- subset(Opossum_HC, (orig.file == "op_neunpos_B2") | (orig.file == "op_neunpos_B5") | (orig.file == "OpNeuNs1") | (orig.file == "OpNeuNs2") | (orig.file == "Oppasum2_neun_pos"))
Opossum_HC <- ClusterSeurat(Opossum_HC,  numPCs = 15, integrate.by = "orig.file")
saveRDS(Opossum_HC, file = "../Species_Objects/Opossum_HC_v1.rds")

DimPlot(Opossum_HC, label = TRUE)
DimPlot(Opossum_HC, group.by = "orig.file", cells = sample(colnames(Opossum_HC)))
VlnPlot(Opossum_HC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Opossum_HC, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Opossum_HC, features = HC_markers, assay = "RNA") + RotatedAxis()
DotPlot(Opossum_HC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "TFAP2B", "TFAP2C", "ISL1", "ONECUT2", "ONECUT3"), assay = "RNA") + RotatedAxis()

Opossum_HC <- DropClusters(Opossum_HC, idents = 3, refactor = FALSE)

Opossum_HC@meta.data$type = "H2"
Idents(Opossum_HC) <- "seurat_clusters"
Opossum_HC@meta.data[WhichCells(Opossum_HC, idents = 0), "type"] <- "H1"
DimPlot(Opossum_HC, group.by = "type", label = TRUE)
DotPlot(Opossum_HC, group.by ="type", features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "TFAP2B", "TFAP2C", "ISL1", "ONECUT2", "ONECUT3"), assay = "RNA") + RotatedAxis()

#keep 0, 1, 2 seperate
#drop 3
```


