---
title: "R Notebook"
output: html_notebook
---

## Preprocessing 

Load libraries
```{r}
library(reshape2)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
source("../utils/utilFxns.R")
source("../utils/plottingFxns.R")
```

Set data directory and load loom files
```{r}
data_dir <- "/Users/joshhahn/Google Drive/My Drive/shekharlab_data/projects/Sheep/"


#Convert(paste0(data_dir, "LambNEUN/velo_outs/full.h5ad"), dest = "h5seurat")
NEUN <- LoadH5Seurat(paste0(data_dir, "LambNEUN/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "lambCHX10/velo_outs/full.h5ad"), dest = "h5seurat")
CHX10 <- LoadH5Seurat(paste0(data_dir, "lambCHX10/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "lambneun (1)/velo_outs/full.h5ad"), dest = "h5seurat")
neun <- LoadH5Seurat(paste0(data_dir, "lambneun (1)/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "lambchx10 (1)/velo_outs/full.h5ad"), dest = "h5seurat")
chx10 <- LoadH5Seurat(paste0(data_dir, "lambchx10 (1)/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "ShChx10/velo_outs/full.h5ad"), dest = "h5seurat")
ShChx10 <- LoadH5Seurat(paste0(data_dir, "ShChx10/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "ShNeuN1/velo_outs/full.h5ad"), dest = "h5seurat")
NeuN1 <- LoadH5Seurat(paste0(data_dir, "ShNeuN1/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "ShNeuN2/velo_outs/full.h5ad"), dest = "h5seurat")
NeuN2 <- LoadH5Seurat(paste0(data_dir, "ShNeuN2/velo_outs/full.h5seurat"))


```

Combine matrices and create new Seurat object
```{r}
Sheep_mat <- cbind(CHX10@assays$matrix@counts, NEUN@assays$matrix@counts, chx10@assays$matrix@counts, neun@assays$matrix@counts, ShChx10@assays$matrix@counts, NeuN1@assays$matrix@counts, NeuN2@assays$matrix@counts)

Sheep <- CreateSeuratObject(Sheep_mat, names.delim = ":")

Sheep@meta.data[colnames(CHX10), 'orig.file'] = "lambCHX10"
Sheep@meta.data[colnames(NEUN), 'orig.file'] = "LambNEUN"
Sheep@meta.data[colnames(chx10), 'orig.file'] = "lambchx10"
Sheep@meta.data[colnames(neun), 'orig.file'] = "lambneun"
Sheep@meta.data[colnames(ShChx10), 'orig.file'] = "ShChx10"
Sheep@meta.data[colnames(NeuN1), 'orig.file'] = "ShNeuN1"
Sheep@meta.data[colnames(NeuN2), 'orig.file'] = "ShNeuN2"


```

```{r}
data_dir <- "/Users/joshhahn/Google Drive/My Drive/New Data/sheep_neunpos_B3/"

Convert(paste0(data_dir, "velo_outs/full.h5ad"), dest = "h5seurat")
neunpos <- LoadH5Seurat(paste0(data_dir, "velo_outs/full.h5seurat"))

old_Sheep <- readRDS("../Species_Initial/Sheep_initial.rds")
Sheep_mat <- cbind(old_Sheep@assays$RNA@counts, neunpos@assays$matrix@counts)

Sheep <- CreateSeuratObject(Sheep_mat)
Sheep@meta.data$orig.file <- "sheep_neunpos_B3"
Sheep@meta.data[rownames(old_Sheep@meta.data), "orig.file"] <- old_Sheep@meta.data$orig.file

```


```{r}
VlnPlot(Sheep, features = "nCount_RNA", pt.size = 0, group.by = "orig.file")
VlnPlot(Sheep, features = "nFeature_RNA", pt.size = 0, group.by = "orig.file")
```



## Cell class annotation

```{r}
Sheep <- ClusterSeurat(Sheep, cluster_resolution = 0.3)
saveRDS(Sheep, file = "../Species_Initial/Sheep_initial.rds")
```


Visualize initial clusters
```{r}
DimPlot(Sheep, label = TRUE)
DimPlot(Sheep, group.by = "orig.file")
VlnPlot(Sheep, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Sheep, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```


Look for major retinal cell classes
```{r}
Sheep <- DendroOrder(Sheep)

RGC_markers= c("RBPMS", "RBPMS2", "SLC17A6", "THY1", "NEFL", "NEFM", "SNCG")
BC_markers=c("VSX2", "OTX2", "GRM6", "TRPM1", "VSX1", "CABP5", "GRIK1", "PRKCA")
AC_markers=c( "POU6F2","TFAP2A", "TFAP2B", "TFAP2C", "GAD1", "GAD2", "SLC6A9", "C1QL1", "C1QL2")
HC_markers=c("ONECUT1", "LHX1", "CALB1", "TPM3")
Cone_markers=c("PDE6H", "CRX", "ARR3")
Rod_markers=c("SAG", "PDC", "RHO")
MG_markers=c("SLC1A3","RLBP1", "APOE")
Other_markers=c("S100B", "GFAP", "C1QA", "C1QB", "CLDN5", "IGFBP7")

DotPlot(Sheep, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), group.by = "dendro_order") + RotatedAxis()
```

Annotate clusters using DE genes and Dot plot of marker genes
```{r}
old_Sheep <- readRDS("../Species_Initial/Sheep_initial.rds")
Sheep@meta.data$cell_class = "N/A"
Sheep@meta.data[rownames(old_Sheep@meta.data), "cell_class"] <- old_Sheep@meta.data$cell_class
table(Sheep@meta.data$dendro_order, Sheep@meta.data$cell_class)
```

Annotate clusters using DE genes and Dot plot of marker genes

```{r}
Sheep@meta.data$cell_class = "Nope"
Sheep@meta.data[WhichCells(Sheep, idents =c(13,15)),]$cell_class = "RGC"
Sheep@meta.data[WhichCells(Sheep, idents = c(12,14,28,41,24,42,34,0,22,9,33)),]$cell_class = "BP"
Sheep@meta.data[WhichCells(Sheep, idents = c(38)),]$cell_class = "HC"
Sheep@meta.data[WhichCells(Sheep, idents = c(32)),]$cell_class = "MG"

Sheep@meta.data[WhichCells(Sheep, idents = c(31,3,11)),]$cell_class = "GlyAC"
Sheep@meta.data[WhichCells(Sheep, idents = c(25,20,35,8,17,26,10,27,30,40,2,5,1,6,7,16,21,39,19,36)),]$cell_class = "GabaAC"

Sheep@meta.data[WhichCells(Sheep, idents = c(18)),]$cell_class = "Rod"
Sheep@meta.data[WhichCells(Sheep, idents = c(37)),]$cell_class = "Cone"

Sheep@meta.data[WhichCells(Sheep, idents = c(23,29,4)),]$cell_class = "Other"
```


Investigate unmarked clusters to see if they have any unique markers
```{r}
Sheep_markers <- FindAllMarkers(Sheep, max.cells.per.ident = 300)
Mouse <- readRDS("../Species_Initial/Mouse_initial.rds")

# 9, 29, 4, 26, 5, 17, 27
ident = 5
mark <- subset(Sheep_markers, cluster == ident)
DotPlot(Mouse, features = head(mark$gene, 20), group.by = "cell_class") + RotatedAxis()

```

```{r}
DimPlot(Sheep, group.by = "cell_class")
counts <- table(Sheep@meta.data$cell_class, Sheep@meta.data$orig.file)
counts
saveRDS(Sheep, "../Species_Initial/Sheep_initial.rds")
saveRDS(Sheep@meta.data, "../Metadata/Sheep_metadata.rds")
```


## RGC clustering analysis
Create separate RGC object
```{r}
Sheep <- readRDS("../Species_Initial/Sheep_initial.rds")

# Only keep RGCs
Idents(Sheep) <- "cell_class"
Sheep_RGC <- subset(Sheep, cells = WhichCells(Sheep, idents ="RGC"))

# Remove cells with low and high counts or features
VlnPlot(Sheep_RGC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Sheep_RGC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

```

```{r}
Sheep_RGC <- subset(Sheep_RGC, subset = (orig.file == "lambneun") | (orig.file == "ShNeuN1") | (orig.file == "ShNeuN2") | (orig.file == "LambNEUN") | (orig.file == "sheep_neunpos_B3"))
Sheep_RGC <- subset(Sheep_RGC, nFeature_RNA > 1000)

Sheep_RGC <- subset(Sheep_RGC, orig.file != "LambNEUN")


saveRDS(Sheep_RGC, file = "../Species_Objects/SheepRGC_v1.rds")
```


Run clustering pipeline

```{r}
Sheep_RGC <- ClusterSeurat(Sheep_RGC, integrate.by = "orig.file")
saveRDS(Sheep_RGC, "../Species_Objects/SheepRGC_int_v1.rds")
```

Visualize initial clusters
```{r}
DimPlot(Sheep_RGC, label = TRUE)
DimPlot(Sheep_RGC, group.by = "orig.file", cells = sample(colnames(Sheep_RGC)))
VlnPlot(Sheep_RGC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Sheep_RGC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

```{r}
DefaultAssay(Sheep_RGC) <- "integrated"

Sheep_RGC <- FindClusters(Sheep_RGC, resolution = 1.4)

DimPlot(Sheep_RGC, label = TRUE, group.by = "integrated_snn_res.0.5")
DimPlot(Sheep_RGC, label = TRUE, group.by = "integrated_snn_res.0.8")
DimPlot(Sheep_RGC, label = TRUE, group.by = "integrated_snn_res.1.1")
DimPlot(Sheep_RGC, label = TRUE, group.by = "integrated_snn_res.1.4")
```

Increasing above 0.8 doesn't impact
```{r}
Sheep_RGC@meta.data$seurat_clusters <- Sheep_RGC@meta.data$integrated_snn_res.0.8
Idents(Sheep_RGC) <- "seurat_clusters"

DefaultAssay(Sheep_RGC) <- "RNA"
Sheep_RGC <- DendroOrder(Sheep_RGC)
```


Examine markers for suspicious clusters, remove contaminant clusters
```{r}
DotPlot(Sheep_RGC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers)) + RotatedAxis()

VlnPlot(Sheep_RGC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Sheep_RGC, "nFeature_RNA", pt.size = 0) + RotatedAxis()

```

Determine if any clusters need to be merged
```{r}
# Drop cluster: low features, doesn't express rbpms
Sheep_RGC <- DropClusters(Sheep_RGC, idents = 1, refactor = TRUE)
DefaultAssay(Sheep_RGC) <- "integrated"
Sheep_RGC <- FindNeighbors(Sheep_RGC, dims = 1:20)
Sheep_RGC <- RunUMAP(Sheep_RGC, dims = 1:20)
DimPlot(Sheep_RGC, label = TRUE)

saveRDS(Sheep_RGC, "../Species_Objects/SheepRGC_int_v2.rds")
saveRDS(Sheep_RGC, "../Metadata/RGC/SheepRGC_int_v2_metadata.rds")
```


## BC clustering analysis
Create separate BC object
```{r}
Sheep <- readRDS("../Species_Initial/Sheep_initial.rds")

# Only keep BCs
Idents(Sheep) <- "cell_class"
Sheep_BC <- subset(Sheep, cells = WhichCells(Sheep, idents ="BP"))

# Remove cells with low and high counts or features
VlnPlot(Sheep_BC, "nCount_RNA", group.by = "orig.file", pt.size = 0.05)
VlnPlot(Sheep_BC, "nFeature_RNA", group.by = "orig.file", pt.size = 0.05)


```

```{r}
Sheep_BC <- subset(Sheep_BC, subset = nFeature_RNA > 100 & nCount_RNA < 15000)

# Only keep cells from chx10-enriched samples
Sheep_BC <- subset(Sheep_BC, subset = (orig.file == "lambchx10") | (orig.file == "lambCHX10") | (orig.file == "ShChx10") )


saveRDS(Sheep_BC, file = "../Species_Objects/SheepBC_v2.rds")
```


Run clustering pipeline

```{r}
Sheep_BC <- ClusterSeurat(Sheep_BC,  numPCs = 20, integrate.by = "orig.file")

saveRDS(Sheep_BC, "../Species_Objects/SheepBC_integrated_v2.rds")
```

Visualize initial clusters
```{r}
Sheep_BC <- readRDS("../Species_Objects/SheepBC_integrated_v2.rds")
DimPlot(Sheep_BC, label = TRUE)
DimPlot(Sheep_BC, group.by = "orig.file", cells = sample(colnames(Sheep_BC)))
VlnPlot(Sheep_BC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Sheep_BC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

Adjust resolution parameter
```{r}
DefaultAssay(Sheep_BC) <- "integrated"

Sheep_BC <- FindClusters(Sheep_BC, resolution = 1.4)

DimPlot(Sheep_BC, label = TRUE, group.by = "integrated_snn_res.0.5")
DimPlot(Sheep_BC, label = TRUE, group.by = "integrated_snn_res.0.8")
DimPlot(Sheep_BC, label = TRUE, group.by = "integrated_snn_res.1.1")
DimPlot(Sheep_BC, label = TRUE, group.by = "integrated_snn_res.1.4")
```


Merge clusters.
```{r}
Sheep_BC@meta.data$seurat_clusters <- Sheep_BC@meta.data$integrated_snn_res.0.8
Idents(Sheep_BC) <- "seurat_clusters"
DefaultAssay(Sheep_BC) <- "RNA"

DotPlot(Sheep_BC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers)) + RotatedAxis()
VlnPlot(Sheep_BC, features = "nCount_RNA", pt.size = 0)
```

```{r}
# 12 expresses PR markers
Sheep_BC <- DropClusters(Sheep_BC, idents = 12, refactor = FALSE)
# Merge rod bipolar clusters
Sheep_BC <- MergeClusters(Sheep_BC, idents = c(0:5), refactor = FALSE)
# 21 expresses all markers of 9, as does 19 with 7
Sheep_BC <- MergeClusters(Sheep_BC, idents = c(9,21), refactor = FALSE)
Sheep_BC <- MergeClusters(Sheep_BC, idents = c(7,19), refactor = TRUE)




# Plenty of markers distinguish cluster 6 and 10
# 5 and 8 have enough markers to distinguish

ident.1 = 7
ident.2 = 19
mark_dif <- FindMarkers(Sheep_BC, ident.1 = ident.1, ident.2 = ident.2)
mark_dif <- mark_dif[order(mark_dif$avg_log2FC),]

DotPlot(Sheep_BC, idents = c(ident.1, ident.2), features = c(head(rownames(mark_dif)), tail(rownames(mark_dif)))) + RotatedAxis()

mark.1 <- FindMarkers(Sheep_BC, ident.1 = ident.1, max.cells.per.ident = 1000)
mark.2 <- FindMarkers(Sheep_BC, ident.1 = ident.2, max.cells.per.ident = 1000)
DotPlot(Sheep_BC, features = unique(c(head(rownames(mark.1)), head(rownames(mark.2))))) + RotatedAxis()

```

```{r}
Sheep_BC@meta.data$type <- Sheep_BC@meta.data$seurat_clusters

DimPlot(Sheep_BC, label = TRUE, group.by = "type")
```



```{r}
DefaultAssay(Sheep_BC) <- "RNA"
Idents(Sheep_BC) <- "type"
Sheep_BC <- DendroOrder(Sheep_BC)

pan_BC <- c("VSX2", "CABP5")
Rod <- c("PRKCA", "GRAMD1B")
Off <- c("GRIK1")
On <- c("GRM6", "ISL1")

DotPlot(Sheep_BC, features = c(pan_BC, Rod, Off, On), group.by = "dendro_order", assay = "RNA") + RotatedAxis()

Sheep_BC@meta.data$on_off <- factor(Sheep_BC@meta.data$seurat_clusters, levels = rev(c(1,4,5,6,7,9,10,2,3,8,11,12)))

DotPlot(Sheep_BC, features = c(pan_BC, Rod, Off, On), group.by = "on_off", assay = "RNA") + RotatedAxis()

```

Examine markers for suspicious clusters, remove contaminant clusters
```{r}
Sheep_BC_markers <- FindAllMarkers(Sheep_BC, max.cells.per.ident = 1000)
```

```{r}
Idents(Sheep_BC) <- "type"
saveRDS(Sheep_BC, "../Species_Objects/SheepBC_integrated_v3.rds")
```

```{r}
Sheep <- readRDS("~Downloads/Species_Initial/Sheep_initial.rds")

# Only keep HCs
Idents(Sheep) <- "cell_class"
Sheep_HC <- subset(Sheep, cells = WhichCells(Sheep, idents ="HC"))

# Remove cells with low and high counts or features
VlnPlot(Sheep_HC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Sheep_HC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)


Sheep_HC <- subset(Sheep_HC, (orig.file == "lambneun") |  (orig.file == "LambNEUN"))
Sheep_HC <- subset(Sheep_HC, subset = nCount_RNA < 20000)
Sheep_HC <- ClusterSeurat(Sheep_HC, integrate.by = "orig.file")
saveRDS(Sheep_HC, file = "../Species_Objects/Sheep_HC_int_v1.rds")

DimPlot(Sheep_HC, label = TRUE)
DimPlot(Sheep_HC, group.by = "orig.file", cells = sample(colnames(Sheep_HC)))
VlnPlot(Sheep_HC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Sheep_HC, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Sheep_HC, features = HC_markers, assay = "RNA") + RotatedAxis()
DotPlot(Sheep_HC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "ISL1", "ONECUT2", "ONECUT3"), assay = "RNA") + RotatedAxis()

Sheep_HC <- FindNeighbors(Sheep_HC, dims = 1:15)
Sheep_HC <- RunUMAP(Sheep_HC, dims = 1:15)

Sheep_HC@meta.data$type = "HC1"
Idents(Sheep_HC) <- "seurat_clusters"
Sheep_HC@meta.data[WhichCells(Sheep_HC, idents = 1), "type"] <- "H2"
Sheep_HC@meta.data[WhichCells(Sheep_HC, idents = 2), "type"] <- "H1"
DimPlot(Sheep_HC, group.by = "type", label = TRUE)
DotPlot(Sheep_HC, group.by ="type", features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "ISL1", "ONECUT2", "ONECUT3"), assay = "RNA") + RotatedAxis()
saveRDS(Sheep_HC, file = "../Species_Objects/Sheep_HC_int_v2.rds")
```

#PR
```{r}
Idents(Sheep) <- "cell_class"
Sheep_PR <- subset(Sheep, cells = WhichCells(Sheep, idents = c("Rod", "Cone")))

# Remove cells with low and high counts or features
VlnPlot(Sheep_PR, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Sheep_PR, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

Sheep_PR <- ClusterSeurat(Sheep_PR,  numPCs = 20, integrate.by = "orig.file")
saveRDS(Sheep_PR, file = "../Species_Objects/Sheep_PR_int_v1.rds")

DimPlot(Sheep_PR, label = TRUE)
DimPlot(Sheep_PR, group.by = "orig.file", cells = sample(colnames(Sheep_PR)))
VlnPlot(Sheep_PR, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Sheep_PR, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Sheep_PR, features = Rod_markers, assay = "RNA") + RotatedAxis()
DotPlot(Sheep_PR, features = Cone_markers, assay = "RNA") + RotatedAxis()
DotPlot(Sheep_PR, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "OPN1SW", "OPN1LW", "OPN"), assay = "RNA") + RotatedAxis()

#0,1 merge (Rods)
#3, 4 drop
#2 cone (just label as cone)` 
Sheep_PR <- MergeClusters(Sheep_PR, idents = c(0,1), refactor = FALSE)
Sheep_PR <- DropClusters(Sheep_PR, idents = c(3,4), refactor = FALSE)

Sheep_PR@meta.data$type = "ML_cone"
Idents(Sheep_PR) <- "seurat_clusters"
Sheep_PR@meta.data[WhichCells(Sheep_PR, idents = 2), "type"] <- "Cone"
Sheep_PR@meta.data[WhichCells(Sheep_PR, idents = 0), "type"] <- "Rod"
DimPlot(Sheep_PR, group.by = "type", label = TRUE)
DotPlot(Sheep_PR, group.by = "type", features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "OPN1SW", "OPN1LW", "OPN"), assay = "RNA") + RotatedAxis()
saveRDS(Sheep_PR, file = "../Species_Objects/Sheep_PR_int_v2.rds")
```

