---
title: "R Notebook"
output: html_notebook
---

## Load libraries
```{r}
library(tidyverse)
library(ggtree)
library(Seurat)
library(ggplot2)
source("../utils/utilFxns.R")
source("../utils/plottingFxns.R")
```

## Prepare Velocyto Data
Read in velocyto matrices
```{r}
data_dir <- "/Users/joshhahn/Google Drive File Stream/My Drive/shekharlab_data/projects/Squirrel/"

Convert(paste0(data_dir, "SquirrelChx10pos/velo_outs/full.h5ad"), dest = "h5seurat")
Chx10 <- LoadH5Seurat(paste0(data_dir, "SquirrelChx10pos/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "SquirrelNeuNpos/velo_outs/full.h5ad"), dest = "h5seurat")
Neun <- LoadH5Seurat(paste0(data_dir, "SquirrelNeuNpos/velo_outs/full.h5seurat"))

```


Combine matrices and create new Seurat object
```{r}
Squirrel_mat <- cbind(Chx10@assays$matrix@counts, Neun@assays$matrix@counts)
Squirrel <- CreateSeuratObject(Squirrel_mat, names.delim = ":")

Squirrel@meta.data[colnames(Chx10), 'orig.file'] = "SquirrelChx10pos"
Squirrel@meta.data[colnames(Neun), 'orig.file'] = "SquirrelNeuNpos"

```

Initial quality control metrics
```{r}
VlnPlot(Squirrel, features = "nCount_RNA", pt.size = 0)
VlnPlot(Squirrel, features = "nFeature_RNA", pt.size = 0)
```


## Cell Class Annotation
Perform initial clustering
```{r}
Squirrel <- ClusterSeurat(Squirrel)
saveRDS(Squirrel, file = "../Species_Objects/Squirrel_initial.rds")
```

Visualize initial clusters
```{r}
Squirrel <- readRDS("../Species_Objects/Squirrel_initial.rds")
DimPlot(Squirrel, label = TRUE)
DimPlot(Squirrel, group.by = "orig.file", cells = sample(colnames(Squirrel)))
VlnPlot(Squirrel, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Squirrel, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

Calculate DE genes for cluster annotation
```{r}
Squirrel_genes <- FindAllMarkers(Squirrel)
saveRDS(Squirrel_genes, file = "../Marker_Genes/Squirrel_genes_initial.rds")
```


Look for major retinal cell classes
```{r}
Idents(Squirrel) <- "seurat_clusters"
Squirrel <- DendroOrder(Squirrel)
RGC_markers= c("RBPMS", "POU4F1", "THY1")
BC_markers=c("VSX1", "VSX2", "OTX2", "GRIK1", "GRIK2","PRKCA")
AC_markers=c("TFAP2A","TFAP2B", "GAD1", "GAD2", "SLC6A9")
MG_markers=c("SLC1A3","RLBP1", "GLUL", "CLU")
HC_markers=c("ONECUT1", "ONECUT2","ONECUT3")
Cone_markers=c("OPN1SW", "RS1", "GNGT2")
Rod_markers=c("RHO", "PDE6B", "PDC")
OL_markers = c("OLIG2")

DotPlot(Squirrel, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, OL_markers), group.by = "dendro_order") + RotatedAxis()
```



Annotate clusters based on expression of cell class markers
```{r}
Squirrel@meta.data$cell_class = "Other"
Squirrel@meta.data[WhichCells(Squirrel, idents =c(31,17,37,43,44,18,47,11,28,25,20,33,13,45,39,24,7,0,4,3,42,40,32)),]$cell_class = "RGC"
Squirrel@meta.data[WhichCells(Squirrel, idents = c(22,14,10,30,26,52,35,12,15,8,6,38)),]$cell_class = "BP"
Squirrel@meta.data[WhichCells(Squirrel, idents = c(51,34,41,19,9,29,1,54,21,27,23)),]$cell_class = "GabaAC"
Squirrel@meta.data[WhichCells(Squirrel, idents = c(16, 36)),]$cell_class = "GlyAC"
Squirrel@meta.data[WhichCells(Squirrel, idents = c(46)),]$cell_class = "HC"

# Photoreceptors do separate into rods and cones, do a round of clustering to separate rods and cones
Squirrel_PR <- subset(Squirrel, cells = WhichCells(Squirrel, idents =  c(50,2)))
Squirrel_PR <- ClusterSeurat(Squirrel_PR)

Rod_markers <- c("RHO" ,"PDC", "NRL", "SAG", "GNAT1", "GNGT1")
Cone_markers <- c("ARR3", "PDE6H", "GNAT2","GNGT2","OPN1SW","OPN1MW")
DotPlot(Squirrel_PR, features = c(Rod_markers, Cone_markers)) + RotatedAxis()

rod_cells <- WhichCells(Squirrel_PR, idents = 5)
cone_cells <- WhichCells(Squirrel_PR, idents = 0:4)

Squirrel@meta.data[rod_cells,]$cell_class = "Rod"
Squirrel@meta.data[cone_cells,]$cell_class = "Cone"

Squirrel@meta.data[WhichCells(Squirrel, idents = c(48,53,5,55)),]$cell_class = "MG"
Squirrel@meta.data[WhichCells(Squirrel, idents = c(49)),]$cell_class = "MicroG"


saveRDS(Squirrel, "../Species_Objects/Squirrel_initial.rds")
```

```{r}
# Call BC due to GRIK2
markers12 <- FindMarkers(Squirrel, ident.1 = 12)

# Debris: low counts, no DE genes
markers49 <- FindMarkers(Squirrel, ident.1 = 49)
```


Plot cell type by file
```{r}
Squirrel <- readRDS("../Species_Initial/Squirrel_initial.rds")

pdf("../Figures/SquirrelUMAP.pdf", w=5, h=4, useDingbats = FALSE)
DimPlot(Squirrel, group.by = "cell_class", cols = c("red","orange","green","blue","black","purple", "plum")) + ggtitle("Squirrel")
dev.off()

counts <- table(Squirrel@meta.data$cell_class, Squirrel@meta.data$orig.file)
counts
counts <- t(t(counts) / colSums(counts))
barplot(counts, legend = rownames(counts), col= rev(c("red","orange","yellow","white","green","blue", "black")))
```

Import labels from study
```{r}
Squirrel <- readRDS("../Species_Objects/Squirrel_initial.rds")
Chick_meta <- read.csv( "../Species_Reference/Squirrel/Chick_retina_atlas_meta.csv")
meta <- data.frame(cluster = Chick_meta$Cluster, row.names = Chick_meta$NAME)

barcode <- function(string){
    code <- strsplit(string, c(":"))
    barcode <- strsplit(code[[1]][2], "x")
    return (paste0(barcode[[1]][1], "-1"))
}

new_names <- unlist(lapply(colnames(Squirrel), FUN = barcode))
Squirrel@meta.data$barcode <- paste0(Squirrel@meta.data$orig.file, "_", new_names)

Squirrel@meta.data$annotated <- meta[Squirrel@meta.data$barcode,]

saveRDS(Squirrel, "../Species_Objects/Squirrel_initial.rds")
```

## Squirrel RGC
### Subset and filter low quality cells
```{r}
Squirrel <- readRDS("../Species_Objects/Squirrel_initial.rds")
Idents(Squirrel) <- "cell_class"
Squirrel_RGC <- subset(Squirrel, idents = "RGC")

VlnPlot(Squirrel_RGC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Squirrel_RGC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

Squirrel_RGC <- subset(Squirrel_RGC, subset = nFeature_RNA > 1000 & nCount_RNA < 12000)

saveRDS(Squirrel_RGC, "../Species_Objects/Squirrel_RGC_v1.rds")
```

### Initial clustering
```{r}
Squirrel_RGC <- NormalizeData(Squirrel_RGC)
Squirrel_RGC <- FindVariableFeatures(Squirrel_RGC)
Squirrel_RGC <- ScaleData(Squirrel_RGC)
Squirrel_RGC <- RunPCA(Squirrel_RGC)

ElbowPlot(Squirrel_RGC, ndims = 50)

# 20 PCs seems sufficient, start with resolution parameter of 1
numPCs <- 20
cluster_resolution <- 1.5

Squirrel_RGC <- FindNeighbors(Squirrel_RGC, dims = 1:numPCs)
Squirrel_RGC <- FindClusters(Squirrel_RGC, resolution = cluster_resolution)
Squirrel_RGC <- RunTSNE(Squirrel_RGC, dims = 1:numPCs)
Squirrel_RGC <- RunUMAP(Squirrel_RGC, dims = 1:numPCs)

saveRDS(Squirrel_RGC, "../Species_Objects/Squirrel_RGC_v1.rds")
```

### Visualize initial clusters
```{r}
# Squirrel_RGC <- readRDS("~/Species_Objects/Squirrel_RGC_v1.rds")

# Increasing resolution parameter from 1 to 1.5 minimally affects cluster assignment, start with 1.5 and merge 
DimPlot(Squirrel_RGC, label = TRUE, group.by = "RNA_snn_res.1")
DimPlot(Squirrel_RGC, label = TRUE, group.by = "RNA_snn_res.1.5")
Idents(Squirrel_RGC) <- "RNA_snn_res.1.5"

# Due to limited number of cells from the Chx10 experiment, integration will not improve clustering
DimPlot(Squirrel_RGC, group.by = "orig.file", cells = sample(colnames(Squirrel_RGC)))

# Vln plots of counts and features
VlnPlot(Squirrel_RGC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Squirrel_RGC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

### Calculate DE markers for contaminant removal
```{r}
SqrRGC_markers <- FindAllMarkers(Squirrel_RGC)
saveRDS(SqrRGC_markers, "../Species_Markers/SquirrelRGC_markers_v1.rds")
```

### Remove clusters
```{r}
RGC_markers= c("RBPMS", "POU4F1", "THY1")
BC_markers=c("VSX1", "VSX2", "OTX2", "GRIK1", "GRIK2","PRKCA")
AC_markers=c("TFAP2A","TFAP2B", "GAD1", "GAD2")
MG_markers=c("SLC1A3","RLBP1", "GLUL", "CLU")
HC_markers=c("ONECUT1", "ONECUT2","ONECUT3")
Cone_markers=c("RS1", "GNGT2")
Rod_markers=c("RHO", "PDE6B", "PDC")
OL_markers = c("OLIG2")
DotPlot(Squirrel_RGC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, OL_markers)) + RotatedAxis()


# Cluster 31 - muller glia, expresses Glul, Clu
head(subset(SqrRGC_markers, cluster == 31), 20)
# Cluster 19 - rod, expresses Arr3, PDE6C
head(subset(SqrRGC_markers, cluster == 19), 20)

Squirrel_RGC@meta.data$seurat_clusters <- Squirrel_RGC@meta.data$RNA_snn_res.1.5

Squirrel_RGC <- DropClusters(Squirrel_RGC, idents = c(19,31), refactor = FALSE)
```

### Merge clusters
```{r}
Squirrel_RGC <- DendroOrder(Squirrel_RGC)
Idents(Squirrel_RGC) <- "dendro_order"

plot_markers <- NULL
num_markers = 2
for(i in levels(Idents(Squirrel_RGC))){
    top_marks <- head(subset(SqrRGC_markers, cluster == i),num_markers)$gene
    plot_markers <-  unique(c(plot_markers, top_marks))
}

DotPlot(Squirrel_RGC, features = plot_markers) + RotatedAxis()

# Clusters 30, 1: no significant differences, merge
mark_30_1 <- FindMarkers(Squirrel_RGC, ident.1 = 1, ident.2 = 30)
DotPlot(Squirrel_RGC, features = rownames(head(mark_30_1, 10))) + RotatedAxis()

Squirrel_RGC <- MergeClusters(Squirrel_RGC, idents = c(1,30), refactor = FALSE)


# Clusters 33, 3: merge, low significant differences
mark_33_3 <- FindMarkers(Squirrel_RGC, ident.1 = 3, ident.2 = 33)
DotPlot(Squirrel_RGC, features = rownames(head(mark_33_3, 10))) + RotatedAxis()

Squirrel_RGC <- MergeClusters(Squirrel_RGC, idents = c(3,33), refactor = FALSE)


# Clusters 5, 12: keep separate
mark_5_12 <- FindMarkers(Squirrel_RGC, ident.1 = 5, ident.2 = 12)
DotPlot(Squirrel_RGC, features = rownames(head(mark_5_12, 10))) + RotatedAxis()

# CLusters 2, 32: cluster 2 lacks distinguishing markers, merge
mark_2_32 <- FindMarkers(Squirrel_RGC, ident.1 = 2, ident.2 = 32)
DotPlot(Squirrel_RGC, features = rownames(head(mark_2_32, 10))) + RotatedAxis()

Squirrel_RGC <- MergeClusters(Squirrel_RGC, idents = c(2,32), refactor = TRUE)

saveRDS(Squirrel_RGC, "../Species_Objects/Squirrel_RGC_v2.rds")
```

```{r}
DotPlot(Squirrel_RGC, features = plot_markers) + RotatedAxis()
```

```{r}
Squirrel_RGC <- readRDS("../Species_Objects/Squirrel_RGC_v2.rds")

pdf("../Figures/Species_UMAPs/SquirrelRGC.pdf", w=4, h=4, useDingbats = FALSE)
DimPlot(Squirrel_RGC,  label = TRUE) + NoLegend()
dev.off()
```

### Alt clustering

```{r}
ref_RGC <- readRDS("../Species_Reference/Squirrel_RGC_cells_count_matrix_and_cluster_assignment.rds")

barcode <- function(string){
    code <- strsplit(string, c(":"))
    barcode <- strsplit(code[[1]][2], "x")
    return (paste0(barcode[[1]][1], "-1"))
}
new_names <- unlist(lapply(colnames(Squirrel_RGC), FUN = barcode))
Squirrel_RGC@meta.data$barcode <- paste0(Squirrel_RGC@meta.data$orig.file, "_", new_names)

Squirrel_RGC@meta.data$annotated <- ref_RGC$clusterid[Squirrel_RGC@meta.data$barcode]

MakePrettyConfusionMatrix(table(Squirrel_RGC@meta.data$annotated, Squirrel_RGC$RNA_snn_res.1.5), xlab.use = "new", ylab.use = "annotated") 

MakePrettyConfusionMatrix(table(Squirrel_RGC@meta.data$seurat_clusters, Squirrel_RGC@meta.data$annotated), xlab.use = "annotated", ylab.use = "new") 

MakePrettyConfusionMatrix(table(Squirrel_RGC@meta.data$annotated, Squirrel_RGC$seurat_clusters), xlab.use = "new", ylab.use = "annotated") 
```

## Squirrel BC
### Subset and filter low quality cells
```{r}
Squirrel <- readRDS("../Species_Initial/Squirrel_initial.rds")
Idents(Squirrel) <- "cell_class"
Squirrel_BC <- subset(Squirrel, idents = "BP")

VlnPlot(Squirrel_BC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Squirrel_BC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)
```

```{r}
Squirrel_BC <- subset(Squirrel_BC, subset = nFeature_RNA > 800 & nCount_RNA < 6000)
Squirrel_BC <- subset(Squirrel_BC, orig.file == "SquirrelChx10pos")
```

### Initial clustering
```{r}
Squirrel_BC <- NormalizeData(Squirrel_BC)
Squirrel_BC <- FindVariableFeatures(Squirrel_BC)
Squirrel_BC <- ScaleData(Squirrel_BC)
Squirrel_BC <- RunPCA(Squirrel_BC)

ElbowPlot(Squirrel_BC, ndims = 50)

# 20 PCs seems sufficient, start with resolution parameter of 1
numPCs <- 20
cluster_resolution <- .8

Squirrel_BC <- FindNeighbors(Squirrel_BC, dims = 1:numPCs)
Squirrel_BC <- FindClusters(Squirrel_BC, resolution = cluster_resolution)
Squirrel_BC <- RunTSNE(Squirrel_BC, dims = 1:numPCs)
Squirrel_BC <- RunUMAP(Squirrel_BC, dims = 1:numPCs)

saveRDS(Squirrel_BC, "../Species_Objects/Squirrel_BC_v4.rds")
```

### Visualize initial clusters
```{r}
Squirrel_BC <- readRDS("../Species_Objects/Squirrel_BC_v1.rds")

# Increasing resolution parameter from 1 to 1.5 minimally affects cluster assignment, start with 1.5 and merge 
DimPlot(Squirrel_BC, label = TRUE, group.by = "RNA_snn_res.1")
DimPlot(Squirrel_BC, label = TRUE, group.by = "RNA_snn_res.1.5")
Idents(Squirrel_BC) <- "RNA_snn_res.1"


# Vln plots of counts and features
VlnPlot(Squirrel_BC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Squirrel_BC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```



### Remove clusters
```{r}
RGC_markers= c("RBPMS", "POU4F1", "THY1")
BC_markers=c("VSX1", "VSX2", "OTX2", "GRIK1", "GRIK2","PRKCA", "GRM6","TRPM1","CABP5")
AC_markers=c("TFAP2A","TFAP2B", "GAD1", "GAD2")
MG_markers=c("SLC1A3","RLBP1", "GLUL", "CLU")
HC_markers=c("ONECUT1", "ONECUT2","ONECUT3")
Cone_markers=c("RS1", "GNGT2")
Rod_markers=c("RHO", "PDE6B", "PDC")
OL_markers = c("OLIG2")
DotPlot(Squirrel_BC, features = c(BC_markers, RGC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, OL_markers)) + RotatedAxis()
```

```{r}
# 6, 10, 14 do not express BP markers
# 20 is an AC doublet
Squirrel_BC@meta.data$seurat_clusters <- Squirrel_BC@meta.data$RNA_snn_res.1
Squirrel_BC <- DropClusters(Squirrel_BC,idents = c(6,10,14,20))
```

```{r}
# Find neighbors again after removing cells
numPCs <- 20
cluster_resolution <- .8

Squirrel_BC <- FindNeighbors(Squirrel_BC, dims = 1:numPCs)
Squirrel_BC <- FindClusters(Squirrel_BC, resolution = cluster_resolution)
Squirrel_BC <- RunUMAP(Squirrel_BC, dims = 1:numPCs)

DimPlot(Squirrel_BC, label = TRUE)
```

```{r}
saveRDS(Squirrel_BC, "../Species_Objects/Squirrel_BC_v4.rds")
```

### Try NOGs to find any clusters to remove
#### Sample count matrices
```{r}
Mouse_BC <- readRDS("../Species_Objects/MouseBC_int_ann_v3.rds")
Squirrel_BC <- readRDS("../Species_Objects/Squirrel_BC_v4.rds")

Mouse_meta <- Mouse_BC@meta.data
rownames(Mouse_meta) <- paste0("Mouse;", rownames(Mouse_meta))
Mouse_counts <- readRDS("../Species_OrthologMatrix/Mouse_orthomat_v3.rds")

Squirrel_meta <- Squirrel_BC@meta.data
rownames(Squirrel_meta) <- paste0("Squirrel;", rownames(Squirrel_meta))
Squirrel_counts <- readRDS("../Species_OrthologMatrix/Squirrel_orthomat_v3.rds")

joint_counts <- cbind(Mouse_counts[,rownames(Mouse_meta)], Squirrel_counts[,rownames(Squirrel_meta)])
Joint <- CreateSeuratObject(joint_counts, names.delim = ";")
Joint@meta.data$type <- "N/A"
Joint@meta.data[rownames(Mouse_meta), "type"] <- as.character(Mouse_meta$annotated)
Joint@meta.data[rownames(Squirrel_meta), "type"] <- as.character(Squirrel_meta$RNA_snn_res.0.8)
```

#### Clustering
Before integration
```{r}
Joint <- ClusterSeurat(Joint, integrate.by = "orig.ident")
```

```{r}
Joint@meta.data$species_type <- paste0(Joint@meta.data$orig.ident, Joint@meta.data$type)
DimPlot(Joint, group.by = "species_type", label = TRUE)
DimPlot(Joint, group.by = "orig.ident")
```

### Refine clusters
```{r}
Squirrel_BC <- readRDS("../Species_Objects/Squirrel_BC_v4.rds")
Squirrel_BC@meta.data$seurat_clusters <- Squirrel_BC@meta.data$RNA_snn_res.0.8
Idents(Squirrel_BC) <- "seurat_clusters"

# Remove cluster 16 
Squirrel_BC <- DropClusters(Squirrel_BC, idents = 16, refactor = FALSE)

# 9 and 10 - merge: top markers are very similar, differences appear to be in expression levels
Squirrel_BC <- MergeClusters(Squirrel_BC, idents = c(9,10), refactor = FALSE)

# 4 and 13 - merge for similar reasons
Squirrel_BC <- MergeClusters(Squirrel_BC, idents = c(4,13), refactor = TRUE)


ident.1 = 4
ident.2 = 13
mark_dif <- FindMarkers(Squirrel_BC, ident.1 = ident.1, ident.2 = ident.2)
mark_dif$pct.dif <- mark_dif$pct.1 - mark_dif$pct.2
mark_dif <- mark_dif[order(mark_dif$avg_log2FC),]

DotPlot(Squirrel_BC, idents = c(ident.1, ident.2), features = c(head(rownames(mark_dif)), tail(rownames(mark_dif)))) + RotatedAxis()

mark.1 <- FindMarkers(Squirrel_BC, ident.1 = ident.1, max.cells.per.ident = 1000)
mark.2 <- FindMarkers(Squirrel_BC, ident.1 = ident.2, max.cells.per.ident = 1000)
DotPlot(Squirrel_BC, features = unique(c(head(rownames(mark.1)), head(rownames(mark.2))))) + RotatedAxis()
```


### Replot
```{r}
Squirrel_BC@meta.data$type <- Squirrel_BC@meta.data$seurat_clusters

Idents(Squirrel_BC) <- "type"
DefaultAssay(Squirrel_BC) <- "RNA"

DimPlot(Squirrel_BC, label = TRUE)
saveRDS(Squirrel_BC, "../Species_Objects/Squirrel_BC_v5.rds")
```

#PR
```{r}
Idents(Squirrel) <- "cell_class"
Squirrel_PR <- subset(Squirrel, cells = WhichCells(Squirrel, idents = c("Rod", "Cone")))

# Remove cells with low and high counts or features
VlnPlot(Squirrel_PR, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Squirrel_PR, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

Squirrel_PR <- ClusterSeurat(Squirrel_PR,  numPCs = 30, integrate.by = "orig.file")

DimPlot(Squirrel_PR, label = TRUE)
DimPlot(Squirrel_PR, group.by = "orig.file", cells = sample(colnames(Squirrel_PR)))
VlnPlot(Squirrel_PR, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Squirrel_PR, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Squirrel_PR, features = Rod_markers, assay = "RNA") + RotatedAxis()
DotPlot(Squirrel_PR, features = Cone_markers, assay = "RNA") + RotatedAxis()
DotPlot(Squirrel_PR, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "OPN1LW", "OPN", "ENSSTOG00000024701"), assay = "RNA") + RotatedAxis()

#5 (Rods)
#4 S cone
#0,1,2 MW cone

Squirrel_PR <- MergeClusters(Squirrel_PR, idents = c(0,1,2), refactor = FALSE)
Squirrel_PR <- DropClusters(Squirrel_PR, idents = c(3), refactor = FALSE)

Squirrel_PR@meta.data$type = "MW_cone"
Idents(Squirrel_PR) <- "seurat_clusters"
Squirrel_PR@meta.data[WhichCells(Squirrel_PR, idents = 4), "type"] <- "S_cone"
Squirrel_PR@meta.data[WhichCells(Squirrel_PR, idents = 5), "type"] <- "Rod"
DimPlot(Squirrel_PR, group.by = "type", label = TRUE)
DotPlot(Squirrel_PR, group.by = "type", features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "OPN1LW", "OPN", "ENSSTOG00000024701"), assay = "RNA") + RotatedAxis()
saveRDS(Squirrel_PR, file = "../Species_Objects/Squirrel_PR_int_v2.rds")
```

#HC
```{r}
Idents(Squirrel) <- "cell_class"
Squirrel_HC <- subset(Squirrel, cells = WhichCells(Squirrel, idents ="HC"))

# Remove cells with low and high counts or features
VlnPlot(Squirrel_HC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Squirrel_HC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

Squirrel_HC <- subset(Squirrel_HC, (orig.file == "SquirrelChx10pos"))
Squirrel_HC <- ClusterSeurat(Squirrel_HC,  numPCs = 20)

DimPlot(Squirrel_HC, label = TRUE)
DimPlot(Squirrel_HC, group.by = "orig.file", cells = sample(colnames(Squirrel_HC)))
VlnPlot(Squirrel_HC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Squirrel_HC, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Squirrel_HC, features = HC_markers, assay = "RNA") + RotatedAxis()
DotPlot(Squirrel_HC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "TFAP2C", "ISL1"), assay = "RNA") + RotatedAxis()

#drop 2
#merge 0, 1
Squirrel_HC <- DropClusters(Squirrel_HC, idents = c(2), refactor = FALSE)
Squirrel_HC <- MergeClusters(Squirrel_HC, idents = c(0,1), refactor = FALSE)
DimPlot(Squirrel_HC, label = TRUE)
Squirrel_HC <- FindNeighbors(Squirrel_HC, dims = 1:15)
Squirrel_HC <- RunUMAP(Squirrel_HC, dims = 1:15)

Squirrel_HC@meta.data$type = "HC1"
Idents(Squirrel_HC) <- "seurat_clusters"
Squirrel_HC@meta.data[WhichCells(Squirrel_HC, idents = 0), "type"] <- "H1"
DimPlot(Squirrel_HC, group.by = "type", label = TRUE)
DotPlot(Squirrel_HC, group.by = "type", features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "TFAP2C", "ISL1"), assay = "RNA") + RotatedAxis()
saveRDS(Squirrel_HC, file = "Squirrel_HC_int_v2")
```

