---
title: "R Notebook"
output: html_notebook
---

## Preprocessing 

Load libraries
```{r}
library(reshape2)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
source("../utils/utilFxns.R")
source("../utils/plottingFxns.R")
```

Set data directory and load loom files
```{r}
data_dir <- "/Users/joshhahn/Google Drive/My Drive/shekharlab_data/projects/Cow/"


#Convert(paste0(data_dir, "CowNEUN (1)/velo_outs/full.h5ad"), dest = "h5seurat")
NEUN <- LoadH5Seurat(paste0(data_dir, "CowNEUN (1)/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "CowCHx10/velo_outs/full.h5ad"), dest = "h5seurat")
CHX10 <- LoadH5Seurat(paste0(data_dir, "CowCHx10/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "cowchx10 (1)/velo_outs/full.h5ad"), dest = "h5seurat")
chx10 <- LoadH5Seurat(paste0(data_dir, "cowchx10 (1)/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "cowneun/velo_outs/full.h5ad"), dest = "h5seurat")
neun <- LoadH5Seurat(paste0(data_dir, "cowneun/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "CowNeuN1/velo_outs/full.h5ad"), dest = "h5seurat")
NeuN1 <- LoadH5Seurat(paste0(data_dir, "CowNeuN1/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "CowNeuN2/velo_outs/full.h5ad"), dest = "h5seurat")
NeuN2 <- LoadH5Seurat(paste0(data_dir, "CowNeuN2/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "CowCHx10_r2/velo_outs/full.h5ad"), dest = "h5seurat")
CHX10_r2 <- LoadH5Seurat(paste0(data_dir, "CowCHx10_r2/velo_outs/full.h5seurat"))




```

Combine matrices and create new Seurat object
```{r}
Cow_mat <- cbind(CHX10@assays$matrix@counts, NEUN@assays$matrix@counts, chx10@assays$matrix@counts, neun@assays$matrix@counts, NeuN1@assays$matrix@counts, NeuN2@assays$matrix@counts, CHX10_r2@assays$matrix@counts)

Cow <- CreateSeuratObject(Cow_mat, names.delim = ":")

Cow@meta.data[colnames(CHX10), 'orig.file'] = "CowCHx10"
Cow@meta.data[colnames(NEUN), 'orig.file'] = "CowNEUN"
Cow@meta.data[colnames(chx10), 'orig.file'] = "cowchx10"
Cow@meta.data[colnames(neun), 'orig.file'] = "cowneun"
Cow@meta.data[colnames(NeuN1), 'orig.file'] = "CowNeuN1"
Cow@meta.data[colnames(NeuN2), 'orig.file'] = "CowNeuN2"
Cow@meta.data[colnames(CHX10_r2), 'orig.file'] = "CowCHx10_r2"

```

```{r}
data_dir <- "/Users/joshhahn/Google Drive/My Drive/New Data/cow_neunpos_B1/"

Convert(paste0(data_dir, "velo_outs/full.h5ad"), dest = "h5seurat")
neunpos <- LoadH5Seurat(paste0(data_dir, "velo_outs/full.h5seurat"))

Cow <- readRDS("../Species_Initial/Cow_initial.rds")
Cow_mat <- cbind(Cow@assays$RNA@counts, neunpos@assays$matrix@counts)

new_Cow <- CreateSeuratObject(Cow_mat)
new_Cow@meta.data$orig.file <- "cow_neunpos_B1"
new_Cow@meta.data[rownames(Cow@meta.data), "orig.file"] <- Cow@meta.data$orig.file

Cow <- new_Cow
```



```{r}
VlnPlot(Cow, features = "nCount_RNA", pt.size = 0, group.by = "orig.file")
VlnPlot(Cow, features = "nFeature_RNA", pt.size = 0, group.by = "orig.file")
```

```{r}
saveRDS(Cow, "../Species_Initial/Cow_initial.rds")
```

## Cell class annotation

```{r}
Cow <- readRDS("../Species_Initial/Cow_initial.rds")
Cow <- ClusterSeurat(Cow, numPCs = 20)
Cow <- FindClusters(Cow, resolution = 0.3)
saveRDS(Cow, file = "../Species_Initial/Cow_initial.rds")
```


Visualize initial clusters
```{r}
DimPlot(Cow, label = TRUE)
DimPlot(Cow, group.by = "orig.file")
VlnPlot(Cow, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Cow, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```


Look for major retinal cell classes
```{r}
Cow <- DendroOrder(Cow)

RGC_markers= c("RBPMS2","RBPMS", "SLC17A6", "THY1", "NEFL", "NEFM", "SNCG")
BC_markers=c("VSX2", "OTX2", "GRM6", "TRPM1", "VSX1", "CABP5", "GRIK1", "PRKCA")
AC_markers=c("POU6F2", "TFAP2A", "TFAP2B", "TFAP2C", "GAD1", "GAD2", "SLC6A9",  "C1QL1", "C1QL2")
HC_markers=c("ONECUT1", "LHX1", "CALB1", "TPM3")
Cone_markers=c("PDE6H", "CRX", "ARR3")
Rod_markers=c("SAG", "PDC", "RHO")
MG_markers=c("SLC1A3","RLBP1", "APOE")
Other_markers=c("S100B", "GFAP", "C1QA", "C1QB", "CLDN5", "IGFBP7")


DotPlot(Cow, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), group.by = "dendro_order") + RotatedAxis()

```

```{r}
plot(Cow@tools$BuildClusterTree)
```


Annotate clusters using DE genes and Dot plot of marker genes
```{r}
old_Cow <- readRDS("../Species_Initial/Cow_initial.rds")
Cow@meta.data$cell_class = "N/A"
Cow@meta.data[rownames(old_Cow@meta.data), "cell_class"] <- old_Cow@meta.data$cell_class
table(Cow@meta.data$dendro_order, Cow@meta.data$cell_class)
```


```{r}
Idents(Cow) <- "dendro_order"
Cow@meta.data$cell_class = "N/A"
Cow@meta.data[WhichCells(Cow, idents =c(35)),]$cell_class = "RGC"
Cow@meta.data[WhichCells(Cow, idents = c(39,0,20,14,27,15,12)),]$cell_class = "BP"
Cow@meta.data[WhichCells(Cow, idents = c(23)),]$cell_class = "HC"
Cow@meta.data[WhichCells(Cow, idents = c(5)),]$cell_class = "Rod"
Cow@meta.data[WhichCells(Cow, idents = c(33)),]$cell_class = "Cone"
Cow@meta.data[WhichCells(Cow, idents = c(17)),]$cell_class = "MG"
Cow@meta.data[WhichCells(Cow, idents = c(29,24,18,37,32,22,25,11,9,19,13,4,2,31,16,40,41,3,34)),]$cell_class = "GabaAC"
Cow@meta.data[WhichCells(Cow, idents = c(30,10, 26,36,7,28)),]$cell_class = "GlyAC"
Cow@meta.data[WhichCells(Cow, idents = c(6, 21, 38, 8, 1)),]$cell_class = "Other"

DimPlot(Cow, group.by = "cell_class")
```

Investigate unmarked clusters to see if they have any unique markers
```{r}

Cow_genes <- readRDS("../Species_Markers/Cow_genes_initial.rds")

# 17, 14: call them BP from GRIK1 and CABP5
# 37: PR. All top DE genes are selectively expressed in Rods in Mouse. Potentially a BP doublet, but annotate as PR
# 18, 19, 21, 34, 30, 11, 29, 10, 7, 12, 28
markers <- FindMarkers(Cow, ident.1 = c(21), max.cells.per.ident = 500)
ordered <- markers[order(markers$avg_log2FC, decreasing = TRUE), ]

DotPlot(Mouse, features = head(rownames(markers), 20), group.by = "cell_class") + RotatedAxis()

DotPlot(Mouse, features = head(rownames(ordered), 20), group.by = "cell_class") + RotatedAxis()
```

```{r}
counts <- table(Cow@meta.data$cell_class, Cow@meta.data$orig.file)
counts

saveRDS(Cow, "../Species_Initial/Cow_initial.rds")
saveRDS(Cow@meta.data, "../Metadata/Cow_metadata.rds")
```


## RGC clustering analysis
Create separate RGC object
```{r}
Cow <- readRDS("../Species_Initial/Cow_initial.rds")

# Only keep RGCs
Idents(Cow) <- "cell_class"
Cow_RGC <- subset(Cow, cells = WhichCells(Cow, idents ="RGC"))

# Remove cells with low and high counts or features
VlnPlot(Cow_RGC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Cow_RGC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

```

```{r}
Cow_RGC <- subset(Cow_RGC, subset = (orig.file == "CowNEUN") | (orig.file == "cowneun") | (orig.file == "CowNeuN1") | (orig.file == "CowNeuN2") | (orig.file == "cow_neunpos_B1)"))


saveRDS(Cow_RGC, file = "../Species_Objects/CowRGC_v1.rds")
```

Run clustering pipeline

```{r}
Cow_RGC <- ClusterSeurat(Cow_RGC, integrate.by = "orig.file", cluster_resolution = .8, numPCs = 20)
saveRDS(Cow_RGC, "../Species_Objects/CowRGC_integrated_v1.rds")

```

Visualize initial clusters
```{r}
# Cow_RGC <- readRDS("~/Species_Objects/CowRGC_integrated_v1.rds")
DimPlot(Cow_RGC, label = TRUE)
DimPlot(Cow_RGC, group.by = "orig.file", cells = sample(colnames(Cow_RGC)))
VlnPlot(Cow_RGC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Cow_RGC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

Examine markers for suspicious clusters, remove contaminant clusters
```{r}
Cow_RGC_markers <- FindAllMarkers(Cow_RGC, max.cells.per.ident = 1000)
```

Determine if any clusters need to be merged
```{r}

```


## BC clustering analysis
Create separate BC object
```{r}
Cow <- readRDS("../Species_Objects/Cow_initial.rds")

# Only keep BCs
Idents(Cow) <- "cell_class"
Cow_BC <- subset(Cow, cells = WhichCells(Cow, idents ="BP"))

# Remove cells with low and high counts or features
VlnPlot(Cow_BC, "nCount_RNA", group.by = "orig.file", pt.size = 0)
VlnPlot(Cow_BC, "nFeature_RNA", group.by = "orig.file", pt.size = 0)


```

```{r}
Cow_BC <- subset(Cow_BC, subset = nFeature_RNA > 100 & nCount_RNA < 10000)

# Only keep cells from chx10-enriched samples
Cow_BC <- subset(Cow_BC, subset = (orig.file == "cowchx10") | (orig.file == "CowCHx10") | (orig.file == "CowCHx10_r2") )


saveRDS(Cow_BC, file = "../Species_Objects/CowBC_v2.rds")

```

Run clustering pipeline

```{r}
Cow_BC <- ClusterSeurat(Cow_BC,  numPCs = 20, integrate.by = "orig.file")
saveRDS(Cow_BC, "../Species_Objects/CowBC_integrated_v2.rds")
```

Visualize initial clusters
```{r}
# Cow_BC <- readRDS("~/Species_Objects/CowBC_integrated_v2.rds")
DimPlot(Cow_BC, label = TRUE)
DimPlot(Cow_BC, group.by = "orig.file", cells = sample(colnames(Cow_BC)))
VlnPlot(Cow_BC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Cow_BC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

Adjust resolution parameter
```{r}
DefaultAssay(Cow_BC) <- "integrated"
Cow_BC <- FindClusters(Cow_BC, resolution = 1.4)

DimPlot(Cow_BC, label = TRUE, group.by = "integrated_snn_res.0.5")
DimPlot(Cow_BC, label = TRUE, group.by = "integrated_snn_res.0.8")
DimPlot(Cow_BC, label = TRUE, group.by = "integrated_snn_res.1.1")
DimPlot(Cow_BC, label = TRUE, group.by = "integrated_snn_res.1.4")

```

```{r}
Cow_BC@meta.data$seurat_clusters <- Cow_BC@meta.data$integrated_snn_res.0.8
Idents(Cow_BC) <- "integrated_snn_res.0.8"
DefaultAssay(Cow_BC) <- "RNA"

DotPlot(Cow_BC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), assay = "RNA") + RotatedAxis()

markers15 <- FindMarkers(Cow_BC, ident.1 = 15)

```

```{r}
Cow_BC <- DropClusters(Cow_BC, idents = 19, refactor = FALSE)
# Markers for 15 include photoreceptor markers, expresses at a much higher level
Cow_BC <- DropClusters(Cow_BC, idents = 15, refactor = FALSE)
Cow_BC <- MergeClusters(Cow_BC, idents = 0:5, refactor = TRUE)

Cow_BC@meta.data$type <- Cow_BC@meta.data$seurat_clusters

DimPlot(Cow_BC, label = TRUE, group.by = "type")
```

Double check whether some clusters should be merged
```{r}
# Merge 4 and 17: markers for 4 are expressed in 17 as well
Cow_BC <- MergeClusters(Cow_BC, idents = c(4,17), refactor = TRUE)

# 5 and 8 have enough markers to distinguish

ident.1 = 5
ident.2 = 8
mark_dif <- FindMarkers(Cow_BC, ident.1 = ident.1, ident.2 = ident.2)
mark_dif <- mark_dif[order(mark_dif$avg_log2FC),]

DotPlot(Cow_BC, idents = c(ident.1, ident.2), features = c(head(rownames(mark_dif)), tail(rownames(mark_dif)))) + RotatedAxis()

mark.1 <- FindMarkers(Cow_BC, ident.1 = ident.1, max.cells.per.ident = 1000)
mark.2 <- FindMarkers(Cow_BC, ident.1 = ident.2, max.cells.per.ident = 1000)
DotPlot(Cow_BC, features = unique(c(head(rownames(mark.1)), head(rownames(mark.2))))) + RotatedAxis()
```

```{r}
Cow_BC@meta.data$type <- Cow_BC@meta.data$seurat_clusters

DimPlot(Cow_BC, label = TRUE, group.by = "type")
```



```{r}
DefaultAssay(Cow_BC) <- "RNA"
Cow_BC <- DendroOrder(Cow_BC)

pan_BC <- c("VSX2", "CABP5", "OTX2")
Rod <- c("PRKCA", "GRAMD1B")
Off <- c("GRIK1")
On <- c("GRM6", "ISL1")

DotPlot(Cow_BC, features = c(pan_BC, Rod, Off, On), group.by = "dendro_order", assay = "RNA") + RotatedAxis()

Cow_BC@meta.data$on_off <- factor(Cow_BC@meta.data$seurat_clusters, levels = rev(c(1,2, 4,5,7,8,9,10,11,3,6,12,13,14,15,16)))

DotPlot(Cow_BC, features = c(pan_BC, Rod, Off, On), group.by = "on_off", assay = "RNA") + RotatedAxis()

```

```{r}
saveRDS(Cow_BC, "../Species_Objects/CowBC_integrated_v3.rds")
```

#Photoreceptors
```{r}
Idents(Cow) <- "cell_class"
Cow_PR <- subset(Cow, cells = WhichCells(Cow, idents = c("Rod", "Cone")))

# Remove cells with low and high counts or features
VlnPlot(Cow_PR, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Cow_PR, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

Cow_PR <- subset(Cow_PR, (orig.file == "cow_neunpos_B1") | (orig.file == "cowchx10") | (orig.file == "CowCHx10_r2") | (orig.file == "cowneun") | (orig.file == "CowNEUN") | (orig.file == "CowNeuN1") | (orig.file == "CowNeuN2"))
Cow_PR <- ClusterSeurat(Cow_PR,  numPCs = 20, integrate.by = "orig.file")

DimPlot(Cow_PR, label = TRUE)
DimPlot(Cow_PR, group.by = "orig.file", cells = sample(colnames(Cow_PR)))
VlnPlot(Cow_PR, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Cow_PR, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Cow_PR, features = Rod_markers, assay = "RNA") + RotatedAxis()
DotPlot(Cow_PR, features = Cone_markers, assay = "RNA") + RotatedAxis()
DotPlot(Cow_PR, group.by = "type", features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "OPN1SW", "OPN1LW", "OPN"), assay = "RNA") + RotatedAxis()

Cow_PR <- DropClusters(Cow_PR, idents = c(3,5), refactor = FALSE)
Cow_PR <- MergeClusters(Cow_PR, idents = c(0,1,6), refactor = FALSE)
DimPlot(Cow_PR, label = TRUE)
Cow_PR@meta.data$type = "ML_cone"
Idents(Cow_PR) <- "seurat_clusters"
Cow_PR@meta.data[WhichCells(Cow_PR, idents = 4), "type"] <- "S_cone"
Cow_PR@meta.data[WhichCells(Cow_PR, idents = 0), "type"] <- "Rod"
DimPlot(Cow_PR, group.by = "type", label = TRUE)
saveRDS(Cow_PR, file = "../Species_Objects/Cow_PR_v1.rds")

```

#HC
```{r}
Idents(Cow) <- "cell_class"
Cow_HC <- subset(Cow, cells = WhichCells(Cow, idents = c("HC")))

# Remove cells with low and high counts or features
VlnPlot(Cow_HC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Cow_HC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

Cow_HC <- subset(Cow_HC, (orig.file == "cow_neunpos_B1") |  (orig.file == "CowCHx10_r2"))
Cow_HC <- ClusterSeurat(Cow_HC,  numPCs = 25, integrate.by = "orig.file")

DimPlot(Cow_HC, label = TRUE)
DimPlot(Cow_HC, group.by = "orig.file", cells = sample(colnames(Cow_HC)))
VlnPlot(Cow_HC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Cow_HC, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Cow_HC, features = HC_markers, assay = "RNA") + RotatedAxis()
DotPlot(Cow_HC, group.by = "type", features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers, "ISL1"), assay = "RNA") + RotatedAxis()

Cow_HC <- DropClusters(Cow_HC, idents = 3, refactor = FALSE)
Cow_HC <- MergeClusters(Cow_HC, idents = c(0,1), refactor = FALSE)
Cow_HC <- MergeClusters(Cow_HC, idents = c(2,4), refactor = FALSE)
DimPlot(Cow_HC, label = TRUE)
Cow_HC <- FindNeighbors(Cow_HC, dims = 1:25)
Cow_HC <- RunUMAP(Cow_HC, dims = 1:25)

Cow_HC@meta.data$type = "HC1"
Idents(Cow_HC) <- "seurat_clusters"
Cow_HC@meta.data[WhichCells(Cow_HC, idents = 2), "type"] <- "H2"
Cow_HC@meta.data[WhichCells(Cow_HC, idents = 0), "type"] <- "H1"
DimPlot(Cow_HC, group.by = "type", label = TRUE)
saveRDS(Cow_HC, file = "../Species_Objects/Cow_HC_v2.rds")
```

