---
title: "R Notebook"
output: html_notebook
---

# Load libraries
```{r}
library(stringr)
library(tidyverse)
library(reshape2)
# library(ggtree)
# library(velocyto.R)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
source("../utils/utilFxns.R")
source("../utils/plottingFxns.R")
```


# Load velocyto data
Set data directory and load loom files
```{r}
data_dir <- "/Users/joshhahn/Google Drive/My Drive/shekharlab_data/projects/HumanSn/"



Convert(paste0(data_dir, "Hu032616OD_macula_NeuNPos/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu035516OS_macula_NeuNPosS1/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu035516OS_macula_NeuNPosS2/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu056316OD_macula_NeuNPos/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu056416OS_macula_NeuNPos/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu082219_macular_All/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu086916OD_macula_NeuNPos/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu088716OS_macula_NeuNPos/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu105916OD_macula_NeuNPos/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu218OSmAll/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu218OSmRGC/velo_outs/full.h5ad"), dest = "h5seurat")

#
Convert(paste0(data_dir, "Hu218OSPeriRetina/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu220235OSmAll/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu220OSmRGC/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu220OSPeriRetina/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu235OSmRGC/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "Hu235OSPeriRetina/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "HuCMixS1/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "HuCMixS2/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "HuGMixS1/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "HuGMixS3/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "HuGPt1398/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "HuPRet1059fc564mc/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "HuPRet326mc355fc/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "HuPRet563mc/velo_outs/full.h5ad"), dest = "h5seurat")

Convert(paste0(data_dir, "HuPRet887mc869fc/velo_outs/full.h5ad"), dest = "h5seurat")

```

Generate count matrices by adding intronic and exonic reads
```{r}
files <- c("Hu035516OS_macula_NeuNPosS1", "Hu035516OS_macula_NeuNPosS2", "Hu056316OD_macula_NeuNPos", "Hu056416OS_macula_NeuNPos", "Hu082219_macular_All", "Hu086916OD_macula_NeuNPos", "Hu088716OS_macula_NeuNPos", "Hu105916OD_macula_NeuNPos", "Hu218OSmAll", "Hu218OSmRGC", "Hu218OSPeriRetina",  "Hu220235OSmAll", "Hu220OSmRGC", "Hu220OSPeriRetina", "Hu235OSmRGC", "Hu235OSPeriRetina", "HuCMixS1", "HuCMixS2", "HuGMixS1", "HuGMixS3",  "HuGPt1398", "HuPRet1059fc564mc", "HuPRet326mc355fc", "HuPRet563mc", "HuPRet887mc869fc")

first_file <- "Hu032616OD_macula_NeuNPos"
obj <- LoadH5Seurat(paste0(data_dir, first_file, "/velo_outs/full.h5seurat"))
obj_mat <- obj@assays$spliced@counts + obj@assays$unspliced@counts
colnames(obj_mat) <- paste0(file, ".", colnames(obj_mat))
Human_mat <- obj_mat

for(file in files){
  obj <- LoadH5Seurat(paste0(data_dir, file, "/velo_outs/full.h5seurat"))
  obj_mat <- obj@assays$spliced@counts + obj@assays$unspliced@counts
  colnames(obj_mat) <- paste0(file, ".", colnames(obj_mat))
  Human_mat <- cbind(Human_mat, obj_mat)
}

```

Create Seurat object and set file information
```{r}
Human <- CreateSeuratObject(Human_mat, names.delim = ":")
Human@meta.data$orig.file <- str_sub(Human@meta.data$orig.ident, 0, -28)


saveRDS(Human, "../Species_Initial/HumanSn_initial.rds")
```

```{r}
VlnPlot(Human, features = "nCount_RNA", pt.size = 0, group.by = "orig.file")
VlnPlot(Human, features = "nFeature_RNA", pt.size = 0, group.by = "orig.file")
```


# Cell class annotation
```{r}
Human <- ClusterSeurat(Human)
  
```

Visualize initial clusters
```{r}
DimPlot(Human, label = TRUE)
DimPlot(Human, group.by = "orig.file", cells = sample(colnames(Human)))
VlnPlot(Human, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Human, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

Look for major retinal cell classes
```{r}
Human <- DendroOrder(Human)
#Human <- UpperCase_genes(Human)

RGC_markers= c("RBPMS", "RBPMS2", "SLC17A6", "POU6F2", "THY1", "NEFL", "NEFM", "SNCG")
BC_markers= c("VSX2", "OTX2", "GRM6", "TRPM1", "VSX1", "CABP5", "GRIK1", "PRKCA")
AC_markers= c("TFAP2A", "TFAP2B", "TFAP2C", "GAD1", "GAD2",  "SLC6A9", "C1QL1", "C1QL2")
HC_markers= c("ONECUT2","ONECUT1", "LHX1", "CALB1", "TPM3")
Cone_markers= c("PDE6H", "CRX", "ARR3")
Rod_markers= c("SAG", "PDC", "RHO")
MG_markers= c("SLC1A3","RLBP1", "APOE")
Other_markers= c("S100B", "GFAP", "C1QA", "C1QB", "CLDN5", "IGFBP7")

DotPlot(Human, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), group.by = "dendro_order") + RotatedAxis()
plot(Human@tools$BuildClusterTree)

```


```{r}
Human@meta.data$cell_class = "nope"
Human@meta.data[WhichCells(Human, idents =c(2,9,27,3,0,15)),]$cell_class = "RGC"
Human@meta.data[WhichCells(Human, idents = c(21,26,8)),]$cell_class = "BP"
Human@meta.data[WhichCells(Human, idents = c(20)),]$cell_class = "HC"
Human@meta.data[WhichCells(Human, idents = c(18,22,12,17)),]$cell_class = "GlyAC"

Human@meta.data[WhichCells(Human, idents = c(7,23,4,11,6,13)),]$cell_class = "GabaAC"
Human@meta.data[WhichCells(Human, idents = c(1,16,25,28)),]$cell_class = "Rod"
Human@meta.data[WhichCells(Human, idents = c(19)),]$cell_class = "Cone"

Human@meta.data[WhichCells(Human, idents = c(29,10)),]$cell_class = "MG"
Human@meta.data[WhichCells(Human, idents = c(24,5,14)),]$cell_class = "Other"

DimPlot(Human, group.by = "cell_class", raster = TRUE)


saveRDS(Human, "../Species_Initial/HumanSn_initial.rds")
saveRDS(Human@meta.data, "../Metadata/HumanSn_metadata.rds")

```



### Import cluster labels
```{r}
Human <- readRDS("../Species_Initial/HumanSn_initial.rds")

meta <- read.csv("../Species_Reference/retina_macula_periphery_meta.csv")
rownames(meta) <- meta$X
meta$NAME <- NULL

# Format velocyto labels to match appropriately
barcode <- function(string){
  file <- Human@meta.data[string, "orig.file"]
  code <- strsplit(string, c(":"))
  barcode <- strsplit(code[[1]][2], "x")
  if(file == "H2FoveaS1"){
    file <- "H2Fovea1S1"
  }
  if(file == "H2FoveaS2"){
    file <- "H2Fovea2S1"
  }
  name <- paste0(file, "_", barcode[[1]][1], "-1")
  return(name)
}
Human@meta.data$barcode <- str_sub(colnames(Human), -17, -2)
Human@meta.data$new_code <- paste0(Human@meta.data$orig.file, "_", Human@meta.data$barcode, "-1")
Human@meta.data$old_code <- rownames(Human@meta.data)

cells_in <- Human@meta.data$new_code %in% rownames(meta)
cell_names <- Human@meta.data$new_code[cells]

Human@meta.data$annotated <- NA
Human@meta.data[cells_in, "annotated_class"] <- meta[cell_names, "class"]
Human@meta.data[cells_in, "annotated_type"] <- meta[cell_names, "type"]
Human@meta.data[cells_in, "tissue"] <- meta[cell_names, "tissue"]
Human@meta.data[cells_in, "donor"] <- meta[cell_names, "donor"]

Human@meta.data$old_code <- NULL

# Remove cells not present in the atlas
Human@meta.data$keep <- !is.na(Human@meta.data$annotated_class)
Idents(Human) <- "keep"
Human <- subset(Human, idents = TRUE)
Human@meta.data$keep <- NULL

saveRDS(Human, "../Species_Initial/Human_initial.rds")
```



### Fovea RGC, BC
```{r}
Human <- readRDS("../Species_Initial/Human_initial.rds")

Fovea <- subset(Human, tissue == "MaculaCtrl")
Fovea@meta.data$type <- factor(Fovea@meta.data$annotated_type, levels = unique(Fovea@meta.data$annotated_type))
Fovea@meta.data$annotated_type <- NULL
Idents(Fovea) <- "type"
saveRDS(Fovea, "../Species_Initial/HumanFovea_initial.rds")
saveRDS(Fovea@meta.data, "../Metadata/HumanFovea_metadata.rds")

# Separate into RGCs, BCs
Fovea_RGC <- subset(Fovea, annotated_class == "RGC")
Fovea_RGC@meta.data$type <- factor(Fovea_RGC@meta.data$type, levels = unique(Fovea_RGC@meta.data$type))
saveRDS(Fovea_RGC, "../Species_Objects/HumanFovea_RGC_ann_v2.rds")
saveRDS(Fovea_RGC@meta.data, "../Metadata/RGC/HumanFovea_RGC_ann_v2_metadata.rds")

Fovea_BC <- subset(Fovea, annotated_class %in% c("BC-ON", "BC-OFF"))
Fovea_BC@meta.data$type <- factor(Fovea_BC@meta.data$type, levels = unique(Fovea_BC@meta.data$type))
saveRDS(Fovea_BC, "../Species_Objects/HumanFovea_BC_ann_v2.rds")
saveRDS(Fovea_BC@meta.data, "../Metadata/BC/HumanFovea_BC_ann_v2_metadata.rds")
```

### Periphery RGC, BC
```{r}
Periphery <- subset(Human, tissue == "PeripheryCtrl")
Periphery@meta.data$type <- factor(Periphery@meta.data$annotated_type, levels = unique(Periphery@meta.data$annotated_type))
Periphery@meta.data$annotated_type <- NULL
Idents(Periphery) <- "type"
saveRDS(Periphery, "../Species_Initial/HumanPeriphery_initial.rds")
saveRDS(Periphery@meta.data, "../Metadata/HumanPeriphery_metadata.rds")

# Separate into RGCs, BCs
Periphery_RGC <- subset(Periphery, annotated_class == "RGC")
Periphery_RGC@meta.data$type <- factor(Periphery_RGC@meta.data$type, levels = unique(Periphery_RGC@meta.data$type))
saveRDS(Periphery_RGC, "../Species_Objects/HumanPeriphery_RGC_ann_v2.rds")
saveRDS(Periphery_RGC@meta.data, "../Metadata/RGC/HumanPeriphery_RGC_ann_v2_metadata.rds")

Periphery_BC <- subset(Periphery, annotated_class %in% c("BC-ON", "BC-OFF"))
Periphery_BC@meta.data$type <- factor(Periphery_BC@meta.data$type, levels = unique(Periphery_BC@meta.data$type))
saveRDS(Periphery_BC, "../Species_Objects/HumanPeriphery_BC_ann_v2.rds")
saveRDS(Periphery_BC@meta.data, "../Metadata/BC/HumanPeriphery_BC_ann_v2_metadata.rds")
```


Fovea




# Old code I don't want to delete yet

### Integrate RGC and BC objects

Batch correct Human RGCs and run clustering pipeline

```{r}
Human_RGC <- readRDS("../Species_Objects/HumanRGC_ann.rds")

obj.list <- SplitObject(Human_RGC, split.by = "animal")
for (i in 1:length(obj.list)) {
      obj.list[[i]] <- NormalizeData(obj.list[[i]], verbose = FALSE)
      obj.list[[i]] <- FindVariableFeatures(obj.list[[i]], selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}
obj.anchors <- FindIntegrationAnchors(object.list = obj.list)
Human_RGC <- IntegrateData(anchorset = obj.anchors)
DefaultAssay(Human_RGC) <- "integrated"

Human_RGC <- ScaleData(Human_RGC, verbose = FALSE)
Human_RGC <- RunPCA(Human_RGC, npcs = 30, verbose = FALSE)
Human_RGC <- RunUMAP(Human_RGC, reduction = "pca", dims = 1:30)

DimPlot(Human_RGC, group.by = "annotated", label = TRUE)

saveRDS(Human_RGC, "~/Species_Objects/HumanRGC_int_ann_v1.rds")
```

Batch correct Human BCs and run clustering pipeline

```{r}
Human_BC <- readRDS("../Species_Objects/HumanBC_ann.rds")

obj.list <- SplitObject(Human_BC, split.by = "animal")
for (i in 1:length(obj.list)) {
      obj.list[[i]] <- NormalizeData(obj.list[[i]], verbose = FALSE)
      obj.list[[i]] <- FindVariableFeatures(obj.list[[i]], selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}
obj.anchors <- FindIntegrationAnchors(object.list = obj.list)
Human_BC <- IntegrateData(anchorset = obj.anchors)
DefaultAssay(Human_BC) <- "integrated"

Human_BC <- ScaleData(Human_BC, verbose = FALSE)
Human_BC <- RunPCA(Human_BC, npcs = 30, verbose = FALSE)
Human_BC <- RunUMAP(Human_BC, reduction = "pca", dims = 1:30)

DimPlot(Human_BC, group.by = "annotated", label = TRUE)

saveRDS(Human_RGC, "~/Species_Objects/HumanRGC_int_ann_v1.rds")
```


### Convert UMAP and diagonal gene plots to files
```{r}
Human_RGC <- readRDS("../Species_Objects/HumanRGC_int_ann_v1.rds")
Human_BC <- readRDS("../Species_Objects/HumanBC_int_ann_v1.rds")

Humanv2 <- readRDS("../Species_Objects/Human_initial_annotated.rds")

Idents(Human_RGC) <- "annotated"
Idents(Human_BC) <- "annotated"

pdf("../Figures/Species_UMAPs/HumanRGC.pdf", w=4, h=4, useDingbats = FALSE)
DimPlot(Human_RGC, label = TRUE) + NoLegend()
dev.off()

pdf("../Figures/Species_UMAPs/HumanBC.pdf", w=4, h=4, useDingbats = FALSE)
DimPlot(Human_BC, label = TRUE) + NoLegend()
dev.off()
```

Save as Fovea
```{r}
Human_RGC <- readRDS("../Species_Objects/HumanRGC_int_ann_v1.rds")
Human_BC <- readRDS("../Species_Objects/HumanBC_int_ann_v1.rds")

saveRDS(Human_RGC, "../Species_Objects/HumanFovea_RGC_int_ann_v1.rds")
saveRDS(Human_BC, "../Species_Objects/HumanFovea_BC_int_ann_v1.rds")
```


## Periphery
### Load velocyto data
Read in velocyto matrices
```{r}
data_dir <- "/Users/joshhahn/Google Drive File Stream/My Drive/shekharlab_data/projects/Human/Velocyto/"

Convert(paste0(data_dir, "H1CD73dpS1/full.h5ad"), dest = "h5seurat")
H1CD73S1 <- LoadH5Seurat(paste0(data_dir, "H1CD73dpS1/full.h5seurat"))

Convert(paste0(data_dir, "H1CD90S1/full.h5ad"), dest = "h5seurat")
H1CD90S1 <- LoadH5Seurat(paste0(data_dir, "H1CD90S1/full.h5seurat"))

Convert(paste0(data_dir, "H3CD73dpS1/full.h5ad"), dest = "h5seurat")
H3CD73S1 <- LoadH5Seurat(paste0(data_dir, "H3CD73dpS1/full.h5seurat"))

Convert(paste0(data_dir, "H3CD73dpS2/full.h5ad"), dest = "h5seurat")
H3CD73S2 <- LoadH5Seurat(paste0(data_dir, "H3CD73dpS2/full.h5seurat"))

Convert(paste0(data_dir, "H3CD90S1/full.h5ad"), dest = "h5seurat")
H3CD90S1 <- LoadH5Seurat(paste0(data_dir, "H3CD90S1/full.h5seurat"))

Convert(paste0(data_dir, "H3CD90S2/full.h5ad"), dest = "h5seurat")
H3CD90S2 <- LoadH5Seurat(paste0(data_dir, "H3CD90S2/full.h5seurat"))

```

Combine matrices and create new Seurat object
```{r}
Human_mat <- cbind(H1CD73S1@assays$matrix@counts, H1CD90S1@assays$matrix@counts, H3CD73S1@assays$matrix@counts, H3CD73S2@assays$matrix@counts, H3CD90S1@assays$matrix@counts, H3CD90S2@assays$matrix@counts)

Human_peri <- CreateSeuratObject(Human_mat, names.delim = ":")

Human_peri@meta.data$orig.file = "N/A"
Human_peri@meta.data[colnames(H1CD73S1), 'orig.file'] = "Human1CD73dpS1"
Human_peri@meta.data[colnames(H1CD90S1), 'orig.file'] = "Human1CD90S1"
Human_peri@meta.data[colnames(H3CD73S1), 'orig.file'] = "Per2CD73dpS1"
Human_peri@meta.data[colnames(H3CD73S2), 'orig.file'] = "Per2CD73dpS2"
Human_peri@meta.data[colnames(H3CD90S1), 'orig.file'] = "Per2CD90S1"
Human_peri@meta.data[colnames(H3CD90S2), 'orig.file'] = "Per2CD90S2"

saveRDS(Human_peri, "../Species_Objects/HumanPeriphery_initial.rds")
```

### Import cluster labels
```{r}
Human_peri <- readRDS("../Species_Objects/HumanPeriphery_initial.rds")

Human_peri@meta.data$library.ID = "N/A"
Human_peri@meta.data[colnames(H1CD73S1), 'library.ID'] = "H1CD73dpS1"
Human_peri@meta.data[colnames(H1CD90S1), 'library.ID'] = "H1CD90S1"
Human_peri@meta.data[colnames(H3CD73S1), 'library.ID'] = "H3CD73dpS1"
Human_peri@meta.data[colnames(H3CD73S2), 'library.ID'] = "H3CD73dpS2"
Human_peri@meta.data[colnames(H3CD90S1), 'library.ID'] = "H3CD90S1"
Human_peri@meta.data[colnames(H3CD90S2), 'library.ID'] = "H3CD90S2"

meta <- read.csv("../Species_Reference/Human_retina_combined_all_meta.csv")
rownames(meta) <- meta$NAME
meta$NAME <- NULL

# Format velocyto labels to match appropriately
barcode <- function(string){
  libID <- Human_peri@meta.data[string, "library.ID"]
  code <- strsplit(string, c(":"))
  barcode <- strsplit(code[[1]][2], "x")
  name <- paste0(libID, "_", barcode[[1]][1], "-1")
  return(name)
}
new_names <- unlist(lapply(colnames(Human_peri), barcode))
Human_peri@meta.data$barcode <- new_names
Human_peri@meta.data$annotated <- meta[Human_peri@meta.data$barcode, "Cluster"]

# Remove cells not present in the atlas
Human_peri@meta.data$keep <- !is.na(Human_peri@meta.data$annotated)
Idents(Human_peri) <- "keep"
Human_peri <- subset(Human_peri, idents = TRUE)
Human_peri@meta.data$keep <- NULL
```


### Reformat cell class
```{r}
Human_peri <- readRDS("../Species_Objects/HumanPeriphery_initial.rds")
Idents(Human_peri) <- "annotated"

Human_peri@meta.data$cell_class <- "Other"

Human_peri@meta.data[WhichCells(Human_peri, idents = paste0("Gaba", 1:10)), "cell_class"] <- "GabaAC"
Human_peri@meta.data[WhichCells(Human_peri, idents = paste0("Gly", 1:8)), "cell_class"] <- "GlyAC"
Human_peri@meta.data[WhichCells(Human_peri, idents = c("RB1", "DB4", "DB5*", "DB6", "BB+GB*","IMB","DB2","DB3b","DB3a","FMB","DB1","OFFx")), "cell_class"] <- "BP"
Human_peri@meta.data[WhichCells(Human_peri, idents = c("H1", "H2")), "cell_class"] <- "HC"
Human_peri@meta.data[WhichCells(Human_peri, idents = "Muller"), "cell_class"] <- "MG"
Human_peri@meta.data[WhichCells(Human_peri, idents = "MicroGlia"), "cell_class"] <- "MicroG"
Human_peri@meta.data[WhichCells(Human_peri, idents = c(paste0("RGC", 5:12), "MG_ON", "MG_OFF", "PG_ON", "PG_OFF")), "cell_class"] <- "RGC"
Human_peri@meta.data[WhichCells(Human_peri, idents = c("mlCones", "sCones")), "cell_class"] <- "Cone"
Human_peri@meta.data[WhichCells(Human_peri, idents = c("Rods")), "cell_class"] <- "Rod"
Human_peri@meta.data[WhichCells(Human_peri, idents = c("Endothelium", "Astrocytes")), "cell_class"] <- "Other"

saveRDS(Human_peri, "../Species_Objects/HumanPeriphery_initial.rds")
```

### Separate into RGC objects and BC objects
```{r}
Human_peri <- readRDS("../Species_Objects/HumanPeriphery_initial_annotated.rds")

Idents(Human_peri) <- "annotated"
Human_peri_RGC <- subset(Human_peri, idents = c(paste0("RGC", 5:12), "MG_ON", "MG_OFF", "PG_ON", "PG_OFF"))

saveRDS(Human_peri_RGC, "../Species_Objects/HumanPeriphery_RGC_ann.rds")

Human_peri_BC <- subset(Human_peri, idents = c("RB1", "DB4", "DB5*", "DB6", "BB+GB*","IMB","DB2","DB3b","DB3a","FMB","DB1","OFFx"))

saveRDS(Human_peri_BC, "../Species_Objects/HumanPeriphery_BC_ann.rds")
```

### Cluster + UMAP 
Note: Did NOT perform batch correction, as only raw RNA counts are used for downstream purposes
```{r}
Human_peri_RGC <- readRDS("../Species_Objects/HumanPeriphery_RGC_ann.rds")
Human_peri_RGC <- ClusterSeurat(Human_peri_RGC)
Idents(Human_peri_RGC) <- "annotated"
saveRDS(Human_peri_RGC, "../Species_Objects/HumanPeriphery_RGC_ann_v1.rds")


Human_peri_BC <- readRDS("../Species_Objects/HumanPeriphery_BC_ann.rds")
Human_peri_BC <- ClusterSeurat(Human_peri_BC)
Idents(Human_peri_BC) <- "annotated"
saveRDS(Human_peri_BC, "../Species_Objects/HumanPeriphery_BC_ann_v1.rds")


```



## Rods
```{r}
Human <- readRDS("../Species_Objects/Human_initial_annotated.rds")

# Subset to rods, do preliminary clustering, and save
Idents(Human) <- "annotated"
Human_Rods <- subset(Human, cells = WhichCells(Human, idents = "Rods"))
Human_Rods <- ClusterSeurat(Human_Rods)
saveRDS(Human_Rods, file = "../Species_Objects/HumanRods_v1.rds")
```


## DE Test
### RGC
```{r}
RGC_Fov <- readRDS("../Species_Objects/HumanRGC_int_ann_v1.rds")
de_rgc <- FindAllMarkers(RGC_Fov, assay = "RNA", only.pos = TRUE)
saveRDS(de_rgc, "../Species_Markers/HumanRGC_markers.rds")

pct_exp = ExpressionByCluster(RGC_Fov, de_rgc$gene)
write.csv(de_rgc, "../Species_Markers/ForKS/HumanRGC_all_de.csv")
write.csv(pct_exp, "../Species_Markers/ForKS/HumanRGC_all_pctexp.csv")
```

```{r}
de_rgc <- readRDS("../Species_Markers/HumanRGC_markers.rds")
de_rgc = de_rgc[de_rgc$pct.2 < .2,]
pct_exp = ExpressionByCluster(RGC_Fov, de_rgc$gene)
write.csv(de_rgc, "../Species_Markers/ForKS/HumanRGC_otherlow_de.csv")
write.csv(pct_exp, "../Species_Markers/ForKS/HumanRGC_otherlow_pctexp.csv")
```

```{r}
de_rgc <- readRDS("../Species_Markers/HumanRGC_markers.rds")
de_rgc = de_rgc[(de_rgc$avg_log2FC > .4 & de_rgc$pct.1 > .4),]
pct_exp = ExpressionByCluster(RGC_Fov, de_rgc$gene)
write.csv(de_rgc, "../Species_Markers/ForKS/HumanRGC_highexp_de.csv")
write.csv(pct_exp, "../Species_Markers/ForKS/HumanRGC_highexp_pctexp.csv")
```

```{r}
de_rgc <- readRDS("../Species_Markers/HumanRGC_markers.rds")
pct_exp = ExpressionByCluster(RGC_Fov, de_rgc$gene)

pct_exp <- as.data.frame(pct_exp)
rownames(pct_exp) <- rownames(de_rgc)


# Calculate number of clusters that express over a certain threshold and subset accordingly
threshold = .05
num_clusters = 3
pct_exp$num = rowSums(pct_exp > threshold)
pct_exp <- pct_exp[pct_exp$num < (num_clusters+1) , ]
de_rgc <- de_rgc[rownames(pct_exp),]

# See how many genes are in each cluster
table(de_rgc$cluster)
sum(table(de_rgc$cluster) == 0)

# Save
write.csv(de_rgc, "../Species_Markers/ForKS/HumanRGC_selective_de.csv")
write.csv(pct_exp, "../Species_Markers/ForKS/HumanRGC_selective_pctexp.csv")
```

## Recluster fovea
### RGCs

```{r}
HumanF <- readRDS("../Species_Initial/HumanFovea_initial.rds")
HumanP <- readRDS("../Species_Initial/HumanPeriphery_initial.rds")
```



Cluster
```{r}
Fovea_RGC <- readRDS("../Species_Objects/HumanRGC_int_ann_v1.rds")
DefaultAssay(Fovea_RGC) <- "integrated"
Fovea_RGC <- FindNeighbors(Fovea_RGC, dims = 1:30)
Fovea_RGC <- FindClusters(Fovea_RGC, resolution = 1.2)
DimPlot(Fovea_RGC, group.by = "integrated_snn_res.0.5", label = TRUE)
```

Investigate markers
```{r}
Idents(Fovea_RGC) <- "integrated_snn_res.0.5"
markers <- FindAllMarkers(Fovea_RGC, assay="RNA")
exp_markers <- subset(markers, pct.2 < .5)
e_top_markers <- TopMarkers(exp_markers, num_markers = 5)
top_markers <- TopMarkers(markers, num_markers = 5)
DotPlot(Fovea_RGC, features = top_markers, assay = "RNA") + RotatedAxis()
```

Merge clusters
```{r}
Fovea_RGC@meta.data$seurat_clusters <- Fovea_RGC@meta.data$integrated_snn_res.0.5
Fovea_RGC <- MergeClusters(Fovea_RGC, c(0,3,5,7), refactor = FALSE)
Fovea_RGC <- MergeClusters(Fovea_RGC, c(1,2,4,6), refactor = TRUE)

DimPlot(Fovea_RGC, group.by = "seurat_clusters")

Fovea_RGC@meta.data$reclustered <- Fovea_RGC@meta.data$seurat_clusters

```

Confusion matrix
```{r}
re_table <- table(Fovea_RGC@meta.data$reclustered, Fovea_RGC@meta.data$annotated)
MakePrettyConfusionMatrix(re_table)
```

### BCs
Cluster
```{r}
Fovea_BC <- readRDS("../Species_Objects/HumanBC_int_ann_v1.rds")
DefaultAssay(Fovea_BC) <- "integrated"
Fovea_BC <- FindNeighbors(Fovea_BC, dims = 1:30)
Fovea_BC <- FindClusters(Fovea_BC, resolution = 0.8)
DimPlot(Fovea_BC, group.by = "integrated_snn_res.0.8", label = TRUE)
```

Investigate markers
```{r}
Idents(Fovea_BC) <- "integrated_snn_res.0.8"
markers <- FindAllMarkers(Fovea_BC, assay="RNA")
exp_markers <- subset(markers, pct.2 < .5)
e_top_markers <- TopMarkers(exp_markers, num_markers = 5)
top_markers <- TopMarkers(markers, num_markers = 5)
DotPlot(Fovea_BC, features = top_markers, assay = "RNA") + RotatedAxis()
```

Merge clusters
```{r}
Fovea_BC@meta.data$seurat_clusters <- Fovea_BC@meta.data$integrated_snn_res.0.8
Fovea_BC <- MergeClusters(Fovea_BC, c(0,13), refactor = TRUE)

DimPlot(Fovea_BC, group.by = "seurat_clusters", label = TRUE)

Fovea_BC@meta.data$reclustered <- Fovea_BC@meta.data$seurat_clusters

```

Confusion matrix
```{r}
re_table <- table(Fovea_BC@meta.data$reclustered, Fovea_BC@meta.data$annotated)
MakePrettyConfusionMatrix(re_table)
```

```{r}
Idents(Fovea_BC) <- "reclustered"
markers1_16 <- FindMarkers(Fovea_BC, ident.1 = 1, ident.2 = 16)
DotPlot(Fovea_BC, features = head(rownames(markers1_16), 10), assay = "RNA", idents = c(1,16)) + RotatedAxis()
DotPlot(Fovea_BC, features = head(rownames(subset(markers1_16, avg_log2FC > 0)), 10), assay = "RNA", idents = c(1,16)) + RotatedAxis()

markers3_13 <- FindMarkers(Fovea_BC, ident.1 = 3, ident.2 = 13)
DotPlot(Fovea_BC, features = head(rownames(markers3_13), 20), assay = "RNA", idents = c(3,13)) + RotatedAxis()

markers7_15 <- FindMarkers(Fovea_BC, ident.1 = 7, ident.2 = 15)
DotPlot(Fovea_BC, features = head(rownames(markers7_15), 20), assay = "RNA", idents = c(7,15)) + RotatedAxis()

```


## Scratch
```{r}
library(readxl)
Human <- readRDS("../Species_Objects/Human_initial_annotated.rds")
csm <- as.data.frame(read_excel("../Human PR Surface Markers/csm_final.xlsx"))
csm <- csm$gene

# Figure out which genes are in the dataset
genes <- c()
for (gene in csm){
  gene <- toupper(gene)
  if(gene %in% rownames(Human)){
    genes <- c(genes, gene)
  }
  else{
    print(paste0("Missing gene: ", gene))
    search <- readline(prompt="Enter name to search: ")
    print(grep(search, rownames(Human), ignore.case = TRUE, value = TRUE))
    include <- toupper(readline(prompt="Gene to include (type s to skip): "))
    if(include %in% rownames(Human)){
    genes <- c(genes, gene)
    }
  }
}

genes <- c(genes, "DSCAM", "FCGR2B", "FCGR1B", "FCGR3B")
genes <- unique(toupper(genes))
saveRDS(genes, "../Human PR Surface Markers/csm_edited.rds")
```

```{r}
Idents(Human) <- "cell_class"
DefaultAssay(Human) <- "RNA"
genes <- genes[genes %in% rownames(Human)]
PR_markers <- FindMarkers(Human, ident.1 = "PR", assay = "RNA", features = genes)

PR_markers$gene <- rownames(PR_markers)
PR_markers <- select(PR_markers, gene, p_val, avg_log2FC, pct.1, pct.2)
PR_markers <- PR_markers[order(-PR_markers$avg_log2FC),]

write_csv(PR_markers, "../Human PR Surface Markers/full_DE.csv")
saveRDS(PR_markers, "../Human PR Surface Markers/full_DE.rds")
```

```{r}
PR_markers <- readRDS("../Human PR Surface Markers/full_DE.rds")
selective <- subset(PR_markers, pct.2 < .1)
write_csv(selective, "../Human PR Surface Markers/selective_DE.csv")


PR_markers[order(-PR_markers$avg_log2FC),]
```



## RGCs from hand annotation
```{r}
Human <- readRDS("../Species_Initial/HumanFovea_initial.rds")
Idents(Human) <- "assigned"
RGC <- subset(Human, idents = "RGC")
RGC@meta.data[is.na(RGC@meta.data$annotated), "annotated"] <- "NA"

RGC <- ClusterSeurat(RGC, numPCs = 20, integrate.by = "orig.file")

```

Remove non-RGCs
```{r}
RGC_markers= c("RBPMS", "SLC17A6", "POU6F2")
DotPlot(RGC, features = RGC_markers)
RGC <- DropClusters(RGC, idents = c(0,4,5))
```

Increase cluster resolution. Stick with 0.8
```{r}
DefaultAssay(RGC) <- "integrated"

numPCs <- 13
RGC <- FindNeighbors(RGC, dims = 1:numPCs)
RGC <- RunUMAP(RGC, dims = 1:numPCs)

RGC <- FindClusters(RGC, resolution = 0.5)
RGC <- FindClusters(RGC, resolution = 0.8)
RGC <- FindClusters(RGC, resolution = 1.1)

DimPlot(RGC, label = TRUE, group.by = "integrated_snn_res.0.5")
DimPlot(RGC, label = TRUE, group.by = "integrated_snn_res.0.8")
DimPlot(RGC, label = TRUE, group.by = "integrated_snn_res.1.1")
```


```{r}
DefaultAssay(RGC) <- "RNA"
DimPlot(RGC, label = TRUE)
DimPlot(RGC, group.by = "annotated", label = TRUE)

VlnPlot(RGC, features = "nCount_RNA")
VlnPlot(RGC, features = "nFeature_RNA")

```


```{r}
RGC_markers= c("RBPMS2","RBPMS", "SLC17A6", "POU6F2", "THY1", "NEFL", "NEFM", "SNCG")
BC_markers=c("VSX2", "OTX2", "GRM6", "TRPM1", "VSX1", "CABP5", "GRIK1", "PRKCA")
AC_markers=c("TFAP2A", "TFAP2B", "TFAP2C", "GAD1", "GAD2", "SLC6A9", "GRIK2", "C1QL1", "C1QL2")
HC_markers=c("ONECUT1", "LHX1", "CALB1", "TPM3")
Cone_markers=c("PDE6H", "CRX", "ARR3")
Rod_markers=c("SAG", "PDC", "RHO")
MG_markers=c("SLC1A3","RLBP1", "APOE")
Other_markers=c("S100B", "GFAP", "C1QA", "C1QB", "CLDN5", "IGFBP7")

Idents(RGC) <- "integrated_snn_res.1.1"

DotPlot(RGC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers)) + RotatedAxis()

```

Remove clusters
```{r}
# 12, 13 express PR markers
# 14 expresses BP markers
# 23 is rods
# 17 is MG
RGC <- DropClusters(RGC, idents = c(12,13,14,17, 23))
```

```{r}
DefaultAssay(RGC) <- "RNA"
DimPlot(RGC, label = TRUE)
DimPlot(RGC, group.by = "annotated", label = TRUE)
```

Merge Clusters
```{r}
RGC <- MergeClusters(RGC, idents = c(1,3,6,7))
RGC <- MergeClusters(RGC, idents = c(2,4,5,9), refactor = TRUE)
RGC <- MergeClusters(RGC, idents = c(3,4,8), refactor = TRUE)

```

```{r}
DimPlot(RGC, label = TRUE)
DimPlot(RGC, group.by = "annotated", label = TRUE)
```
```{r}
RGC@meta.data$type <- RGC@meta.data$seurat_clusters
saveRDS(RGC, "../Species_Objects/HumanFovea_RGC_byhand.rds")
```

# 5HT Receptors
```{r}
Mouse <- readRDS("../Species_Objects/MouseRGC_integrated_v2.rds")
Human <- readRDS("../Species_Objects/HumanFovea_RGC_int_ann_v1.rds")
H_Peri <- readRDS("../Species_Objects/HumanPeriphery_RGC_ann_v1.rds")
Macaque <- readRDS("../Species_Objects/MacaqueFovea_RGC_velo_ann_v1.rds")
Mac_Peri <- readRDS("../Species_Objects/MacaquePeri_RGC_velo_ann_v1.rds")
Marmoset <- readRDS("../Species_Objects/MarmosetFovea_RGC_int_v3.rds")
Peri <- readRDS("../Species_Objects/MarmosetPeriphery_RGC_int_v2.rds")


genes <- c("HTR1A", "HTR1B","HTR1D", "HTR1E", "HTR1F")

DotPlot(Mouse, features = genes, assay = "RNA", group.by = "annotated") + RotatedAxis()
DotPlot(Human, features = genes, assay = "RNA", group.by = "annotated") + RotatedAxis()
DotPlot(H_Peri, features = genes, assay = "RNA", group.by = "annotated") + RotatedAxis()

DotPlot(Macaque, features = genes, assay = "RNA", group.by = "annotated") + RotatedAxis()
DotPlot(Mac_Peri, features = genes, assay = "RNA", group.by = "annotated") + RotatedAxis()

DotPlot(Marmoset, features = genes, assay = "RNA", group.by = "annotated") + RotatedAxis()
DotPlot(Peri, features = genes, assay = "RNA", group.by = "annotated") + RotatedAxis()





```
