---
title: "R Notebook"
output: html_notebook
---

## Preprocessing 

Load libraries
```{r}
library(reshape2)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
source("../utils/utilFxns.R")
source("../utils/plottingFxns.R")
```

Set data directory and load loom files
```{r}
data_dir <- "/Users/joshhahn/Google Drive/My Drive/shekharlab_data/projects/Lizard/"
 

#Convert(paste0(data_dir, "LizChx10high1/velo_outs/full.h5ad"), dest = "h5seurat")
Chx10_1 <- LoadH5Seurat(paste0(data_dir, "LizChx10high1/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "LizChx10high2/velo_outs/full.h5ad"), dest = "h5seurat")
Chx10_2 <- LoadH5Seurat(paste0(data_dir, "LizChx10high2/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "LizardNEUNpos/velo_outs/full.h5ad"), dest = "h5seurat")
NEUNpos <- LoadH5Seurat(paste0(data_dir, "LizardNEUNpos/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "LizardNEUNneg/velo_outs/full.h5ad"), dest = "h5seurat")
NEUNneg <- LoadH5Seurat(paste0(data_dir, "LizardNEUNneg/velo_outs/full.h5seurat"))

#Convert(paste0(data_dir, "S14_Anole/velo_outs/full.h5ad"), dest = "h5seurat")
S14 <- LoadH5Seurat(paste0(data_dir, "S14_Anole/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "LizChx10/velo_outs/full.h5ad"), dest = "h5seurat")
Chx10 <- LoadH5Seurat(paste0(data_dir, "LizChx10/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "LizNueun1/velo_outs/full.h5ad"), dest = "h5seurat")
Nueun1 <- LoadH5Seurat(paste0(data_dir, "LizNueun1/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "LizNueun2/velo_outs/full.h5ad"), dest = "h5seurat")
Nueun2 <- LoadH5Seurat(paste0(data_dir, "LizNueun2/velo_outs/full.h5seurat"))

```


LizChx10high1 / 2 are low quality, do not include. 
```{r}
Lizard_mat <- cbind(NEUNpos@assays$matrix@counts, NEUNneg@assays$matrix@counts, S14@assays$matrix@counts, Chx10@assays$matrix@counts, Nueun1@assays$matrix@counts, Nueun2@assays$matrix@counts)

Lizard <- CreateSeuratObject(Lizard_mat, names.delim = ":")

Lizard@meta.data[colnames(NEUNpos), 'orig.file'] = "LizardNEUNpos"
Lizard@meta.data[colnames(NEUNneg), 'orig.file'] = "LizardNEUNneg"
Lizard@meta.data[colnames(S14), 'orig.file'] = "S14_Anole"
Lizard@meta.data[colnames(Chx10), 'orig.file'] = "LizChx10"
Lizard@meta.data[colnames(Nueun1), 'orig.file'] = "LizNueun1"
Lizard@meta.data[colnames(Nueun2), 'orig.file'] = "LizNueun2"


```



```{r}
VlnPlot(Lizard, features = "nCount_RNA", pt.size = 0, group.by = "orig.file", y.max = 10000)
VlnPlot(Lizard, features = "nFeature_RNA", pt.size = 0, group.by = "orig.file")
```

```{r}
saveRDS(Lizard, file = "../Species_Initial/Lizard_initial.rds")
```

## Cell class annotation

```{r}
Lizard <- readRDS("../Species_Initial/Lizard_initial.rds")
Lizard <- ClusterSeurat(Lizard, numPCs = 20, cluster_resolution = 0.5)
saveRDS(Lizard, file = "../Species_Objects/Lizard_initial.rds")
```


Visualize initial clusters
```{r}
DimPlot(Lizard, label = TRUE)
DimPlot(Lizard, group.by = "orig.file")
VlnPlot(Lizard, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Lizard, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

```{r}

```

Look for major retinal cell classes
```{r}
Lizard <- DendroOrder(Lizard)

RGC_markers= tolower(c("RBPMS", "RBPMS2", "SLC17A6", "POU6F2", "THY1", "NEFL", "NEFM", "SNCG"))
BC_markers=tolower(c("VSX2", "OTX2", "GRM6", "TRPM1", "VSX1", "CABP5", "GRIK1", "PRKCA"))
AC_markers=tolower(c("TFAP2A", "TFAP2B", "TFAP2C", "GAD1", "GAD2", "SLC6A9", "C1QL1", "C1QL2"))
HC_markers=tolower(c("ONECUT2","ONECUT3", "LHX1", "CALB1", "TPM3"))
Cone_markers=tolower(c("PDE6H", "CRX", "ARR3"))
Rod_markers=tolower(c("SAG", "PDC", "RHO"))
MG_markers=tolower(c("SLC1A3","RLBP1", "APOE"))
Other_markers=tolower(c("S100B", "GFAP", "C1QA", "C1QB", "CLDN5", "IGFBP7"))

DotPlot(Lizard, features = toupper(c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers)), group.by = "dendro_order") + RotatedAxis()
```

```{r}
plot(Lizard@tools$BuildClusterTree)
```


Remove clusters with low counts / features
```{r}
Lizard <- DropClusters(Lizard, idents = c(), refactor = FALSE)
```


Annotate clusters using DE genes and Dot plot of marker genes

```{r}
Lizard@meta.data$cell_class = "Nope"
Lizard@meta.data[WhichCells(Lizard, idents =c(25,4,5)),]$cell_class = "RGC"
Lizard@meta.data[WhichCells(Lizard, idents = c(24,14,20, 9, 12)),]$cell_class = "BP"
Lizard@meta.data[WhichCells(Lizard, idents = c(17)),]$cell_class = "HC"
Lizard@meta.data[WhichCells(Lizard, idents = c(26,19)),]$cell_class = "MG"
Lizard@meta.data[WhichCells(Lizard, idents = c(10)),]$cell_class = "Rod"
Lizard@meta.data[WhichCells(Lizard, idents = c(3,13,11)),]$cell_class = "Cone"

Lizard@meta.data[WhichCells(Lizard, idents = c(7)),]$cell_class = "GlyAC"
Lizard@meta.data[WhichCells(Lizard, idents = c(15, 23, 16, 8, 6, 22, 2)),]$cell_class = "GabaAC"
Lizard@meta.data[WhichCells(Lizard, idents = c(0,1, 18, 21)),]$cell_class = "Other"


DimPlot(Lizard, group.by = "cell_class")


```


Investigate unmarked clusters to see if they have any unique markers
```{r}
Mouse <- readRDS("../Species_Initial/Mouse_initial.rds")

# 
ident = 18
mark <- FindMarkers(Lizard, ident.1 = ident, max.cells.per.ident = 1000)
DotPlot(Mouse, features = toupper(head(rownames(mark), 20)), group.by = "cell_class") + RotatedAxis()

```

```{r}
DimPlot(Lizard, group.by = "cell_class")
DotPlot(Lizard, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), group.by = "cell_class") + RotatedAxis()
counts <- table(Lizard@meta.data$cell_class, Lizard@meta.data$orig.file)
counts
saveRDS(Lizard, "../Species_Initial/Lizard_initial.rds")
saveRDS(Lizard@meta.data, "../Metadata/Lizard_metadata.rds")

```

### Refine photoreceptors
```{r}
Lizard <- readRDS("../Species_Initial/Lizard_initial.rds")
Idents(Lizard) <- "cell_class"
PR <- subset(Lizard, idents = c("Rod", "Cone"))
DefaultAssay(PR) <- "RNA"
PR <- ClusterSeurat(PR)
DotPlot(PR, features = toupper(Rod_markers)) + RotatedAxis()
Rods <- WhichCells(PR, idents = 14)
PR_cells <- colnames(PR)

Lizard@meta.data[PR_cells, "cell_class"] <- "Cone"
Lizard@meta.data[Rods, "cell_class"] <- "Rod"

saveRDS(Lizard@meta.data, "../Metadata/Lizard_metadata.rds")
saveRDS(Lizard, "../Species_Initial/Lizard_initial.rds")
```

## RGC clustering analysis
Create separate RGC object
```{r}
Lizard <- readRDS("../Species_Initial//Lizard_initial.rds")


# Only keep RGCs
Idents(Lizard) <- "cell_class"
Lizard_RGC <- subset(Lizard, cells = WhichCells(Lizard, idents ="RGC"))

# Remove cells with low and high counts or features
VlnPlot(Lizard_RGC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Lizard_RGC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

```
```{r}
Lizard_RGC <- subset(Lizard_RGC, subset = nFeature_RNA > 500 & nCount_RNA < 10000)
Lizard_RGC <- subset(Lizard_RGC, subset = (orig.file == "LizardNEUNpos") | (orig.file == "LizNueun1") | (orig.file == "LizNueun2") | (orig.file == "S14_Anole"))



saveRDS(Lizard_RGC, file = "../Species_Objects/LizardRGC_v2.rds")
```

### Run clustering pipeline

```{r}
Lizard_RGC <- ClusterSeurat(Lizard_RGC, integrate.by = "orig.file", numPCs = 20)
```

### Visualize initial clusters
```{r}
# Lizard_RGC <- readRDS("~/Species_Objects/LizardRGC_integrated_v1.rds")
DimPlot(Lizard_RGC, label = TRUE)
DimPlot(Lizard_RGC, group.by = "orig.file", cells = sample(colnames(Lizard_RGC)))
VlnPlot(Lizard_RGC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Lizard_RGC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

Examine markers for suspicious clusters. There don't appear to be contaminant clusters.
```{r}
DotPlot(Lizard_RGC, features = tolower(RGC_markers), assay = "RNA") + RotatedAxis()

DotPlot(Lizard_RGC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), assay = "RNA") + RotatedAxis()
```



### Refine Clusters
```{r}
DefaultAssay(Lizard_RGC) <- "integrated"

Lizard_RGC <- FindClusters(Lizard_RGC, resolution = 1.4)

DimPlot(Lizard_RGC, label = TRUE, group.by = "integrated_snn_res.0.5")
DimPlot(Lizard_RGC, label = TRUE, group.by = "integrated_snn_res.0.8")
DimPlot(Lizard_RGC, label = TRUE, group.by = "integrated_snn_res.1.1")
DimPlot(Lizard_RGC, label = TRUE, group.by = "integrated_snn_res.1.4")

```
Proceed with resolution of 0.8 Determine if any clusters need to be merged.


```{r}
DefaultAssay(Lizard_RGC) <- "RNA"

Lizard_RGC@meta.data$seurat_clusters <- Lizard_RGC@meta.data$integrated_snn_res.0.8
Idents(Lizard_RGC) <- "seurat_clusters"


# 2 and 3: merge
Lizard_RGC <- MergeClusters(Lizard_RGC, idents = c(3,2), refactor = TRUE)



ident.1 = 2
ident.2 = 3
mark_dif <- FindMarkers(Lizard_RGC, ident.1 = ident.1, ident.2 = ident.2)
mark_dif <- mark_dif[order(mark_dif$avg_log2FC),]

DotPlot(Lizard_RGC, idents = c(ident.1, ident.2), features = c(head(rownames(mark_dif)), tail(rownames(mark_dif)))) + RotatedAxis()

mark.1 <- FindMarkers(Lizard_RGC, ident.1 = ident.1, max.cells.per.ident = 1000)
mark.2 <- FindMarkers(Lizard_RGC, ident.1 = ident.2, max.cells.per.ident = 1000)
DotPlot(Lizard_RGC, features = unique(c(head(rownames(mark.1)), head(rownames(mark.2))))) + RotatedAxis()
```

### Summary Plots
```{r}
Lizard_RGC@meta.data$type <- Lizard_RGC@meta.data$seurat_clusters
DimPlot(Lizard_RGC, label = TRUE, group.by = "type")
```

```{r}
Lizard_markers <- FindAllMarkers(Lizard_RGC, only.pos = TRUE, max.cells.per.ident = 1000)

Lizard_RGC <- DendroOrder(Lizard_RGC)
Idents(Lizard_RGC) <- "type"

PlotUniqueMarkers(Lizard_RGC, Lizard_markers)

DotPlot(Lizard_RGC, features = TopMarkers(Lizard_markers, num_markers = 2), group.by = "type") + RotatedAxis()


```

```{r}
DefaultAssay(Lizard_RGC) <- "RNA"
Idents(Lizard_RGC) <- "type"
saveRDS(Lizard_RGC, "../Species_Objects/LizardRGC_int_v2.rds")

```


## BC clustering analysis
```{r}
Lizard <- readRDS("../Species_Objects/Lizard_initial.rds")

# Only keep BCs
Idents(Lizard) <- "cell_class"
Lizard_BC <- subset(Lizard, cells = WhichCells(Lizard, idents ="BP"))

# Remove cells with low and high counts or features
VlnPlot(Lizard_BC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Lizard_BC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)


```

```{r}
# Remove cells from NEUNpos due to low number
Lizard_BC <- subset(Lizard_BC, subset = (orig.file != "LizardNEUNpos"))


saveRDS(Lizard_BC, file = "../Species_Objects/LizardBC_v1.rds")
LizBC <- readRDS("../Species_Objects/LizardBC_integrated_v1.rds")
```

### Run clustering pipeline

```{r}
Lizard_BC <- readRDS(file = "../Species_Objects/LizardBC_v1.rds")

Lizard_BC <- ClusterSeurat(Lizard_BC, integrate.by = "orig.file", numPCs = 20)
```

### Visualize initial clusters
```{r}
DimPlot(Lizard_BC, label = TRUE)
DimPlot(Lizard_BC, group.by = "orig.file", cells = sample(colnames(Lizard_BC)))
VlnPlot(Lizard_BC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Lizard_BC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```


Examine markers for suspicious clusters. 
```{r}
DotPlot(Lizard_BC, features = tolower(BC_markers), assay = "RNA") + RotatedAxis()

DotPlot(Lizard_BC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), assay = "RNA") + RotatedAxis()
```

### Refine Clusters
Remove cluster 0 since it doesn't express BC markers and has low counts
```{r}
Lizard_BC <- DropClusters(Lizard_BC, idents = 0)
```

Increase cluster resolution. Stick with 0.8
```{r}
DefaultAssay(Lizard_BC) <- "integrated"

numPCs <- 20
Lizard_BC <- FindNeighbors(Lizard_BC, dims = 1:numPCs)
Lizard_BC <- FindClusters(Lizard_BC, resolution = 0.8)
Lizard_BC <- RunUMAP(Lizard_BC, dims = 1:numPCs)

DimPlot(Lizard_BC, label = TRUE)
```

```{r}
DefaultAssay(Lizard_BC) <- "RNA"
Lizard_BC <- DendroOrder(Lizard_BC)

pan_BC <- tolower(c("VSX2", "CABP5", "OTX2"))
Rod <- tolower(c("PRKCA", "GRAMD1B"))
Cone <- tolower(c("SLC24A3", "APBA2B"))

Off <- tolower(c("GRIK1"))
On <- tolower(c("GRM6", "ISL1"))

DotPlot(Lizard_BC, features = c(pan_BC, Rod, Cone, Off, On), group.by = "dendro_order", assay = "RNA") + RotatedAxis()

Lizard_BC@meta.data$on_off <- factor(Lizard_BC@meta.data$seurat_clusters, levels = rev(c(4,8,3,6,7,10,11,13,15, 0, 1,2,5,9,12,14,16)))

DotPlot(Lizard_BC, features = c(pan_BC, Rod, Cone, Off, On), group.by = "on_off", assay = "RNA") + RotatedAxis()



```

Check for my cone markers
```{r}
Mouse_BC <- readRDS("../Species_Objects/MouseBC_int_ann_v3.rds")
Idents(Mouse_BC) <- "annotated"
DefaultAssay(Mouse_BC) <- "RNA"
rod_markers <- FindMarkers(Mouse_BC, ident.1 = "RBC")
rod_markers <- rod_markers[order(rod_markers$avg_log2FC),]
DotPlot(Mouse_BC, features = head(rownames(rod_markers), 20)) + RotatedAxis()

DotPlot(Lizard_BC, features = tolower(head(rownames(rod_markers), 20)), group.by = "on_off", assay = "RNA") + RotatedAxis()


```

```{r}
Lizard_BC@meta.data$type <- Lizard_BC@meta.data$seurat_clusters
Idents(Lizard_BC) <- "type"
DefaultAssay(Lizard_BC) <- "RNA"
saveRDS(Lizard_BC, "../Species_Objects/LizardBC_integrated_v1.rds")
```

### Second RB type
```{r}
Zfish_BC <- readRDS("../Species_Objects/ZebrafishBC_v3.rds")
Lizard_BC <- readRDS("../Species_Objects/LizardBC_integrated_v1.rds")


pan_BC <- c("VSX1", "CABP5A")
Rod <- c("PRKCA", "GRAMD1B")
Cone <- c("SLC24A3", "APBA2B")
Off <- c("GRIK1A", "GRIK1B")
On <- c("GRM6A","GRM6B", "ISL1", "GNB3A")

# Cluster 19 in zebrafish is second RB type
R2_markers <- FindMarkers(Zfish_BC, ident.1 = 19)
DotPlot(Zfish_BC, features = head(rownames(R2_markers)), assay = "RNA", group.by = "bipolar_order") + RotatedAxis()

DotPlot(Lizard_BC, features = tolower(head(rownames(R2_markers))), assay = "RNA", group.by = "on_off") + RotatedAxis()


```

```{r}
Mouse_BC <- readRDS("../Species_Objects/MouseBC_int_ann_v3.rds")

Idents(Mouse_BC) <- "annotated"
RB_markers <- FindMarkers(Mouse_BC, ident.1 = "RBC")
RB_markers <- RB_markers[order(RB_markers$avg_log2FC), ]


DotPlot(Mouse_BC, features = head(rownames(RB_markers), 20), assay = "RNA", group.by = "annotated") + RotatedAxis()

DotPlot(Lizard_BC, features = tolower(head(rownames(RB_markers), 20)), assay = "RNA", group.by = "on_off") + RotatedAxis()

DotPlot(Mouse_BC, features = tail(rownames(RB_markers), 20), assay = "RNA", group.by = "annotated") + RotatedAxis()

DotPlot(Lizard_BC, features = tolower(tail(rownames(RB_markers), 20)), assay = "RNA", group.by = "on_off") + RotatedAxis()




```


#Photoreceptors
```{r}
Lizard <- readRDS("../Species_Initial/Lizard_initial.rds")
Idents(Lizard) <- "cell_class"
Lizard_PR <- subset(Lizard, cells = WhichCells(Lizard, idents = c("Rod", "Cone")))

VlnPlot(Lizard_PR, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Lizard_PR, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

Lizard_PR <- subset(Lizard_PR, orig.file %in% c("LizardNEUNneg",  "LizardNEUNpos",  "LizChx10", "S14_Anole"))
Lizard_PR <- ClusterSeurat(Lizard_PR,  numPCs = 20, integrate.by = "orig.file")
saveRDS(Lizard_PR, "../Species_Objects/LizardPR_int_v1.rds")

DimPlot(Lizard_PR, label = TRUE)
DimPlot(Lizard_PR, group.by = "orig.file", cells = sample(colnames(Lizard_PR)))
VlnPlot(Lizard_PR, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Lizard_PR, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Lizard_PR, features = toupper(Rod_markers), assay = "RNA") + RotatedAxis()
DotPlot(Lizard_PR, features = toupper(Cone_markers), assay = "RNA") + RotatedAxis()
DotPlot(Lizard_PR, features = toupper(c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "OPN1SW", "OPN1LW", "OPN")), assay = "RNA") + RotatedAxis()
```

#HC
```{r}
Idents(Lizard) <- "cell_class"
Lizard_HC <- subset(Lizard, cells = WhichCells(Lizard, idents = "HC"))

VlnPlot(Lizard_HC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Lizard_HC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

Lizard_HC <- subset(Lizard_HC, orig.file %in% c("LizardNEUNneg", "LizChx10", "S14_Anole"))
Lizard_HC <- ClusterSeurat(Lizard_HC,  numPCs = 20, integrate.by = "orig.file")
saveRDS(Lizard_HC, "../Species_Objects/LizardHC_int_v1.rds")

DimPlot(Lizard_HC, label = TRUE)
DimPlot(Lizard_HC, group.by = "orig.file", cells = sample(colnames(Lizard_HC)))
VlnPlot(Lizard_HC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Lizard_HC, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Lizard_HC, features = toupper(HC_markers), assay = "RNA") + RotatedAxis()
DotPlot(Lizard_HC, features = toupper(c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers, "ISL1")), assay = "RNA") + RotatedAxis()
```

