---
title: "R Notebook"
output: html_notebook
---

## Preprocessing 

Load libraries
```{r}
library(reshape2)
library(Seurat)
library(SeuratDisk)
library(ggplot2)
source("../utils/utilFxns.R")
source("../utils/plottingFxns.R")
```

Set data directory and load loom files
```{r}
data_dir <- "/Users/joshhahn/Google Drive/My Drive/shekharlab_data/projects/Rhabdomys/"

Convert(paste0(data_dir, "RhabAllnoStb2/velo_outs/full.h5ad"), dest = "h5seurat")
All_nSt <- LoadH5Seurat(paste0(data_dir, "RhabAllnoStb2/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "RhabAllnoStnoSob2/velo_outs/full.h5ad"), dest = "h5seurat")
All_nStSo <- LoadH5Seurat(paste0(data_dir, "RhabAllnoStnoSob2/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "RhabNeunNeg1b2/velo_outs/full.h5ad"), dest = "h5seurat")
NN1b2 <- LoadH5Seurat(paste0(data_dir, "RhabNeunNeg1b2/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "RhabNeunNegb1/velo_outs/full.h5ad"), dest = "h5seurat")
NNb1 <- LoadH5Seurat(paste0(data_dir, "RhabNeunNegb1/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "RhabNeunPos1b1/velo_outs/full.h5ad"), dest = "h5seurat")
NP1b1 <- LoadH5Seurat(paste0(data_dir, "RhabNeunPos1b1/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "RhabNeunPos1b2/velo_outs/full.h5ad"), dest = "h5seurat")
NP1b2 <- LoadH5Seurat(paste0(data_dir, "RhabNeunPos1b2/velo_outs/full.h5seurat"))

Convert(paste0(data_dir, "RhabNeunPos2b1/velo_outs/full.h5ad"), dest = "h5seurat")
NP2b1 <- LoadH5Seurat(paste0(data_dir, "RhabNeunPos2b1/velo_outs/full.h5seurat"))
```

Combine matrices and create new Seurat object. 
```{r}
Rhabdomys_mat <- cbind(All_nSt@assays$matrix@counts, All_nStSo@assays$matrix@counts, NN1b2@assays$matrix@counts, NNb1@assays$matrix@counts, NP1b1@assays$matrix@counts, NP1b2@assays$matrix@counts, NP2b1@assays$matrix@counts)

rownames(Rhabdomys_mat) <- toupper(rownames(Rhabdomys_mat))

Rhabdomys <- CreateSeuratObject(Rhabdomys_mat, names.delim = ":")

Rhabdomys@meta.data[colnames(All_nSt), 'orig.file'] = "RhabAllnoStb2"
Rhabdomys@meta.data[colnames(All_nStSo), 'orig.file'] = "RhabAllnoStnoSob2"
Rhabdomys@meta.data[colnames(NN1b2), 'orig.file'] = "RhabNeunNeg1b2"
Rhabdomys@meta.data[colnames(NNb1), 'orig.file'] = "RhabNeunNegb1"
Rhabdomys@meta.data[colnames(NP1b1), 'orig.file'] = "RhabNeunPos1b1"
Rhabdomys@meta.data[colnames(NP1b2), 'orig.file'] = "RhabNeunPos1b2"
Rhabdomys@meta.data[colnames(NP2b1), 'orig.file'] = "RhabNeunPos2b1"

```


```{r}
VlnPlot(Rhabdomys, features = "nCount_RNA", pt.size = 0, group.by = "orig.file", y.max = 25000)
VlnPlot(Rhabdomys, features = "nFeature_RNA", pt.size = 0, group.by = "orig.file")
```

```{r}
saveRDS(Rhabdomys, file = "../Species_Initial/Rhabdomys_initial.rds")
```

## Cell class annotation

```{r}
Rhabdomys <- ClusterSeurat(Rhabdomys, numPCs = 20)
saveRDS(Rhabdomys, file = "../Species_Initial/Rhabdomys_initial.rds")
```


Visualize initial clusters
```{r}
DimPlot(Rhabdomys, label = TRUE)
DimPlot(Rhabdomys, group.by = "orig.file")
VlnPlot(Rhabdomys, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Rhabdomys, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```


Look for major retinal cell classes
```{r}
# Rhabdomys <- DendroOrder(Rhabdomys)

RGC_markers= c("RBPMS","RBPMS2", "SLC17A6", "POU6F2", "THY1", "NEFL", "NEFM", "SNCG")
BC_markers= c("VSX2", "OTX2", "TRPM1", "CABP5", "GRIK1", "PRKCA", "GRM6")
AC_markers= c("TFAP2A" ,"TFAP2B", "TFAP2C", "GAD1","GAD2", "SLC6A9", "C1QL1", "C1QL2")
HC_markers= c("ONECUT1", "ONECUT2","ONECUT3", "LHX1", "CALB1", "TPM3")
Cone_markers= c("PDE6H", "CRX", "ARR3")
Rod_markers= c("SAG", "PDC", "RHO")
MG_markers= c("SLC1A3","RLBP1", "APOE")
Other_markers= c("S100B", "GFAP", "C1QA", "C1QB", "CLDN5", "IGFBP7")

DotPlot(Rhabdomys, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), group.by = "dendro_order") + RotatedAxis()
```

```{r}
plot(Rhabdomys@tools$BuildClusterTree)
```

Annotate clusters using DE genes and Dot plot of marker genes

```{r}
Rhabdomys@meta.data$cell_class = "Nope"
Rhabdomys@meta.data[WhichCells(Rhabdomys, idents =c(7,10,0,2,4,5)),]$cell_class = "RGC"
Rhabdomys@meta.data[WhichCells(Rhabdomys, idents = c(9,19,27,29,8,23,25,32,11,15)),]$cell_class = "BP"
Rhabdomys@meta.data[WhichCells(Rhabdomys, idents = c(12)),]$cell_class = "HC"
Rhabdomys@meta.data[WhichCells(Rhabdomys, idents = c(24,30,16,26,34)),]$cell_class = "MG"
Rhabdomys@meta.data[WhichCells(Rhabdomys, idents = c(31,17)),]$cell_class = "GlyAC"
Rhabdomys@meta.data[WhichCells(Rhabdomys, idents = c(21,20,13,6)),]$cell_class = "GabaAC"
Rhabdomys@meta.data[WhichCells(Rhabdomys, idents = c(3)),]$cell_class = "Rod"
Rhabdomys@meta.data[WhichCells(Rhabdomys, idents = c(1,28,33)),]$cell_class = "Cone" 
Rhabdomys@meta.data[WhichCells(Rhabdomys, idents = c(22,14,18)),]$cell_class = "Other"

```


Investigate unmarked clusters to see if they have any unique markers
```{r}
Mouse <- readRDS("../Species_Initial/Mouse_initial.rds")

ortho3[grep("Tfap", ortho3$Gene.name), c("Rhabdomys.gene.name", "Rhabdomys.gene.stable.ID")]


# 
ident = c(26,31)
mark <- FindMarkers(Rhabdomys, ident.1 = ident, max.cells.per.ident = 1000)
ortho3[ortho3$Rhabdomys.gene.stable.ID %in% head(rownames(mark),10), "Gene.name"]

```

```{r}
DimPlot(Rhabdomys, group.by = "cell_class")
DotPlot(Rhabdomys, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), group.by = "cell_class") + RotatedAxis()
counts <- table(Rhabdomys@meta.data$cell_class, Rhabdomys@meta.data$orig.file)
counts
```

```{r}
saveRDS(Rhabdomys, "../Species_Initial/Rhabdomys_initial.rds")
saveRDS(Rhabdomys@meta.data, "../Metadata/Rhabdomys_metadata.rds")
```




## RGC clustering analysis
Create separate RGC object
```{r}
Rhabdomys <- readRDS("../Species_Initial/Rhabdomys_initial.rds")

# Only keep RGCs
Idents(Rhabdomys) <- "cell_class"
Rhabdomys_RGC <- subset(Rhabdomys, cells = WhichCells(Rhabdomys, idents ="RGC"))

# Remove cells with low and high counts or features
VlnPlot(Rhabdomys_RGC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Rhabdomys_RGC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

```


```{r}
# Only keep samples enriched for RGCs
Rhabdomys_RGC <- subset(Rhabdomys_RGC, (orig.file ==  "RhabNeunPos1b1") | (orig.file ==  "RhabNeunPos1b2") | (orig.file ==  "RhabNeunPos2b1"))

# Remove cells with too high counts
Rhabdomys_RGC <- subset(Rhabdomys_RGC, subset = nCount_RNA < 15000)

VlnPlot(Rhabdomys_RGC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Rhabdomys_RGC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

saveRDS(Rhabdomys_RGC, file = "../Species_Objects/RhabdomysRGC_v1.rds")
```

### Run clustering pipeline

```{r}
Rhabdomys_RGC <- ClusterSeurat(Rhabdomys_RGC, integrate.by = "orig.file", numPCs = 20)
saveRDS(Rhabdomys_RGC, file = "../Species_Objects/RhabdomysRGC_int_v1.rds")
```

### Visualize initial clusters
```{r}
DimPlot(Rhabdomys_RGC, label = TRUE)
DimPlot(Rhabdomys_RGC, group.by = "orig.file", cells = sample(colnames(Rhabdomys_RGC)))
VlnPlot(Rhabdomys_RGC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Rhabdomys_RGC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```

### Refine Clusters
Examine markers for suspicious clusters. 
```{r}
DotPlot(Rhabdomys_RGC, features = RGC_markers, assay = "RNA") + RotatedAxis()

DotPlot(Rhabdomys_RGC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), assay = "RNA") + RotatedAxis()
```


Try different resolution parameters
```{r}
DefaultAssay(Rhabdomys_RGC) <- "integrated"
Rhabdomys_RGC <- FindClusters(Rhabdomys_RGC, resolution = 1.4)

DimPlot(Rhabdomys_RGC, label = TRUE, group.by = "integrated_snn_res.0.5")
DimPlot(Rhabdomys_RGC, label = TRUE, group.by = "integrated_snn_res.0.8")
DimPlot(Rhabdomys_RGC, label = TRUE, group.by = "integrated_snn_res.1.1")
DimPlot(Rhabdomys_RGC, label = TRUE, group.by = "integrated_snn_res.1.4")
```

Start with 1.1, as it seems to find the most clusters without many likely spurious splits
```{r}
DefaultAssay(Rhabdomys_RGC) <- "RNA"
Idents(Rhabdomys_RGC) <- "integrated_snn_res.1.1"
Rhabdomys_RGC@meta.data$seurat_clusters <- Idents(Rhabdomys_RGC)
Rhabdomys_RGC <- DendroOrder(Rhabdomys_RGC)
```

Merge any clusters
```{r}
Idents(Rhabdomys_RGC) <- "dendro_order"
Rhab_RGC_markers <- FindAllMarkers(Rhabdomys_RGC, assay = "RNA", max.cells.per.ident = 1000,only.pos = TRUE)
```

```{r}
DotPlot(Rhabdomys_RGC, assay = "RNA", features = TopMarkers(Rhab_RGC_markers, 2)) + RotatedAxis()
```


```{r}
# Drop 34 due to low features
Rhabdomys_RGC <- DropClusters(Rhabdomys_RGC, idents = c(34), refactor = FALSE)

# Merge 1, 3, 19
Rhabdomys_RGC <- MergeClusters(Rhabdomys_RGC, idents = c(1,3,19), refactor = TRUE)

ident.1 = 15
ident.2 = 22
mark_dif <- FindMarkers(Rhabdomys_RGC, ident.1 = ident.1, ident.2 = ident.2)
mark_dif$pct.dif <- mark_dif$pct.1 - mark_dif$pct.2
mark_dif <- mark_dif[order(mark_dif$avg_log2FC),]
mark_dif <- mark_dif[order(mark_dif$pct.dif),]



DotPlot(Rhabdomys_RGC, idents = c(ident.1, ident.2), features = c(head(rownames(mark_dif)), tail(rownames(mark_dif)))) + RotatedAxis()

mark.1 <- subset(Rhab_RGC_markers, cluster == ident.1)
mark.2 <- subset(Rhab_RGC_markers, cluster == ident.2)
DotPlot(Rhabdomys_RGC, features = unique(c(head(mark.1$gene), head(mark.2$gene))), group.by = "dendro_order") + RotatedAxis()
```

### Summary Plots
```{r}
Rhabdomys_RGC@meta.data$type <- Rhabdomys_RGC@meta.data$seurat_clusters
Idents(Rhabdomys_RGC) <- "type"
DefaultAssay(Rhabdomys_RGC) <- "RNA"
DimPlot(Rhabdomys_RGC, label = TRUE, group.by = "type")
saveRDS(Rhabdomys_RGC, "../Species_Objects/RhabdomysRGC_int_v2.rds")
```


```{r}
Rhabdomys_markers <- FindAllMarkers(Rhabdomys_RGC, only.pos = TRUE, max.cells.per.ident = 1000)

Rhabdomys_RGC <- DendroOrder(Rhabdomys_RGC)
Idents(Rhabdomys_RGC) <- "dendro_order"

PlotUniqueMarkers(Rhabdomys_RGC, Rhabdomys_markers)

DotPlot(Rhabdomys_RGC, features = TopMarkers(Rhabdomys_markers, num_markers = 2), group.by = "dendro_order") + RotatedAxis()

```

```{r}
DefaultAssay(Rhabdomys_RGC) <- "RNA"
Idents(Rhabdomys_RGC) <- "type"
saveRDS(Rhabdomys_RGC, "../Species_Objects/RhabdomysRGC_int_v2.rds")

Rhabdomys_RGC <- readRDS( "../Species_Objects/RhabdomysRGC_int_v2.rds")



```


## BC clustering analysis
```{r}
Rhabdomys <- readRDS("../Species_Objects/Rhabdomys_initial.rds")

# Only keep BCs
Idents(Rhabdomys) <- "cell_class"
Rhabdomys_BC <- subset(Rhabdomys, cells = WhichCells(Rhabdomys, idents ="BP"))

# Remove cells with low and high counts or features
VlnPlot(Rhabdomys_BC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Rhabdomys_BC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)
```

```{r}
# Remove cells from NEUNpos samples
Rhabdomys_BC <- subset(Rhabdomys_BC, !((orig.file ==  "RhabNeunPos1b1") | (orig.file ==  "RhabNeunPos1b2") | (orig.file ==  "RhabNeunPos2b1")))

Rhabdomys_BC <- subset(Rhabdomys_BC, subset = nCount_RNA < 6000)

saveRDS(Rhabdomys_BC, file = "../Species_Objects/RhabdomysBC_v1.rds")
```

### Run clustering pipeline

```{r}
# Rhabdomys_BC <- readRDS(file = "../Species_Objects/RhabdomysBC_v1.rds")
Rhabdomys_BC <- ClusterSeurat(Rhabdomys_BC, integrate.by = "orig.file", numPCs = 20)
saveRDS(Rhabdomys_BC, "../Species_Objects/RhabdomysBC_int_v1.rds")
```

### Visualize initial clusters
```{r}
DimPlot(Rhabdomys_BC, label = TRUE)
DimPlot(Rhabdomys_BC, group.by = "orig.file", cells = sample(colnames(Rhabdomys_BC)))
VlnPlot(Rhabdomys_BC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Rhabdomys_BC, "nFeature_RNA", pt.size = 0) + RotatedAxis()
```


Examine markers for suspicious clusters. 
```{r}
DotPlot(Rhabdomys_BC, features = BC_markers, assay = "RNA") + RotatedAxis()

DotPlot(Rhabdomys_BC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, Other_markers), assay = "RNA") + RotatedAxis()
```

```{r}
# 13 expreses PR markers
Rhabdomys_BC <- DropClusters(Rhabdomys_BC, idents = c(13))
```

### Refine Clusters

Try cluster resolutions
```{r}
DefaultAssay(Rhabdomys_BC) <- "integrated"

numPCs <- 18
Rhabdomys_BC <- FindNeighbors(Rhabdomys_BC, dims = 1:numPCs)
Rhabdomys_BC <- RunUMAP(Rhabdomys_BC, dims = 1:numPCs)

Rhabdomys_BC <- FindClusters(Rhabdomys_BC, resolution = 1.1)

DimPlot(Rhabdomys_BC, label = TRUE, group.by = "integrated_snn_res.0.5")
DimPlot(Rhabdomys_BC, label = TRUE, group.by = "integrated_snn_res.0.8")
DimPlot(Rhabdomys_BC, label = TRUE, group.by = "integrated_snn_res.1.1")
```

Stick with 0.8, merge as needed
```{r}
Idents(Rhabdomys_BC) <- "integrated_snn_res.0.8"
Rhabdomys_BC@meta.data$seurat_clusters <- Idents(Rhabdomys_BC)
DefaultAssay(Rhabdomys_BC) <- "RNA"
```

```{r}
# 5 and 8 appear to be separate
# 1 and 7

# Merge 1, 3, 19
Rhabdomys_RGC <- MergeClusters(Rhabdomys_RGC, idents = c(1,3,19), refactor = TRUE)

ident.1 = 1
ident.2 = 7
mark_dif <- FindMarkers(Rhabdomys_BC, ident.1 = ident.1, ident.2 = ident.2)
mark_dif$pct.dif <- mark_dif$pct.1 - mark_dif$pct.2
mark_dif <- mark_dif[order(mark_dif$avg_log2FC),]
mark_dif <- mark_dif[order(mark_dif$pct.dif),]



DotPlot(Rhabdomys_BC, idents = c(ident.1, ident.2), features = c(head(rownames(mark_dif)), tail(rownames(mark_dif)))) + RotatedAxis()

mark.1 <- FindMarkers(Rhabdomys_BC, ident.1 = ident.1, only.pos = TRUE, max.cells.per.ident = 1000)
mark.2 <- subset(Rhab_RGC_markers, cluster == ident.2)
DotPlot(Rhabdomys_RGC, features = unique(c(head(mark.1$gene), head(mark.2$gene))), group.by = "dendro_order") + RotatedAxis()
```

```{r}
Rhabdomys_BC <- MergeClusters(Rhabdomys_BC, idents = c(), refactor = TRUE)
Rhabdomys_BC@meta.data$type <- Rhabdomys_BC@meta.data$seurat_clusters
Idents(Rhabdomys_BC) <- Rhabdomys_BC@meta.data$type
Rhabdomys_BC <- MergeClusters(Rhabdomys_BC, idents = c(0,1), refactor = TRUE)
Rhabdomys_BC@meta.data$type <- Rhabdomys_BC@meta.data$seurat_clusters
```


### Summary Plots
```{r}
pan_BC <- c("VSX2")
Rod <- c("PRKCA", "GRAMD1B")
Off <- c("GRIK1")
On <- c("GRM6", "ISL1")

DotPlot(Rhabdomys_BC, features = c(pan_BC, Rod, Off, On), group.by = "dendro_order", assay = "RNA") + RotatedAxis()

Rhabdomys_BC@meta.data$on_off <- factor(Rhabdomys_BC@meta.data$type, levels = rev(c(12, 5,6,7,9,11,13,14,16,17,1,2,3,4,8,10,15,18)))

DotPlot(Rhabdomys_BC, features = c(pan_BC, Rod, Off, On), group.by = "on_off", assay = "RNA") + RotatedAxis()
```


```{r}
saveRDS(Rhabdomys_BC, "../Species_Objects/RhabdomysBC_int_v2.rds")
Rhabdomys_BC <- readRDS("../Species_Objects/RhabdomysBC_int_v2.rds")

```
#Photoreceptor
```{r}
Idents(Rhabdomys) <- "cell_class"
Rhabdomys_PR <- subset(Rhabdomys, cells = WhichCells(Rhabdomys, idents = c("Rod", "Cone")))


# Remove cells with low and high counts or features
VlnPlot(Rhabdomys_PR, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Rhabdomys_PR, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

Rhabdomys_PR <- subset(Rhabdomys_PR, (orig.file == "RhabAllnoStb2") | (orig.file == "RhabAllnoStnoSob2") | (orig.file == "RhabNeunNeg1b2") | (orig.file == "RhabNeunNegb1") | (orig.file == "RhabNeunPos1b2"))
Rhabdomys_PR <- ClusterSeurat(Rhabdomys_PR,  numPCs = 40, integrate.by = "orig.file")

DimPlot(Rhabdomys_PR, label = TRUE)
DimPlot(Rhabdomys_PR, group.by = "orig.file", cells = sample(colnames(Rhabdomys_PR)))
VlnPlot(Rhabdomys_PR, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Rhabdomys_PR, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Rhabdomys_PR, features = Rod_markers, assay = "RNA") + RotatedAxis()
DotPlot(Rhabdomys_PR, features = Cone_markers, assay = "RNA") + RotatedAxis()
DotPlot(Rhabdomys_PR, group.by = "type", features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "OPN1SW", "OPN1LW", "OPN", "OPN5"), assay = "RNA") + RotatedAxis()

#3, 4, 5 merge (Rods)
#9, 10, 11 ML cone
#7 S cone
#0, 1, 2 L
#drop 6, 8, 12, 13
Rhabdomys_PR <- MergeClusters(Rhabdomys_PR, idents = c(3,4,5), refactor = FALSE)
Rhabdomys_PR <- DropClusters(Rhabdomys_PR, idents = c(6,8,12,13), refactor = FALSE)
Rhabdomys_PR <- MergeClusters(Rhabdomys_PR, idents = c(0,1,2,9,10,11), refactor = FALSE)

Rhabdomys_PR@meta.data$type = "ML_cone"
Idents(Rhabdomys_PR) <- "seurat_clusters"
Rhabdomys_PR@meta.data[WhichCells(Rhabdomys_PR, idents = 7), "type"] <- "S_cone"
Rhabdomys_PR@meta.data[WhichCells(Rhabdomys_PR, idents = 3), "type"] <- "Rod"
DimPlot(Rhabdomys_PR, group.by = "type", label = TRUE)

saveRDS(Rhabdomys_PR, file = "../Species_Objects/Rhabdomys_PR_v1")
```
#HC
```{r}
# Only keep HCs
Idents(Rhabdomys) <- "cell_class"
Rhabdomys_HC <- subset(Rhabdomys, cells = WhichCells(Rhabdomys, idents ="HC"))

#Remove cells with low and high counts or features
VlnPlot(Rhabdomys_HC, "nCount_RNA", group.by = "orig.file", pt.size = .1)
VlnPlot(Rhabdomys_HC, "nFeature_RNA", group.by = "orig.file", pt.size = .1)

Rhabdomys_HC <- ClusterSeurat(Rhabdomys_HC,  numPCs = 30, integrate.by = "orig.file")

DimPlot(Rhabdomys_HC, label = TRUE)
DimPlot(Rhabdomys_HC, group.by = "orig.file", cells = sample(colnames(Rhabdomys_HC)))
VlnPlot(Rhabdomys_HC, "nCount_RNA", pt.size = 0) + RotatedAxis()
VlnPlot(Rhabdomys_HC, "nFeature_RNA", pt.size = 0) + RotatedAxis()

DotPlot(Rhabdomys_HC, features = HC_markers, assay = "RNA") + RotatedAxis()
DotPlot(Rhabdomys_HC, features = c(RGC_markers, BC_markers, AC_markers, HC_markers, Cone_markers, Rod_markers, MG_markers, "ISL1"), assay = "RNA") + RotatedAxis()

FindMarkers(Rhabdomys_HC, ident.1 = "3", assay = "RNA")

Rhabdomys_HC <- DropClusters(Rhabdomys_HC, idents = c(1,3,4), refactor = FALSE)
Rhabdomys_HC <- MergeClusters(Rhabdomys_HC, idents = c(0,2), refactor = FALSE)
Rhabdomys_HC@meta.data$type = "H1"
Idents(Rhabdomys_HC) <- "seurat_clusters"
Rhabdomys_HC@meta.data[WhichCells(Rhabdomys_HC, idents = 0), "type"] <- "H1"
DimPlot(Rhabdomys_HC, group.by = "type", label = TRUE)
saveRDS(Rhabdomys_HC, file = "../Species_Objects/Rhabdomys_HC_v1.rds")


```

